### 2016-06-28 pres_solar_prod.tex
add files

% pres_solar_prod.tex

\documentclass{beamer}
\usepackage{url}
\usepackage{amsmath}
\usepackage{esdiff} %for writing partial derivatives
\usetheme{default}

\title{Lease or Buy?}
\subtitle{Quality differences and asymmetric information in California solar panels}
\author[J. Mauritzen]{Johannes Mauritzen \\ NHH Norwegian School of Economics}
\institute[NHH]{
 % Department of Business and Management Science\\
 % NHH Norwegian School of Economics\\[1ex]
  \texttt{jmaurit@gmail.com}
}

\date{\today}

\begin{document}

%--- the titlepage frame -------------------------%
\begin{frame}[plain]
  \titlepage
\end{frame}

%--- the presentation begins here ----------------%

\begin{frame}[plain]
	\begin{itemize}
	\item[] What's behind the surge in the California rooftop solar market? The role of Chinese panels, uncertainty and new business models.
	\item[] \url{jmaurit.github.io\#solar_lemons}
	\end{itemize}
\end{frame}

\begin{frame}[plain]
	\begin{figure}
	\includegraphics[width=1\textwidth]{figures/cost_over_time.png}
	%\caption{After about 2010, prices for solar systems fell dramatically and installations nearly tripled.}
	%\label{cost_over_time}	
	\end{figure}
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/top_contractors_plot.png}
		 %\caption{Marked concentration increased markedly over the period studied.  How did Solar City and Verengo gain market share?}
		%\label{top_contractors}	
	\end{figure}	
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=.8\textwidth]{figures/panel_makers_scty.png}
		 %\caption{The top panel makers used by the leading solar system contractor, Solar City. 2011 the company shifted to using panels from relatively unknown chinese manufacturers.}
		%\label{panel_makers_scty}	
	\end{figure}
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/scty_china_lease.png}
		 %\caption{Contractors that gained market share, notably Solar City, did so not only by switching to cheaper Chinese panels but also by moving to a leasing model of sales.}
		%\label{cont_market_share}	
	\end{figure}
\end{frame}

\begin{frame}[plain]
	\begin{figure}
	\includegraphics[width=1\textwidth]{figures/chinese_panel_switch.png}
	%\caption{It seems a switch to cheaper Chinese panels can at least partly explain the increased market share of Solar City and Verengo.  But why did so few other contractors adopt Chinese panels?}
	%\label{chinese_panel_switch}	
	\end{figure}	
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/solar_panel_review.png}
		%\caption{Websites are available providing reviews of various solar panels, however information is not available on new Chinese manufacturers that have not been on the market.}
		%\label{solar_panel_review}	
	\end{figure}	
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=.8\textwidth]{figures/solar_quality_nyt.png}
		%\caption{Websites are available providing reviews of various solar panels, however information is not available on new Chinese manufacturers that have not been on the market.}
		%\label{solar_panel_review}	
	\end{figure}
\end{frame}

\begin{frame}[plain]
	\begin{equation}
	\begin{split}
		lease_i = & invlogit(\alpha + \beta timeYears_i + nationality_i +\\ 
		& \sigma timeYears_i*nationality_i + \epsilon)
		%\label{equation:lease_model}
	\end{split}
	\end{equation}	
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/logit_plot.png}
		%\caption{Systems with Chinese panels, mostly absent from the market before 2010, were substantially more likely to be used in leased systems than panels from more established Japanese and German manufacturers.}
		%\label{logit_plot}
	\end{figure}	
\end{frame}

\begin{frame}[plain]
	\begin{equation}
		log(costPerKw_i) = \alpha + \gamma china_i + \tau lease_i + \beta timeYears_i + \sigma inter_i + \epsilon
		%\label{equation:slope1}
	\end{equation}	
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/cost_plot.png}
		 %\caption{Systems with Chinese panels are substantially cheaper, but only became widely popular after 2011.}
		%\label{cost_plot}
	\end{figure}
\end{frame}

\begin{frame}
	Quality and Asymmetric Info in Solar Panels
	\begin{itemize}
	\item ``Experience goods''
	\item Hard to ascertain quality, even after the fact
	\item One shot vs. repeat buyers
	\item Warranties may not provide much assurance
	\end{itemize}	
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=1\textwidth]{tot_production.png}
		% \caption{}
		%\label{tot_production}
	\end{figure}
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=1\textwidth]{example_prod.png}
		%\caption{}
		\label{example_prod}
	\end{figure}
\end{frame}

\begin{frame}
	Hierarchical Bayesian Model
	\begin{itemize}
	\item Natural modeling of structure of data
	\item Natural interpretation of uncertainty
	\item Flexibility of simulation methods
	\end{itemize}	
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=1\textwidth]{solar_prod_bayes_diag.png}		
		%\label{example_prod}
	\end{figure}
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=.8\textwidth]{figures/post_mu_b1.png}

		\label{post_mu_b1}
	\end{figure}
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=.8\textwidth]{figures/predicted_deg.png}
		\label{predicted_deg}
	\end{figure}
\end{frame}

\begin{frame}[plain]
\textbf{Main Points:}
	\begin{itemize}
	\item Cheaper Chinese panels AND the introduction of a leasing business model are directly linked to the surge in solar panel installations. 
	\end{itemize}
\end{frame}

\begin{frame}[plain]
\textbf{Main Points:}
	\begin{itemize}
	\item Cheaper Chinese panels AND the introduction of a leasing business model are directly linked to the surge in solar panel installations. 
	\item Leasing could be an effective way of getting over issues of asymmetric information on quality. 
	\end{itemize}
\end{frame}

\begin{frame}[plain]
\textbf{Main Points:}
	\begin{itemize}
	\item Cheaper Chinese panels AND the introduction of a leasing business model are directly linked to the surge in solar panel installations. 
	\item Leasing could be an effective way of getting over issues of asymmetric information on quality. 
	\item Leased systems tend to have lower degradation over time than those sold outright, consistent with the economic theory on the subject.
	\end{itemize}
\end{frame}

\begin{frame}[plain]
	\begin{itemize}
	\item[] \textbf{Main Lesson for Managers:} Managers need to address uncertainty about quality when selling energy investment goods to consumers and small businesses with limited financial and engineering resources.
	\end{itemize}
\end{frame}

\begin{frame}[plain]
	\begin{itemize}
	\item[] \textbf{Policy Implications:} Flexible support schemes are important for allowing business model innovations and increased investment in renewable energy.
	\end{itemize}
\end{frame}

\begin{frame}[plain]
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/Northcarolina_3rd_party_law.png}

	\end{figure}
\end{frame}

\begin{frame}[plain]
	\begin{itemize}
	\item[] \textbf{Indirect Policy Implication:} Simultaneously subsidizing an investment and adding import duties appear to be contradictory.  
	\end{itemize}
\end{frame}

\end{document}

### 2016-06-28 solar_letters.tex
add files

%solar_letters.tex

% solar_prod.tex
\documentclass[12pt]{article}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{natbib} %for citet and citep
\usepackage{syntonly}
\usepackage{esdiff} %for writing partial derivatives
\usepackage{url} %for inserting urls
\usepackage{placeins}
\usepackage{textcomp} 
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}

\doublespacing % from package setspacs

 \title{Lease or Buy? Quality differences and asymmetric information in California solar panels}

 \date{\today}

 \author{Johannes Mauritzen\\ BI Norwegian Business School\\ Trondheim Campus \\Department of Accounting, Auditing and Business Analytics \\johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}

 \begin{document}
  \begin{spacing}{1} %sets spacing to single for title page
 	\maketitle

 \begin{abstract}
  I provide one of few empirical case studies of asymmetric information of quality in a durable good. This study is unique in observing quality directly. I use data from approximately 1000 rooftop solar photovoltaic systems in California with 4 to 5 years of monthly production data. Using a Bayesian hierarchical model, I find evidence that leased systems show a higher level of quality than those bought out-right.

%  JEL Codes: L15, Q42
 \end{abstract}

 % \thanks{*I would like to thank...}
 % 
  \end{spacing}

\section{Introduction}

The theoretical literature on information asymmetry of quality is well established \citep{akerlof_market_1970, chan_prices_1982,tirole_theory_1988}. The theory has in turn spawned a broad empirical literature, especially in the finance \citep{michaely_pricing_1994, petersen_benefits_1994,adams_liquidity_2009,dobbie_information_2013}, accounting \citep{healy_information_2001} and marketing literature \citep{kirmani_no_2000}.

Empirical studies of information asymmetry in the market for durable goods are scarcer. A durable good can be seen as a one-time purchase that provides a stream of services. Thus, even if transaction data is available on the purchase of the durable good, quality is judged by the subsequent stream of services. The existing empirical literature on information asymmetry in durable goods often makes use of proxies for quality. In a study of information asymmetry in the used aircraft market, \citet{gilligan_lemons_2004} uses aircraft brands as a proxy. \citet{peterson_adverse_2014} and \citet{emons_market_2009} also use brand reputation as a proxy in studies of adverse selection in used cars. This approach does not capture the substantial individual variation in quality both within and between brands and categories.

I present a case-study using data on roof-top solar photovoltaic systems where quality can be inferred directly through monthly production data. A testable implication of the theory of asymmetric information of quality is that in the presence of asymmetric information, poor quality goods will tend to push out high quality \citep{akerlof_market_1970, tirole_theory_1988}. I compare production over time between panels that are leased and sold-outright, which both theoretical and empirical studies suggest may differ with respect to the presence of information asymmetry \citep{johnson_leasing_2003,johnson_leasing_2010, johnson_role_2014,hendel_adverse_1997,gilligan_lemons_2004}.

Photovoltaic systems that were sold outright are plausibly subject to substantial information asymmetry of quality. Homeowners and small business owners can not be expected to have the expertise to judge the quality of panels. Concerns about quality were likely extenuated by the introduction of panels from Chinese manufacturers that were nearly completely new to the US market in the time-frame studied and did not have established reputations. Anecdotally, panels from these new manufacturers varied greatly in quality.\footnote{\url{http://www.nytimes.com/2013/05/29/business/energy-environment/solar-powers-dark-side.html?pagewanted=all&_r=0
}}

The market for rooftop solar panels can be expected to be vulnerable to adverse selection due to asymmetric information on quality. Solar panels are experience goods, where a consumer learns about quality through use. In particular, poor quality panels will tend to show a higher degradation of output over time. Even then, consumers may find it difficult to measure the degradation as it can happen gradually, over many years.

More so, solar panel systems are expected to last at least 20 years, thus their purchase can be considered a one-shot investment. This eliminates repeat buying as a mechanism for ensuring quality. Warranties may also be a relatively weak assurance of quality in the market for solar panels as both contractors and manufacturers are relatively new, tend to be heavily indebted and have recently shown a tendency to go bankrupt.

However, over the period from 2010 to 2014 many of the larger solar contractors moved to leasing solar power systems to homeowners. Shifting ownership to the large contractors that install and finance the solar systems could alleviate the information asymmetry, as these contractors can take steps such as testing panels and visiting manufacturing sites to ensure quality.

In this article, I compare production levels over time between systems that are leased and sold outright, while controlling for other observable factors that could affect production.

Testing for the degradation in production of nearly 1000 panels is problematic using traditional econometric models. The data has a grouped structure with many monthly observations for each solar panel. Traditional econometric models using maximum likelihood estimation can only approximately account for this structure, and inference from models with many fixed effects may be unreliable\citep{gelman_bayesian_2013}. I use a Bayesian hierarchical model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques. The results of the Bayesian model provide evidence that solar panel systems that were leased tend to show less degradation over time than those that were sold outright.

\section{California Solar Initiative, Data and Empirical Model}

The data used in this study is from the California Solar Initiative (CSI). CSI gives incentives for installing solar panels systems in California. For most smaller systems the incentive was given in the form of an upfront payout based on the expected production of the system. Larger systems were required to accept a performance based incentive based on actual production over 5 years (60 months). From the beginning of the program in 2007 this was defined as those over 100kW, but was lowered to 50kW from January 2008 and 30kW from January 2010. Solar panel installations of all sizes have the option of getting an incentive based on actual performance. The data is openly available on the website of CSI.\footnote{\url{http://www.californiasolarstatistics.ca.gov/current_data_files/}}

Below are the key variables present in the CSI data:

\begin{itemize}
\item Installation date
\item Location: address, zip code, county
\item System capacity
\item Name of system owner, host, and contractor
\item Third party owner (leased)
\item Use of single or double axis trackers
\item Year of installation
\end{itemize}

I include only those installations that have been producing for at least 4 years (48 monthly observations), where the maximum number of observations is 60 months of production data. The data covers similar vintages of solar panels -- those that were installed from 2007 through 2009, and which continued to produce through 2013. I removed systems where production was reported to be higher than what would be theoretically possible from a solar power system with a given nameplate capacity. I also removed systems that reported 0 production in a period. This could also be an indication of quality if zero production indicates a malfunction in the system. However, it is not possible to identify which zero observations reflect malfunctions versus reporting issues.

The data has a clear grouped structure. Individual production data are grouped by the different solar panel systems. Each system is then in turn grouped into categories of leased or host-owned. Seasonality must also be adequately accounted for as well as other factors such as geographic location, use of tracking and idiosyncratic variation between systems.

I fit a Bayesian multilevel regression model using Markov-Chain Monte Carlo (MCMC). Bayesian models are especially useful in estimating multi-level models with many parameters. In a multi-level model, lower-level parameters within a certain group are themselves modeled as coming from a distribution characterized by ``meta-parameters.'' The estimated group-level parameters are a weighted average of both the observations within the group as well as the full set of observations in the dataset -- the feature of multilevel models called partial pooling. Partial pooling also serves as a natural form of parameter shrinkage, eliminating the need for corrections for multiple comparisons in inference \citep{gelman_data_2006}.

For estimation we use the Stan Bayesian programming language \citep{stan_development_team_stan_2014}, which uses Hamiltonian MCMC and a No U-Turn Sampler (NUTS) to estimate a joint posterior distribution of the probability model. \citet{kruschke_doing_2014} provides an accessible explanation of the Hamiltonian MCMC algorithm, while more detailed technical descriptions are provided by \citet{gelman_bayesian_2013} and the Stan reference manual \citep{stan_development_team_stan_2014}.

Figure \ref{solar_prod_bayes_diag} shows the structure of the model. Starting from the bottom, the individual production data are log transformed and modeled as having a Cauchy distribution with mean $\hat{y}$ and variance $\sigma$. $\sigma$ is in turn given a half-Cauchy prior.

The use of the Cauchy and half-Cauchy distribution reflects that a large-in-magnitude effect of around 5 or greater on the logit scale is highly unlikely in logistic regressions where all non-binary data has been transformed to have mean zero and standard deviation 1. The Cauchy distribution will also give answers even under complete separation, and avoids computational problems inherent in assigning completely non-informative priors in multilevel models \citep{gelman_weakly_2008}.

Going up a level, $\hat{y}$ is modeled as having random group-level intercepts $\beta_{0,j}$, where j represents the j-th of $J$ solar panel system. The $\beta_{0,j}$'s are in turn modeled as function of system size, county, whether solar trackers are used, and a random effects term $re_j^0$.

In a similar manner the J $B_j$ coefficients are modeled as a function of whether the system is leased, $\mu_{lease}^1$, the host sector of the installation, the county, use of tracking, installation year and a random effects term $re^1_j$.  The question of interest -- whether leased solar systems display a lower degree of degradation over time than those sold outright -- can then be expressed as the distribution of the difference of the $\mu_{lease}$ parameters: $\mu_{lease=1} - \mu_{lease=0}$.

Finally, to take into account the seasonality of the monthly data, random effects for each of the 12 months are estimated with a $Cauchy(0,4)$ prior on each parameter. The remaining hyper-parameters are given Cauchy or half-Cauchy priors.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/solar_prod_bayes_diag.png}
	\caption{The data and the research question combined suggest a hierarchical Bayesian model. The figure shows the structure of the model. Production data is grouped by individual solar panel systems.}
	\label{solar_prod_bayes_diag}
\end{figure}

\section{Results}

Summary statistics for the posterior distributions of the main parameters of the model are shown in table \ref{table_results}. The parameters of interest are $\mu_{lease=1}$ and $\mu_{lease=0}$, which are the average slope of production over time in, respectively, systems that were sold outright and systems that were leased. These coefficients should be interpreted for a marginal change of 1 standard deviation change in time.

\begin{table}
\footnotesize
\resizebox{12cm}{!} {
\begin{tabular}{lrrrr}
\toprule
{} &  2.5\%&   97.5\% &   mean &  median \\
params           &          &          &         &          \\
\midrule
apr              &   0.1452 &  -0.0838 &  0.0069 &  -0.0013 \\
aug              &   0.2065 &  -0.0223 &  0.0681 &   0.0594 \\
dec              &  -0.1005 &  -0.3295 & -0.2386 &  -0.2471 \\
feb              &  -0.1037 &  -0.3346 & -0.2437 &  -0.2523 \\
jan              &  -0.1935 &  -0.4234 & -0.3319 &  -0.3404 \\
jul              &   0.2364 &   0.0058 &  0.0968 &   0.0886 \\
jun              &   0.2584 &   0.0282 &  0.1191 &   0.1105 \\
mar              &  -0.0257 &  -0.2535 & -0.1631 &  -0.1717 \\
may              &   0.2067 &  -0.0225 &  0.0687 &   0.0602 \\
nov              &   0.0546 &  -0.1748 & -0.0839 &  -0.0923 \\
oct              &   0.1256 &  -0.1034 & -0.0124 &  -0.0211 \\
sep              &   0.2094 &  -0.0205 &  0.0704 &   0.0619 \\
mu1\_commercial   &   0.0014 &  -0.0041 & -0.0007 &  -0.0005 \\
mu1\_government   &   0.0035 &  -0.0026 &  0.0004 &   0.0002 \\
mu1\_non\_profit   &   0.0018 &  -0.0057 & -0.0008 &  -0.0003 \\
mu1\_not\_tracking &   0.0044 &  -0.0074 & -0.0016 &  -0.0009 \\
mu1\_residential  &   0.0039 &  -0.0023 &  0.0005 &   0.0003 \\
mu1\_tracking     &   0.0039 &  -0.0092 & -0.0021 &  -0.0014 \\
mu\_lease         &  -0.0001 &  -0.0115 & -0.0045 &  -0.0042 \\
mu\_own           &  -0.0001 &  -0.0125 & -0.0054 &  -0.0054 \\
beta0\_size       &   0.7501 &   0.6948 &  0.7173 &   0.7141 \\
beta1\_ins\_year   &  -0.0007 &  -0.0032 & -0.0020 &  -0.0020 \\
beta1\_size       &  51.1251 & -45.7583 &  0.1312 &   0.0594 \\
sigma            &   0.0295 &   0.0287 &  0.0291 &   0.0291 \\
\bottomrule
\end{tabular}
}
\label{table_results}
\caption{Summary statistics of the distributions over the main parameters in the model. Confidence intervals are estimated as quantiles at 2.5 and 97.5 percent of the estimated posterior distribution of the parameters.}
\end{table}

In figure \ref{diff_lease}, I show a histogram of the posterior of the difference between the two parameters, $\mu_{lease=1} - \mu_{lease=0}$. A direct interpretation is then that there is an approximately 80\% probability that on average solar power systems that are sold outright have a higher rate of degradation than those that are leased.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/diff_lease.png}
	\caption{The figure shows the distribution of the difference of $\mu_{lease=1}$ and $\mu_{lease=0}$. The black vertical line is placed at zero. The distribution of the difference of the two parameters suggests that leased panels had about a 80\% chance of being higher quality than those sold outright, as measured by the slope of production over time.}
	\label{diff_lease}
\end{figure}

The magnitude of the effect can best be seen by plotting average predicted values over time, which are shown for systems that are leased and sold-outright in figure \ref{predicted_deg}. The solid lines represent the median value of the posterior distribution on the $\mu_{lease=1}$ and $\mu_{lease=0}$ parameters, where the light lines represent random draws from the respective posterior distributions. The model results suggest that a leased solar panel system will on average experience degradation of about 6\% after 20 years while those owned outright experience on average 5\% of degradation. However, as the figure makes clear, the average values hide significant variation in the overall distribution.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/predicted_deg.png}
	\caption{The figure shows the average predicted values of degradation over time of solar panel systems that are leased and sold out-right. The solid lines represent the mean value of the posterior where the lighter lines represent random draws from the posterior distribution.}
	\label{predicted_deg}
\end{figure}

\section{Conclusion}
The theoretical literature on information asymmetry and quality is established and deep. The empirical literature for consumer durable investments is, however, sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability. This article provides one of few case studies that both identifies a market that could be expected to have issues of information asymmetry of quality, and provides a direct empirical test of the implications from the theoretical literature.

The use of Bayesian MCMC simulation techniques in this article also provides an illustrative example of the use and usefulness of a powerful emerging tool-set. The Bayesian frameworks and recently developed software such as Stan provide major practical and theoretical advantages over many traditional econometric methods.

\begin{spacing}{1}
\bibliographystyle{plainnat}
\bibliography{solar_prod}

\FloatBarrier

\end{spacing}
\end{document}

### 2016-06-28 solar_letters_cover.tex
add files

--- 2016-06-28 solar_letters.tex
@@
-%solar_letters.tex
-
-% solar_prod.tex
 \documentclass[12pt]{article}
@@
 
-\doublespacing % from package setspacs
+%\doublespacing % from package setspacs
 
@@
 
- \author{Johannes Mauritzen\\ BI Norwegian Business School\\ Trondheim Campus \\Department of Accounting, Auditing and Business Analytics \\johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}
+\author{Johannes Mauritzen \\ BI Norwegian Business School \\ Trondheim Campus \\ Department of Accounting, Auditing and Business Analytics \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}
 
@@
 
-%  JEL Codes: L15, Q42
+ JEL Codes: L15, Q42
 \end{abstract}
@@
  \end{spacing}
-
-\section{Introduction}
-
-The theoretical literature on information asymmetry of quality is well established \citep{akerlof_market_1970, chan_prices_1982,tirole_theory_1988}.
-The theory has in turn spawned a broad empirical literature, especially in the finance \citep{michaely_pricing_1994, petersen_benefits_1994,adams_liquidity_2009,dobbie_information_2013}, accounting \citep{healy_information_2001} and marketing literature \citep{kirmani_no_2000}.
-
-Empirical studies of information asymmetry in the market for durable goods are scarcer.
-A durable good can be seen as a one-time purchase that provides a stream of services.
-Thus, even if transaction data is available on the purchase of the durable good, quality is judged by the subsequent stream of services.
-The existing empirical literature on information asymmetry in durable goods often makes use of proxies for quality.
-In a study of information asymmetry in the used aircraft market, \citet{gilligan_lemons_2004} uses aircraft brands as a proxy.
-\citet{peterson_adverse_2014} and \citet{emons_market_2009} also use brand reputation as a proxy in studies of adverse selection in used cars.
-This approach does not capture the substantial individual variation in quality both within and between brands and categories.
-
-I present a case-study using data on roof-top solar photovoltaic systems where quality can be inferred directly through monthly production data.
-A testable implication of the theory of asymmetric information of quality is that in the presence of asymmetric information, poor quality goods will tend to push out high quality \citep{akerlof_market_1970, tirole_theory_1988}.
-I compare production over time between panels that are leased and sold-outright, which both theoretical and empirical studies suggest may differ with respect to the presence of information asymmetry \citep{johnson_leasing_2003,johnson_leasing_2010, johnson_role_2014,hendel_adverse_1997,gilligan_lemons_2004}.
-
-Photovoltaic systems that were sold outright are plausibly subject to substantial information asymmetry of quality.
-Homeowners and small business owners can not be expected to have the expertise to judge the quality of panels.
-Concerns about quality were likely extenuated by the introduction of panels from Chinese manufacturers that were nearly completely new to the US market in the time-frame studied and did not have established reputations.
-Anecdotally, panels from these new manufacturers varied greatly in quality.\footnote{\url{http://www.nytimes.com/2013/05/29/business/energy-environment/solar-powers-dark-side.html?pagewanted=all&_r=0
-}}
-
-The market for rooftop solar panels can be expected to be vulnerable to adverse selection due to asymmetric information on quality.
-Solar panels are experience goods, where a consumer learns about quality through use.
-In particular, poor quality panels will tend to show a higher degradation of output over time.
-Even then, consumers may find it difficult to measure the degradation as it can happen gradually, over many years.
-
-More so, solar panel systems are expected to last at least 20 years, thus their purchase can be considered a one-shot investment.
-This eliminates repeat buying as a mechanism for ensuring quality.
-Warranties may also be a relatively weak assurance of quality in the market for solar panels as both contractors and manufacturers are relatively new, tend to be heavily indebted and have recently shown a tendency to go bankrupt.
-
-However, over the period from 2010 to 2014 many of the larger solar contractors moved to leasing solar power systems to homeowners.
-Shifting ownership to the large contractors that install and finance the solar systems could alleviate the information asymmetry, as these contractors can take steps such as testing panels and visiting manufacturing sites to ensure quality.
-
-In this article, I compare production levels over time between systems that are leased and sold outright, while controlling for other observable factors that could affect production.
-
-Testing for the degradation in production of nearly 1000 panels is problematic using traditional econometric models.
-The data has a grouped structure with many monthly observations for each solar panel.
-Traditional econometric models using maximum likelihood estimation can only approximately account for this structure, and inference from models with many fixed effects may be unreliable\citep{gelman_bayesian_2013}.
-I use a Bayesian hierarchical model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
-The results of the Bayesian model provide evidence that solar panel systems that were leased tend to show less degradation over time than those that were sold outright.
-
-\section{California Solar Initiative, Data and Empirical Model}
-
-The data used in this study is from the California Solar Initiative (CSI).
-CSI gives incentives for installing solar panels systems in California.
-For most smaller systems the incentive was given in the form of an upfront payout based on the expected production of the system.
-Larger systems were required to accept a performance based incentive based on actual production over 5 years (60 months).
-From the beginning of the program in 2007 this was defined as those over 100kW, but was lowered to 50kW from January 2008 and 30kW from January 2010.
-Solar panel installations of all sizes have the option of getting an incentive based on actual performance.
-The data is openly available on the website of CSI.\footnote{\url{http://www.californiasolarstatistics.ca.gov/current_data_files/}}
-
-Below are the key variables present in the CSI data:
-
-\begin{itemize}
-\item Installation date
-\item Location: address, zip code, county
-\item System capacity
-\item Name of system owner, host, and contractor
-\item Third party owner (leased)
-\item Use of single or double axis trackers
-\item Year of installation
-\end{itemize}
-
-I include only those installations that have been producing for at least 4 years (48 monthly observations), where the maximum number of observations is 60 months of production data.
-The data covers similar vintages of solar panels -- those that were installed from 2007 through 2009, and which continued to produce through 2013.
-I removed systems where production was reported to be higher than what would be theoretically possible from a solar power system with a given nameplate capacity.
-I also removed systems that reported 0 production in a period.
-This could also be an indication of quality if zero production indicates a malfunction in the system.
-However, it is not possible to identify which zero observations reflect malfunctions versus reporting issues.
-
-The data has a clear grouped structure.
-Individual production data are grouped by the different solar panel systems.
-Each system is then in turn grouped into categories of leased or host-owned.
-Seasonality must also be adequately accounted for as well as other factors such as geographic location, use of tracking and idiosyncratic variation between systems.
-
-I fit a Bayesian multilevel regression model using Markov-Chain Monte Carlo (MCMC).
-Bayesian models are especially useful in estimating multi-level models with many parameters.
-In a multi-level model, lower-level parameters within a certain group are themselves modeled as coming from a distribution characterized by ``meta-parameters.'' The estimated group-level parameters are a weighted average of both the observations within the group as well as the full set of observations in the dataset -- the feature of multilevel models called partial pooling.
-Partial pooling also serves as a natural form of parameter shrinkage, eliminating the need for corrections for multiple comparisons in inference \citep{gelman_data_2006}.
-
-For estimation we use the Stan Bayesian programming language \citep{stan_development_team_stan_2014}, which uses Hamiltonian MCMC and a No U-Turn Sampler (NUTS) to estimate a joint posterior distribution of the probability model.
-\citet{kruschke_doing_2014} provides an accessible explanation of the Hamiltonian MCMC algorithm, while more detailed technical descriptions are provided by \citet{gelman_bayesian_2013} and the Stan reference manual \citep{stan_development_team_stan_2014}.
-
-Figure \ref{solar_prod_bayes_diag} shows the structure of the model.
-Starting from the bottom, the individual production data are log transformed and modeled as having a Cauchy distribution with mean $\hat{y}$ and variance $\sigma$.
-$\sigma$ is in turn given a half-Cauchy prior.
-
-The use of the Cauchy and half-Cauchy distribution reflects that a large-in-magnitude effect of around 5 or greater on the logit scale is highly unlikely in logistic regressions where all non-binary data has been transformed to have mean zero and standard deviation 1.
-The Cauchy distribution will also give answers even under complete separation, and avoids computational problems inherent in assigning completely non-informative priors in multilevel models \citep{gelman_weakly_2008}.
-
-Going up a level, $\hat{y}$ is modeled as having random group-level intercepts $\beta_{0,j}$, where j represents the j-th of $J$ solar panel system.
-The $\beta_{0,j}$'s are in turn modeled as function of system size, county, whether solar trackers are used, and a random effects term $re_j^0$.
-
-In a similar manner the J $B_j$ coefficients are modeled as a function of whether the system is leased, $\mu_{lease}^1$, the host sector of the installation, the county, use of tracking, installation year and a random effects term $re^1_j$.
-The question of interest -- whether leased solar systems display a lower degree of degradation over time than those sold outright -- can then be expressed as the distribution of the difference of the $\mu_{lease}$ parameters: $\mu_{lease=1} - \mu_{lease=0}$.
-
-Finally, to take into account the seasonality of the monthly data, random effects for each of the 12 months are estimated with a $Cauchy(0,4)$ prior on each parameter.
-The remaining hyper-parameters are given Cauchy or half-Cauchy priors.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{figures/solar_prod_bayes_diag.png}
-	\caption{The data and the research question combined suggest a hierarchical Bayesian model. The figure shows the structure of the model. Production data is grouped by individual solar panel systems.}
-	\label{solar_prod_bayes_diag}
-\end{figure}
-
-\section{Results}
-
-Summary statistics for the posterior distributions of the main parameters of the model are shown in table \ref{table_results}.
-The parameters of interest are $\mu_{lease=1}$ and $\mu_{lease=0}$, which are the average slope of production over time in, respectively, systems that were sold outright and systems that were leased.
-These coefficients should be interpreted for a marginal change of 1 standard deviation change in time.
-
-\begin{table}
-\footnotesize
-\resizebox{12cm}{!} {
-\begin{tabular}{lrrrr}
-\toprule
-{} &  2.5\%&   97.5\% &   mean &  median \\
-params           &          &          &         &          \\
-\midrule
-apr              &   0.1452 &  -0.0838 &  0.0069 &  -0.0013 \\
-aug              &   0.2065 &  -0.0223 &  0.0681 &   0.0594 \\
-dec              &  -0.1005 &  -0.3295 & -0.2386 &  -0.2471 \\
-feb              &  -0.1037 &  -0.3346 & -0.2437 &  -0.2523 \\
-jan              &  -0.1935 &  -0.4234 & -0.3319 &  -0.3404 \\
-jul              &   0.2364 &   0.0058 &  0.0968 &   0.0886 \\
-jun              &   0.2584 &   0.0282 &  0.1191 &   0.1105 \\
-mar              &  -0.0257 &  -0.2535 & -0.1631 &  -0.1717 \\
-may              &   0.2067 &  -0.0225 &  0.0687 &   0.0602 \\
-nov              &   0.0546 &  -0.1748 & -0.0839 &  -0.0923 \\
-oct              &   0.1256 &  -0.1034 & -0.0124 &  -0.0211 \\
-sep              &   0.2094 &  -0.0205 &  0.0704 &   0.0619 \\
-mu1\_commercial   &   0.0014 &  -0.0041 & -0.0007 &  -0.0005 \\
-mu1\_government   &   0.0035 &  -0.0026 &  0.0004 &   0.0002 \\
-mu1\_non\_profit   &   0.0018 &  -0.0057 & -0.0008 &  -0.0003 \\
-mu1\_not\_tracking &   0.0044 &  -0.0074 & -0.0016 &  -0.0009 \\
-mu1\_residential  &   0.0039 &  -0.0023 &  0.0005 &   0.0003 \\
-mu1\_tracking     &   0.0039 &  -0.0092 & -0.0021 &  -0.0014 \\
-mu\_lease         &  -0.0001 &  -0.0115 & -0.0045 &  -0.0042 \\
-mu\_own           &  -0.0001 &  -0.0125 & -0.0054 &  -0.0054 \\
-beta0\_size       &   0.7501 &   0.6948 &  0.7173 &   0.7141 \\
-beta1\_ins\_year   &  -0.0007 &  -0.0032 & -0.0020 &  -0.0020 \\
-beta1\_size       &  51.1251 & -45.7583 &  0.1312 &   0.0594 \\
-sigma            &   0.0295 &   0.0287 &  0.0291 &   0.0291 \\
-\bottomrule
-\end{tabular}
-}
-\label{table_results}
-\caption{Summary statistics of the distributions over the main parameters in the model.
-Confidence intervals are estimated as quantiles at 2.5 and 97.5 percent of the estimated posterior distribution of the parameters.}
-\end{table}
-
-In figure \ref{diff_lease}, I show a histogram of the posterior of the difference between the two parameters, $\mu_{lease=1} - \mu_{lease=0}$.
-A direct interpretation is then that there is an approximately 80\% probability that on average solar power systems that are sold outright have a higher rate of degradation than those that are leased.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{figures/diff_lease.png}
-	\caption{The figure shows the distribution of the difference of $\mu_{lease=1}$ and $\mu_{lease=0}$.
-The black vertical line is placed at zero.
-The distribution of the difference of the two parameters suggests that leased panels had about a 80\% chance of being higher quality than those sold outright, as measured by the slope of production over time.}
-	\label{diff_lease}
-\end{figure}
-
-The magnitude of the effect can best be seen by plotting average predicted values over time, which are shown for systems that are leased and sold-outright in figure \ref{predicted_deg}.
-The solid lines represent the median value of the posterior distribution on the $\mu_{lease=1}$ and $\mu_{lease=0}$ parameters, where the light lines represent random draws from the respective posterior distributions.
-The model results suggest that a leased solar panel system will on average experience degradation of about 6\% after 20 years while those owned outright experience on average 5\% of degradation.
-However, as the figure makes clear, the average values hide significant variation in the overall distribution.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{figures/predicted_deg.png}
-	\caption{The figure shows the average predicted values of degradation over time of solar panel systems that are leased and sold out-right.
-The solid lines represent the mean value of the posterior where the lighter lines represent random draws from the posterior distribution.}
-	\label{predicted_deg}
-\end{figure}
-
-\section{Conclusion}
-The theoretical literature on information asymmetry and quality is established and deep.
-The empirical literature for consumer durable investments is, however, sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability.
-This article provides one of few case studies that both identifies a market that could be expected to have issues of information asymmetry of quality, and provides a direct empirical test of the implications from the theoretical literature.
-
-The use of Bayesian MCMC simulation techniques in this article also provides an illustrative example of the use and usefulness of a powerful emerging tool-set.
-The Bayesian frameworks and recently developed software such as Stan provide major practical and theoretical advantages over many traditional econometric methods.
-
-\begin{spacing}{1}
-\bibliographystyle{plainnat}
-\bibliography{solar_prod}
-
-\FloatBarrier
-
-\end{spacing}
  \end{document}

### 2016-06-28 solar_prod.tex
add files

% solar_prod.tex

\documentclass[12pt]{article}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{natbib} %for citet and citep
\usepackage{syntonly}
\usepackage{esdiff} %for writing partial derivatives
\usepackage{url} %for inserting urls
\usepackage{placeins}
\usepackage{textcomp} 
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}

\doublespacing % from package setspacs

\title{Lease or Buy? Quality differences and asymmetric information in California solar panels}

\date{\today}

\author{Johannes Mauritzen \\ BI Norwegian Business School \\ Trondheim Campus \\ Department of Economics \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}

\begin{document}
 \begin{spacing}{1} %sets spacing to single for title page
	\maketitle

\begin{abstract}
 The market for rooftop solar panel systems are vulnerable to information asymmetry issues of quality. In such a market, poor quality solar panel systems could be expected to push out high quality ones. However, solar systems that are leased can be expected to overcome this asymmetric information problem. Using production data on approximately 1000 solar panel systems in California and a Bayesian hierarchical regression model, I find some evidence that leased systems show a higher level of quality - in the form of degradation over time - than those bought out-right.
\end{abstract}

% \thanks{*I would like to thank...}
% JEL Codes: Q4, L71
 \end{spacing}

\section{Introduction}
In an earlier paper \citep{mauritzen_whats_2015} I show that the boom in rooftop solar panel investment in California coincided with an adoption of a leasing business model by contractors. With leased panels, homeowners and small businesses do not pay upfront or own the panels on their roof. Instead, the contractor or a third-party own the panels and either collect a fixed monthly payment from the owner or charge a fixed price per kilowatt-hour (kWh) of electricity produced.\footnote{Charging for energy produced is often referred to as a power purchase agreement (PPA), but I here refer to all arrangements where the home- or business-owner does not own the solar panel system as a lease.}

Leased panels could be attractive for homeowners and businesses for several reasons. Not having to invest anything upfront would be attractive to those that are cash constrained. Lack of ownership could also be attractive to those concerned about the practicalities of maintaining and operating a small power station. 
In addition, I argue that an issue of information asymmetry of quality could exist in the purchase of rooftop solar panel systems, and that leasing could provide a mechanism for alleviating this asymmetry. Homeowners and small business owners can not generally be expected to have the expertise to judge the quality of panels on their own. This issue could have been extenuated by the introduction of Chinese panels from manufacturers that were nearly completely new to the US and European markets in the time-frame studied and had little to build a reputation on.

Generally, the market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality. Solar panels can be characterized as an ``experience'' good, where an investor needs to learn about the quality through use. In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels. Even then, solar panel owners may find it difficult to measure the degradation as it can happen gradually, over many years.

More so, solar panel systems are expected to last at least 20 years, thus for all practical purposes, their purchase can be considered a one-shot investment. This eliminates repeat buying as a mechanism for ensuring quality. In the literature, warranties are often suggested as a strong signal of quality. However, warranties may be a relatively weak assurance of quality in the market for solar panels as both contractors and manufacturers are relatively new, tend to be heavily indebted and some have recently shown a tendency to go bankrupt.

Given the difficulty of judging quality and lack of market mechanisms to signal quality, the established economic theory on the subject would suggest that the market may tend to provide low-quality panels \citet{tirole_theory_1988}.

However, over the period from 2010 to 2014, many of the larger solar contractors moved to leasing solar power systems to homeowners. The introduction of a leasing model could potentially get over this information asymmetry problem by shifting ownership to the large contractors that install and finance the solar systems. These contractors can then in turn take steps such as testing panels and visiting manufacturing sites to ensure quality of suppliers.

A testable implication that emerges is then that the quality of panels - as measured by degradation of output over time - is better in systems that are leased.  In this paper, I use a data set of California solar power systems that includes monthly production data. I measure the average degradation of production over time as a proxy for the inherent quality of solar panel system.

Testing for the degradation in production of nearly 1000 panels is however problematic using traditional econometric models based on frequentist approaches. The production patterns of the solar panels could be expected to vary greatly depending on location, equipment, size as well as other factors. The inclusion of many fixed effects to control for these factors would likely lead to over-fitting. That is, the model may have a good fit to the data, but will tend to have poor out-of-sample predictive power \citep{gelman_bayesian_2013}.

As a solution, I use a Bayesian hierarchical model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques. The main benefit of using a Bayesian multilevel model is that we can fit a complex model with many hundreds of parameters, while avoiding problems of over-fitting and issues of multiple comparison in inference by relying on partial pooling of the parameters. The main finding is that solar panel systems that were leased tend to show significantly less degradation over time than those that were sold outright.

The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organisation literature. Ch. 2 of \citet{tirole_theory_1988} and accompanying citation list provides a good overview of the theoretical foundations of this topic.

Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller - in other words the seller can choose the level of quality of its product. Under these assumptions, the authors predict several equilibrium - essentially markets for both high and low quality products depending on the extent to which the buyer is informed.

Reputation - or lack thereof - plays a particularly important role in this market. Most of both the solar installers and panel manufacturers are relatively new firms - especially the increasingly dominant Chinese solar panel producers,  \citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.

Expanding empirical literatures also exist on asymmetric information of quality - especially in the marketing, finance and accounting fields. In accounting, the issue of firm quality and voluntary disclosure in regulatory filings as a signal has received broad attention by empirical researchers. \citet{healy_information_2001} review the early literature. Empirical studies in finance have, for example, looked at issues of adverse-selection in Initial Public Offerings \citep{michaely_pricing_1994} and small business lending \citep{petersen_benefits_1994} among many other topics.

The literature in the marketing field may be more relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables. The effects of signaling on quality - usually in the form of advertising - is particularly important in the field. \citet{kirmani_no_2000} provide an overview of this literature.

The use of leasing, discussed in this article, can be seen as a form of costly signaling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of poor quality to the contractor.

Traditionally, energy generation has been an activity undertaken by large, specialized firms with specialized knowledge and resources. The role of information asymmetry in investment and purchase decisions has likely been seen as a secondary issue, and to my knowledge, no literature exists on the role of such informational issues in the industry.

However, the growth of distributed energy technologies, like solar panels, has added new focus on behavioral and information issues related to energy investments by homeowners and small businesses. In turn, a growing literature is forming around solar power investment behavior of homeowners and small businesses.

For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption - showing evidence for a ``solar price premium'' in homes with solar panels. \citet{bollinger_peer_2012} study the the role of peer effects in solar power adoption. They find evidence that the adoption of solar panels by homeowners in a certain zip-code will increase the probability that other households in that zip-code will install solar panels.

Despite the growth of the literature on investments in distributed energy, to my knowledge this is the first paper to identify and directly test implications of information asymmetry in the expanding industry. The paper has implications for regulation and subsidy policy for the expanding distributed energy industry.

\section{California Solar Initiative, Data and Empirical Model}

The data used in this paper is an intersection of two datasets from the California Solar Initiative (CSI). The California Solar Initiative is a state-wide program that gives incentives for installing grid-connected solar panels systems. The incentives are based on performance. For smaller systems, typical of most residential and small business installations, this incentive was given in the form of an upfront payout based on the capacity and the expected performance of the system.

Larger systems were required to accept a performance based incentive based on actual production over 5 years (60 months). From the beginning of the program in 2007 this was defined as those over 100kW, but was lowered to 50kW from January 2008 and 30kW from January 2010. Solar panel installations of all sizes have the option of getting an incentive based on actual performance.

CSI provides data on all grid-connected solar panel systems installed in California since January 2007. In addition, CSI provides another data set of monthly production data from the solar panel systems that received production incentives. The data is openly available on the website of CSI. \footnote{\url{http://www.californiasolarstatistics.ca.gov/current_data_files/}}. A cleaned and merged data set that I use in the following analysis can be found on my website at \url{jmaurit.github.io#buy_or_lease}.

Below are the key variables present in the CSI installation data:

\begin{itemize}
\item Installation date
\item Location: address, zip code, county
\item System capacity
\item Name of system owner, host, and contractor
\item Panel and inverter manufacturer 
\item Third party owner (leased)
\item Use of single or double axis trackers
\item Year of installation
\end{itemize}

The installation data can be matched with the production data for those installations receiving a production subsidy. I include only those installations that have been producing for at least 4 years (48 monthly observations), where the maximum number of observations in this data is 5 years (60 months) of production data. The data then covers similar vintage of solar panels - those that were installed from 2007 through 2009 and which continued to produce through 2013.

I removed systems from the dataset that may have had reporting errors. for example I removed systems where production was reported to be higher than what would be theoretically possible from a solar power system with a given nameplate capacity. I also removed systems that reported 0 production in a period. This could also be an indication of quality if zero production indicates a malfunction in the system. However it is not possible to identify which zero observations reflect malfunctions versus reporting issues or other non-quality issues.

Figure \ref{tot_production} shows the production for the included solar panel systems over time. The seasonality of the systems over the calendar year is clear, and needs to be accounted for in the statistical analysis.

\begin{figure}
	\includegraphics[width=1\textwidth]{tot_production.png}
	\caption{Production over time from California solar panels, Host owned and leased. Production index reflects monthly production in kWh normalized by capacity multiplied by 30.5*12 to reflect a rough estimate of maximum day-light hours in a month.}
	\label{tot_production}
\end{figure}

The main methodological goal of this paper is to make an unbiased estimate of the slope of the average degradation of solar panels - comparing the slopes between groups of systems that are leased and bought outright. These estimates should also control for seasonality and idiosyncratic factors between solar power systems.  
A simple example of fitting a straight lines through two production series is shown in figure \ref{example_prod} - here as a simple single variable OLS estimate.

\begin{figure}
	\includegraphics[width=1\textwidth]{example_prod.png}
	\caption{The main methodological task in this paper is to make an unbiased estimate of the slope of the average degradation of production over time. The figure demonstrates a straight line fitted to a time series from a single solar power system.}
	\label{example_prod}
\end{figure}

The question of interest and the structure of the available data suggests a hierarchical structure to the empirical model. The individual production data are grouped by the different solar panel systems. Each system is then in turn grouped into categories of leased or host-owned. As mentioned, seasonality must also be adequately accounted for as well as other factors such as geographic location, use of tracking and idiosyncratic variation from one solar panel system to another.

For later comparison, consider results from a traditional single-level regression of production over time estimated by maximum likelihood shown in table \ref{table:max_likelihood}. Fixed effects are included for month, host sector, county\footnote{coefficients for county are not shown}, and an indicator variable for leased systems. Variables for year of installation and size of the system are also included. The variable of interest is the interaction of the lease indicator with the time variable - which is transformed from monthly to fractional year units.

The results indicate a good fit with an $R^2$ of 90 percent. The coefficient of interest indicates that leased systems show 4\% less degradation over time than those sold outright - with a p-value of .001. However, even with the relatively large number of fixed effects included in this regression, much of the idiosyncratic variation from one photovoltaic system to another is not accounted for. I will show that when the full idiosyncratic variation is better accounted for, the evidence for a substantial difference in performance over time between leased and systems sold outright is substantially weaker than the below regression results indicate.

\begin{table}
\begin{center}
\begin{tabular}{l c }
\hline
 & Maximum Likelihood Results\\
\hline
(Intercept)                                            & $8.62^{***}$  \\
                                                       & $(0.08)$      \\
I(months\_operation/12)                                & $-0.03^{***}$ \\
                                                       & $(0.00)$      \\
Leased                         & $0.19^{***}$  \\
                                                       & $(0.01)$      \\
sys\_size\_dc                                          & $0.00^{***}$  \\
                                                       & $(0.00)$      \\        & $(0.08)$      \\
February                                         & $0.14^{***}$  \\
                                                       & $(0.02)$      \\
March                                        & $0.27^{***}$  \\
                                                       & $(0.02)$      \\
April                                        & $0.59^{***}$  \\
                                                       & $(0.02)$      \\
May                                         & $0.70^{***}$  \\
                                                       & $(0.02)$      \\
June                                         & $0.79^{***}$  \\
                                                       & $(0.02)$      \\
July                                         & $0.75^{***}$  \\
                                                       & $(0.02)$      \\
August                                         & $0.70^{***}$  \\
                                                       & $(0.02)$      \\
September                                         & $0.69^{***}$  \\
                                                       & $(0.02)$      \\
October                                       & $0.54^{***}$  \\
                                                       & $(0.01)$      \\
November                                       & $0.41^{***}$  \\
                                                       & $(0.01)$      \\
December                                        & $0.15^{***}$  \\
                                                       & $(0.01)$      \\
Fixed Array                            & $-0.52^{***}$ \\
                                                       & $(0.03)$      \\
Mixed                                  & $-0.34^{***}$ \\
                                                       & $(0.06)$      \\
Single-Axis Tracker                    & $-0.31^{***}$ \\
                                                       & $(0.04)$      \\
Government                               & $0.07^{***}$  \\
                                                       & $(0.01)$      \\
Non-Profit                               & $-0.17^{***}$ \\
                                                       & $(0.02)$      \\
Residential                              & $-2.12^{***}$ \\
                                                       & $(0.01)$      \\
I(months\_operation/12):leased & $0.04^{***}$  \\
                                                       & $(0.01)$      \\
\hline
R$^2$                                                  & 0.90          \\
Adj. R$^2$                                             & 0.90          \\
Num. obs.                                              & 32979         \\
RMSE                                                   & 0.57          \\
\hline
\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$}}
\end{tabular}
\caption{Results from a standard linear regression model with fixed effects for }
\label{table:max_likelihood}
\end{center}
\end{table}

Instead of standard single-level maximum likelihood estimation,  I use a hierarchical Bayesian model where parameters can be estimated through Markov Chain Monte Carlo (MCMC) simulation techniques. The main advantage of such models are the flexibility of taking into account the multi-level structure of the data and improved inference that comes from partial pooling.

Consider first a no pooling model. The slope of production over time from each solar panel system would be estimated independently - allowing for large amounts of solar-power-system level variance, but excluding information from data on all the other plants from each of the individual system-level inferences. On the other hand, in a fully pooled model, the data from all of the observations of all the solar power systems in each group (leased, sold outright) are used in the estimation, but idiosyncratic variation between solar-power-systems are not taken into account. An outlying solar power system may, for instance, have too much influence and bias the results.

A hierarchical model with partial pooling is a compromise between these two extremes. Individual estimators of the solar-power-system slopes make use of the information in the full dataset with weights established by the relative variance between and within solar power systems. A full discussion of partial pooling can be found in \citet{gelman_bayesian_2013} or \citet{kruschke_doing_2014}.

Figure \ref{solar_prod_bayes_diag} shows the structure of the model. Starting from the bottom, the individual production data are log transformed and modeled as having a Cauchy distribution with mean $y_hat$ and variance $sigma^2$. $sigma^2$ is in turn given a half-Cauchy prior.

The use of the Cauchy and half-Cauchy distribution reflects that a large in magnitude effect of around 5 or greater on the logit scale is highly unlikely in logistic regressions where all non-binary data has been transformed to have mean zero and standard deviation 1 - which I have done. In addition, the Cauchy distribution will give answers even under complete separation, and avoids computational problems inherent in assigning completely non-informative priors in multilevel models \citep{gelman_weakly_2008}.

Going up a level, $y_hat$ is modeled as having random group-level intercepts $\beta_{0,j}$, where j represents the j-th of $J$ solar panel system. The $\beta_{0,j}s$ are in turn modeled as function of system size, county, whether solar trackers are used, and a random effects term $re_j^0$.

In a similar manner the J $B_j$ coefficients are modeled as a function of whether the system is leased, $\mu_{lease}^1$, the host sector of the installation, the county, use of tracking, installation year and a random effects term $re^1_j$.  The question of interest - whether leased solar systems display a lower degree of degradation over time than those sold outright - can then be expressed as the distribution of the difference of the $mu_{lease}$ parameters: $\mu_{lease=1} - \mu_{lease=0}$.

Finally, to take into account the seasonality of the monthly data, random effects for each of the 12 months are estimated with a $Cauchy(0,4)$ prior on each parameter. The remaining hyper-parameters are given Cauchy or half-Cauchy priors.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/solar_prod_bayes_diag.png}
	\caption{The data and the research question combined suggest a hierarchical Bayesian model. The figure shows the structure of the model. Production data is grouped by individual solar panel systems (index i).}
	\label{solar_prod_bayes_diag}
\end{figure}

To estimate the parameters of the model, I use the Stan Bayesian programming language and simulator \citep{stan_development_team_stan_2014} which uses Hamiltonian MCMC to estimate a joint posterior distribution of the parameters. \citet{kruschke_doing_2014} provides an accessible explanation of the Hamiltonian MCMC algorithm, while more detailed technical descriptions are provided by \citet{gelman_bayesian_2013} and \citet{stan_development_team_stan_2014}.

Bayesian simulation methods are still emerging - especially in Economics. So it is worth spending a moment to motivate their use in this case over more common asymptotic methods. As mentioned, the main reason for using the Bayesian model is the flexible treatment it allows for modeling the inherent hierarchy of the data and allowing for partial pooling in estimation.

The usefulness of Bayesian models is most apparent in the estimation of complex hierarchical models with many parameters.  
In a hierarchical model, lower-level parameters within a certain group are themselves modeled as coming from a distribution characterized by ``meta-parameters''. Thus, the estimated group-level parameters are a weighted function of both the observations within the group as well as the full set of observations in the dataset. For example, the estimated mean parameter on a group with a few, outlying observations would be pulled towards the full sample mean. This partial pooling also serves as a natural form of parameter shrinkage, which mostly eliminates the need for using corrections for multiple comparisons in inference \citep{gelman_data_2006}.

With Bayesian multilevel model, I also have greater flexibility in specifying the model without having to rely on assumptions like constant group-level variances and a Gaussian noise distribution.  Because Bayesian simulation techniques result in an estimate of the full joint probability distribution of the model, the inference has the potential to be more informative than the typical point estimates and p-values of of the standard hypothesis testing frameworks \citep{kruschke_doing_2014}.

Another advantage is that the posterior distributions of the parameters have a natural and direct interpretation as probabilities, as opposed to standard hypothesis testing where the concepts of p-values and significance are often ill-defined. See \citet{kruschke_doing_2014} or \citet{gelman_bayesian_2013} for further discussion.

% \begin{align}
% 		log(production_{ij}) &\sim N(\hat{y}, \sigma^2), \quad i=1,...,n \\
% 		\hat{y} &= \beta_{0j} + \beta_{1j}*time_{ij} + month_{m}
% 		\beta_{1,j} &= \mu_{\beta, l} + \zeta_{b1, j}, \quad j=1...j, \quad l = 0,1 \\
% 		\mu_{\beta1, l} &\sim N(0,10) \\
% 		\beta_{0,j} &\sim N(\mu_{b0}, \sigma_{b0}) \\
% 		\month_{m} &\sim N(1,10)\\
% 		\sigma &\sim uniform(0,100) \\
% 		\zeta_{b1, j} &\sim N(0,\sigma_{b1}) 
% 		\label{equation:model}
% 		\caption{}
% \end{align}

\section{Results}

I run 1000 iterations with 4 chains of the Hamiltonian MCMC algorithm in order to sample the joint posterior probability distribution of the model. I can then extract the the estimates of the marginal posterior distributions of the individual parameters in the form of a list of simulated parameter estimates.

Summary statistics for the posterior distributions of the main parameters of the model are shown in table \ref{table_results}. The parameters of interest are $\mu_{own}, \mu_{lease}$ which are the average slope of production over time in, respectively, systems that were sold outright and systems that were leased. These coefficients as well as the other component coefficients of $\beta1$, should be interpreted for a marginal change of 1 standard deviation change in time. Intuitive illustrations of the magnitudes of the coefficients of interest are shown further below.

\begin{table}
\resizebox{12cm}{!} {
\begin{tabular}{lrrrr}
\toprule
{} &  2.5\%&   97.5\% &   mean &  median \\
params           &          &          &         &          \\
\midrule
apr              &   0.1452 &  -0.0838 &  0.0069 &  -0.0013 \\
aug              &   0.2065 &  -0.0223 &  0.0681 &   0.0594 \\
dec              &  -0.1005 &  -0.3295 & -0.2386 &  -0.2471 \\
feb              &  -0.1037 &  -0.3346 & -0.2437 &  -0.2523 \\
jan              &  -0.1935 &  -0.4234 & -0.3319 &  -0.3404 \\
jul              &   0.2364 &   0.0058 &  0.0968 &   0.0886 \\
jun              &   0.2584 &   0.0282 &  0.1191 &   0.1105 \\
mar              &  -0.0257 &  -0.2535 & -0.1631 &  -0.1717 \\
may              &   0.2067 &  -0.0225 &  0.0687 &   0.0602 \\
nov              &   0.0546 &  -0.1748 & -0.0839 &  -0.0923 \\
oct              &   0.1256 &  -0.1034 & -0.0124 &  -0.0211 \\
sep              &   0.2094 &  -0.0205 &  0.0704 &   0.0619 \\
mu1\_commercial   &   0.0014 &  -0.0041 & -0.0007 &  -0.0005 \\
mu1\_government   &   0.0035 &  -0.0026 &  0.0004 &   0.0002 \\
mu1\_non\_profit   &   0.0018 &  -0.0057 & -0.0008 &  -0.0003 \\
mu1\_not\_tracking &   0.0044 &  -0.0074 & -0.0016 &  -0.0009 \\
mu1\_residential  &   0.0039 &  -0.0023 &  0.0005 &   0.0003 \\
mu1\_tracking     &   0.0039 &  -0.0092 & -0.0021 &  -0.0014 \\
mu\_lease         &  -0.0001 &  -0.0115 & -0.0045 &  -0.0042 \\
mu\_own           &  -0.0001 &  -0.0125 & -0.0054 &  -0.0054 \\
beta0\_size       &   0.7501 &   0.6948 &  0.7173 &   0.7141 \\
beta1\_ins\_year   &  -0.0007 &  -0.0032 & -0.0020 &  -0.0020 \\
beta1\_size       &  51.1251 & -45.7583 &  0.1312 &   0.0594 \\
sigma            &   0.0295 &   0.0287 &  0.0291 &   0.0291 \\
\bottomrule
\end{tabular}
}
\label{table_results}
\caption{Summary statistics of the distributions over the main parameters in the model. Confidence intervals are estimated as quantiles at 2.5 and 97.5 percent of the estimated posterior distribution of the parameters.}
\end{table}

I display the histograms representing the posterior distributions of these two parameters in the top panel of figure \ref{diff_lease}. The histograms indicate that the posterior distributions appear to have means that are different from each other.

In the lower panel, I show a histogram of the posterior of the difference between the two parameters, $\mu_{b1,l=1} - \mu_{b1,l=2} $ which is the direct test of the question of interest. 95\% of the probability lies below the vertical line, which stands at approximately zero. A direct interpretation is then that there is a 95 \% probability that on average solar power systems that are sold outright have a higher rate of degradation than those that are leased.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/diff_lease.png}
	\caption{The figure shows the distribution of the difference of $\mu_{lease}$ and $\mu_{own} $. The black vertical line is placed at zero. The distribution of the difference of the two parameters suggests that leased panels had about a 80\% chance of being higher quality than those sold outright, as measured by the slope of production over time.}
	\label{diff_lease}
\end{figure}

The magnitude of the effect can best be seen by plotting average predicted values over time, which are shown for systems that are leased and sold-outright in figure \ref{predicted_deg}. The solid lines represent the median value of the posterior distribution on the $mu_lease$ and $mu_old$ parameters, where the light lines represent random draws from the respective posterior distributions to represent uncertainty. The model suggests that a leased solar panel system will on average experience degradation of about 6\% after 20 years while those owned outright experience on average of 5\% of degradation. However, as the figure makes clear, the average values hide significant variation in the overall distribution.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/predicted_deg.png}
	\caption{The figure shows the average predicted values of degradation over time of solar panel systems that are leased and sold out-right. The solid lines represent the mean value of the posterior where the lighter lines represent random draws from the posterior distribution}
	\label{predicted_deg}
\end{figure}

\section{Discussion and Conclusion}
The theoretical literature on information asymmetry and quality is established and deep. The empirical literature for consumer durable investments is, however, more sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability. This article provides one of few case studies that both identifies a market that could be expected to have issues of information asymmetry of quality, and provides an indirect empirical test for the presence of information asymmetry of quality.

The structure of the emerging solar industry in the US, as well as other parts of the world suggests that issues of information asymmetry may play an important role. The established theory on the subject suggests that under information asymmetry of quality, poor quality products may push out good quality products, at least in some sub-markets. In this article I have tested an implication of that theory for the case of solar panels in California. The results provide some weak evidence for the existence of information asymmetry issues of quality.

The use of Bayesian MCMC simulation techniques in this article also provides an illustrative example of the use and usefullnes of a powerful emerging toolset. The Bayesian frameworks and recently developed software such as Stan provide major practical and theoretical advantages over many traditional econometric methods. These methods and tools are also increasingly accessible to the practicioner. \citet{kruschke_doing_2014} and \citet{stan_development_team_stan_2014} provide accessible and thorough introductions and guides.

This article should also be considered as part of a broader literature on the special characteristics of new distributed energy generation technologies. The energy investment behavior of home owners, farmers, and small cooperatives are bound to be substantially different than those of large, specialized energy companies that have traditionally done most of the investment in electricity generation. Understanding the quickly evolving energy and power industry requires taking into account informational and behavioral factors, and this is a rich field for further empirical research

\begin{spacing}{1}
\bibliographystyle{plainnat}
\bibliography{solar_prod}

\FloatBarrier

\appendix
\section{Appendix: Tables}

\end{spacing}
\end{document}

### 2016-06-28 analysis.py

@@ -1,0 +1,116 @@
+#analysis.py
+
+import pandas as pd
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import pickle
+import scipy
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+
+solar_data= pd.read_csv("research/solar_prod/solar_data.csv")
+
+#get rid of NAs
+
+solar_data = solar_data[solar_data.prod_index.notnull()]
+solar_data = solar_data[solar_data.months_operation.notnull()]
+
+N = solar_data.count(axis=0)[0]
+
+mean_prod_index = solar_data["prod_index"].mean()
+mean_months_operation = solar_data["months_operation"].mean()
+sd_prod_index = solar_data["prod_index"].std()
+sd_months_operation = solar_data["months_operation"].std()
+
+
+solar_stan_data = {"prod_index":solar_data["prod_index"],
+"months_operation":solar_data["months_operation"],
+"N":N, 
+"mean_prod_index":mean_prod_index,
+"mean_months_operation":mean_months_operation,
+"sd_prod_index":sd_prod_index,
+"sd_months_operation":sd_months_operation
+}
+
+solar_stan = """
+data {
+	int<lower = 0> N; // number of observations
+	vector[N] prod_index; //response variable
+	vector[N] months_operation; //main predictor
+	real mean_prod_index;
+	real mean_months_operation;
+	real sd_prod_index;
+	real sd_months_operation;
+}
+transformed data{ //for constants in priors
+	real unifLo;
+	real unifHi;
+	real expLambda;
+	real alphasigma;
+	real beta1sigma;
+	unifLo <- sd_prod_index/1000;
+	unifHi <- sd_prod_index*1000;
+	expLambda<-1/29.0;
+	alphasigma <- 10*fabs(mean_months_operation*sd_prod_index/sd_months_operation);
+	beta1sigma <- 10*fabs(sd_prod_index/sd_months_operation);
+}
+parameters {
+	real a; //intercept
+	real b_1; //slope
+	real<lower=0> sigma; //standard deviation
[…]

### 2016-06-28 analysis.r

@@ -1,0 +1,66 @@
+#analysis.r
+
+rm(list = ls())
+
+library(ggplot2)
+library(reshape2)
+library(dplyr)
+library(plyr) # upgrade to dplyr
+library(arm) #special functions for Gelman & Hill
+library(lme4)
+library(mgcv)
+library(texreg)
+library(lubridate) #for extracting date
+# library(rstan)
+library(sandwich) #robust standard errors
+library(grid)
+
+solar_data <- read.csv("/Users/johannesmauritzen/research/solar_prod/solar_data.csv", stringsAsFactor = FALSE)
+
+#lmer model
+mlm_production<-lmer(log(prod_index + .0001) ~ months_operation + factor(month) + factor(year) + installation_year + own_sect*months_operation + 
+	(1+months_operation|app_num), data=solar_data)
+summary(mlm_production)
+
+mlm_production2<-lmer(log(prod_index + .0001) ~ months_operation + factor(month) + factor(year) + installation_year + onw_sect + cost_per_max_kwh*months_operation + 
+	(1+months_operation|app_num), data=solar_data)
+summary(mlm_production2)
+
+
+reg_prod_coef <- fixef(mlm_production2)
+reg_prod_se <- se.fixef(mlm_production2)
+reg_prod <- data.frame(reg_prod_coef, reg_prod_se)
+
+
+#simulations from posterior:
+mlm_prod_sim <- sim(mlm_production2)
+meta_sim1 <- mlm_prod_sim@fixef
+meta_sim1 <- as.data.frame(meta_sim1)
+
+a_hat<-coef(mlm_production2)$app_num[,1]
+b_hat<-coef(mlm_production2)$app_num[,2]
+a_se<-se.ranef(mlm_production2)$app_num[,1]
+b_se<-se.ranef(mlm_production2)$app_num[,2]
+
+app_num <- rownames(coef(mlm_production)$app_num)
+
+cross_section<-read.csv("/Users/johannesmauritzen/research/solar_prod/solar_cross_section.csv", stringsAsFactor = FALSE)
+
+
+ab_s <-data.frame(a_hat = a_hat, b_hat=b_hat, 
+	a_se=a_se, b_se=b_se, app_num = app_num)
+
+reg_data <- merge(ab_s, cross_section, by="app_num")
+reg_data <- reg_data[order(reg_data$b_hat),]
+
+production_slope<-ggplot(reg_data) +
[…]

### 2016-06-28 analysis_wide.py

@@ -1,0 +1,141 @@
+#analysis_wide.py
+
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+
+#data file with many covariates, but limited sample
+prod_wide=pd.read_csv("research/solar_prod/prod_wide_new.csv")
+prod_wide["prod_end_date"]=pd.to_datetime(prod_wide["prod_end_date"])
+prod_wide["app_received"]=pd.to_datetime(prod_wide["app_received"])
+prod_wide["app_complete"]=pd.to_datetime(prod_wide["app_complete"])
+prod_wide["app_approved"]=pd.to_datetime(prod_wide["app_approved"])
+
+
+wide_cross = prod_wide[~prod_wide.app_num.duplicated()]
+wide_cross.groupby("program").size()
+wide_cross.groupby("own_sect").size()
+wide_cross.groupby("host_cust_sect").size()
+wide_cross.groupby("cust_county").size()
+wide_cross.groupby("technology").size()
+wide_cross.groupby("sector").size()
+wide_cross.groupby("pace").size()
+wide_cross.groupby("ev").size()
+wide_cross.groupby("utility").size()
+wide_cross.groupby(["third_party_owned", "host_cust_sect"]).size()
+
+#prod_wide.reset_index(inplace=True)
+#calculate theoretical maximul
+prod_wide["prod_max"]=prod_wide["sys_size_ac"]*12*30.5
+prod_wide["prod_index"] = prod_wide["prod_kwh"]/prod_wide["prod_max"]
+
+#create months of operation variable
+count_index = lambda v: [i+1 for i, x in enumerate(v)]
+prod_wide["months_operation"] = prod_wide.groupby("app_num")["pbi_payment_num"].transform(count_index)
+
+#look for outliers
+median_prod_index= prod_wide.groupby("app_num")["prod_index"].apply(lambda x: x.median())
+outlier = median_prod_index.index[median_prod_index<.1]
+outlier=outlier.get_values()
[…]

### 2016-06-28 analysis_wide.r

@@ -1,0 +1,99 @@
+#analysis_wide.r
+
+rm(list = ls())
+
+library(ggplot2)
+library(reshape2)
+library(dplyr)
+library(arm) #special functions for Gelman & Hill
+library(lme4)
+library(mgcv)
+library(texreg)
+library(lubridate) #for extracting date
+# library(rstan)
+library(sandwich) #robust standard errors
+library(grid)
+#library(plm) #for panel data
+
+prod_wide <- read.csv("/Users/johannesmauritzen/research/solar_files/prod_wide_clean.csv", stringsAsFactor = FALSE)
+
+#linear model
+
+prod_wide <- prod_wide[!is.na(prod_wide$prod_index),]
+prod_wide <- prod_wide[!is.na(prod_wide$months_operation),]
+
+
+
+prod_wide$log_prod_kwh_st <- (log(prod_wide$prod_kwh) - mean(log(prod_wide$prod_kwh)))/sd(log(prod_wide$prod_kwh))
+prod_wide$months_operation_st <- prod_wide$months_operation - mean(prod_wide$months_operation)
+
+lm_prod <- lm(prod_index ~ months_operation*lease + 
+	factor(month)*factor(year), data=prod_wide)
+summary(lm_prod)
+
+lm_prod2 <- lm(log_prod_kwh_st ~ I(months_operation_st/12)*factor(third_party_owned) + sys_size_dc +
+	factor(cust_county) + factor(month) + factor(tracking) + I(months_operation_st/12)*factor(sector), data=prod_wide)
+summary(lm_prod2)
+
+texreg(lm_prod2, robust=TRUE, digits=3)
+
+plm_prod2 <- plm(I(log(prod_kwh)) ~ I(months_operation/12)*factor(third_party_owned) + 
+	sys_size_dc + factor(cust_county) + factor(month)  + factor(tracking) + factor(sector), data=prod_wide, model = "pooling")
+summary(plm_prod2)
+
+
+
+mlm_production<-lmer(log(prod_index + .0001) ~ months_operation*lease + factor(month) + 
+	(1+months_operation|app_num), data=prod_wide)
+summary(mlm_production)
+
+reg_new_coef<-fixef(mlm_production)
+reg_new_se<-se.fixef(mlm_production)
+reg_new<-data.frame(reg_new_coef, reg_new_se)
+
+mlm_production_sim<-sim(mlm_production, 1000)
[…]

### 2016-06-28 data_clean.py

@@ -1,0 +1,93 @@
+#data_clean.py
+
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+pd.set_option('display.max_columns', None)
+pd.set_option("display.max_rows", None)
+
+#import data solar initiative
+#Measured production data
+#from http://www.californiasolarstatistics.ca.gov/current_data_files/
+
+#"Projects that receive Performance-Based Incentives (PBI) 
+# also provide monthly system production data, which can be found in 
+# the Measured Production Data Set below. ""
+prod=pd.read_csv("research/solar_prod/MeasuredProduction_2-25-2015.csv")
+
+prod.columns=["app_num", "prog_admin", "program", "tot_cost", "own_sect", 
+"host_cust_sect","cust_city", "cust_county", "cust_zip", "pbi_payment_num",
+"prod_end_date", "prod_kwh"]
+
+#convert date to date format
+prod["prod_end_date"]=pd.to_datetime(prod["prod_end_date"], format="%Y-%m-%d")
+
+#convert production from int to float
+prod["prod_kwh"]=prod["prod_kwh"].astype(float)
+
+
+#link to data on installations
+installations=pd.read_csv("/Users/johannesmauritzen/Google Drive/Research/CalSolar/multilevel_data.csv")
+installations.rename(columns={'Application.Number': 'app_num'}, inplace=True)
+
+interconnected = pd.read_csv("/Users/johannesmauritzen/research/solar_prod/interconnected_solar.csv")
+interconnected = interconnected[['Application Id', 'Utility', 
+		'Service City', 'Service Zip',
+     	 'Service County', 'Technology Type', 'System Size DC',
+     	  'System Size AC','Mounting Method', 'Tracking',
+       	 'Customer Sector', 'App Received Date', 'App Complete Date', 
+       	 'App Approved Date', 'Installer Name', 'Third Party Owned',
+       	  'Third Party Owned Type', 'Third Party Name', 
+       	  'Pace Financed', 'Pace Financier',
+       'Electric Vehicle', 'Electric Vehicle Count', 
+       'Total System Cost','Itc Cost Basis', 
+       'Matched CSI Application Number', 'Application Status',
+       'Project is VNEM, NEM-V, NEM-Agg?', 'Module Model 1',
+       'Module Manufacturer 1']]
+
[…]

### 2016-06-28 descriptives.py

@@ -1,0 +1,191 @@
+#descriptives.py
+
+import pandas as pd
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+
+
+#larger sample but fewer covariates
+prod_long=pd.read_csv("research/solar_prod/prod_long.csv")
+prod_long["prod_end_date"]=pd.to_datetime(prod_long["prod_end_date"])
+
+
+
+#data file with many covariates, but limited sample
+prod_wide=pd.read_csv("research/solar_prod/prod_wide.csv")
+prod_wide["prod_end_date"]=pd.to_datetime(prod_wide["prod_end_date"])
+prod_wide["complete_date"]=pd.to_datetime(prod_wide["complete_date"])
+
+wide_cross = prod_wide.groupby("app_num").apply(lambda x: x.iloc[0])
+wide_cross.groupby("program").size()
+
+#just the individual cells
+
+
+
+
+#create max-production variable
+prod_long.reset_index(inplace=True)
+prod_long["prod_max"]=prod_long.groupby("app_num")["prod_kwh"].transform(max)
+prod_long["prod_index"] = prod_long["prod_kwh"]/prod_long["prod_max"]
+
+#create months of operation variable
+count_index = lambda v: [i+1 for i, x in enumerate(v)]
+prod_long["months_operation"] = prod_long["index"].groupby(prod_long["app_num"]).transform(count_index)
+
+prod_long["month"] = prod_long["prod_end_date"].dt.month
+month_dummies = pd.get_dummies(prod_long["month"])
+month_dummies.columns = ["jan","feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct" ,"nov", "dec"]
+prod_long["year"]=prod_long["prod_end_date"].dt.year
+year_dummies = pd.get_dummies(prod_long["year"])
+year_dummies.columns = ["y2007", "y2008", "y2009","y2010", "y2011", "y2012", "y2013", "y2014", "y2015"]
+
+solar_data=pd.concat([prod_long, month_dummies], axis=1)
+solar_data=pd.concat([solar_data, year_dummies], axis=1)
+
[…]

### 2016-06-28 solar_prod_MLM.py

@@ -1,0 +1,400 @@
+
+# coding: utf-8
+
+# In[2]:
+
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+
+# In[3]:
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+#get_ipython().magic('matplotlib inline')
+
+
+# In[4]:
+
+def get_percentiles(ser):
+    return(np.percentile(ser, [2.5,15,50,85,97.5]))
+
+def logit_to_prob(value):
+    return(math.exp(value)/(1+math.exp(value)))
+
+
+# In[5]:
+
+solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_wide_clean.csv")
+
+solar_data = solar_data[solar_data.prod_index.notnull()]
+solar_data = solar_data[solar_data.months_operation.notnull()]
+
+app_nums  = solar_data.app_num.unique()
+installation=[i + 1 for i in range(0,len(app_nums))]
+installations = pd.DataFrame({"app_num":app_nums, "installation":installation})
+solar_data = solar_data.merge(installations, how="left", on="app_num")
+
+
+# In[6]:
+
+#age of panel system in data
+N = solar_data.count(axis=0)[0]
+
+app_num = solar_data["app_num"]
+unique_app_num = app_num.unique()
+J = unique_app_num.size
+L = 2
+
+solar_data["log_prod_kwh"] = np.log(solar_data["prod_kwh"])
+
+
+# In[7]:
+
+#observation data
+log_prod_kwh = solar_data["log_prod_kwh"] 
+month = solar_data["month"]
+
+
+# In[8]:
+
+#installation level variables
+installations = solar_data[["app_num", "installation", "cust_county", "sector", "tracking", "sys_size_dc", "third_party_owned", "installation_year"]][~solar_data.installation.duplicated()]
+
+
+# In[9]:
+
+#lease variables
[…]

### 2016-06-28 stan_analysis.py

@@ -1,0 +1,240 @@
+#stan_analysis.py
+
+#analysis_wide.py
+
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+
+solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_wide_clean.csv")
+
+solar_data = solar_data[solar_data.prod_index.notnull()]
+solar_data = solar_data[solar_data.months_operation.notnull()]
+
+app_nums  = solar_data.app_num.unique()
+installation=[i + 1 for i in range(0,len(app_nums))]
+installations = pd.DataFrame({"app_num":app_nums, "installation":installation})
+
+solar_data = solar_data.merge(installations, how="left", on="app_num")
+
+
+#age of panel system in data
+N = solar_data.count(axis=0)[0]
+
+app_num = solar_data["app_num"]
+unique_app_num = app_num.unique()
+J = unique_app_num.size
+L = 2
+
+solar_data["log_prod_kwh"] = np.log(solar_data["prod_kwh"])
+#sns.lmplot(x="months_operation", y="log_prod_kwh", data = solar_data[solar_data.app_num =="PGE-CSI-00002"])
+#lt.show()
+
+#observation level data
+log_prod_kwh = solar_data["log_prod_kwh"] 
+month = solar_data["month"]
+
+fig, ax = plt.subplots()
+plt.scatter
+
+#installation level variables
+installations = solar_data[["app_num", "installation", "cust_county", "sector", "tracking", "sys_size_dc", "third_party_owned", "installation_year"]][~solar_data.installation.duplicated()]
+
+#lease variables
+pd.get_dummies(solar_data["third_party_owned"])["yes"]
+lease = pd.get_dummies(installations.third_party_owned)
+lease = [int(i+1) for i in lease.yes]
+
[…]

### 2016-06-28 stan_helpers.py

@@ -1,0 +1,54 @@
+import pandas as pd
+import numpy as np
+
+def get_cent_interval(sample, conf=.95):
+	p=1-conf
+	sample = np.sort(sample)
+	n_sample = len(sample)
+	low_ci = sample[round(n_sample*(p/2))]
+	high_ci = sample[round(n_sample*(1-p/2))]
+	return([low_ci, high_ci])
+
+def summary_info(post_sample, lim=5):	
+	#post_sample = solar_params
+	#lim=5
+	"""
+	Input extracted STAN model sample: post_sample
+	lim - confidence interval ie lim = 5 => 95 central posterior interval  
+	"""
+
+	params = []
+	means = []
+	medians = []
+	modes = []
+	low_ci = []
+	high_ci = []
+	for i in post_sample:
+		param_sim = post_sample[i]
+		if (param_sim.shape[0] == param_sim.size):
+			params.append(i)
+			means.append(np.mean(param_sim))
+			medians.append(np.median(param_sim))
+			n_samples = len(param_sim)
+			n_tail = n_samples*(lim/100)/2
+			param_sim=np.sort(param_sim)
+			low_ci.append(param_sim[n_tail])
+			high_ci.append(param_sim[n_samples-n_tail])
+		else:	
+			for par_case in param_sim.T:
+				params.append(i)
+				means.append(np.mean(par_case))
+				medians.append(np.median(par_case))
+				modes.append(np.mode(parcase))
+				n_samples = len(par_case)
+				n_tail = n_samples*(lim/100)/2
+				par_case=np.sort(par_case)
+				low_ci.append(par_case[n_tail])
+				high_ci.append(par_case[n_samples-n_tail])
+	output = {"params":params,
+		"means" : means,
+		"medians" : medians,
+		"low_ci" : low_ci,
+		"high_ci" : high_ci}
+	output = pd.DataFrame(output)
+	return(output)

### 2017-12-10 solar_commodities.tex
Initial ML estimation done

--- 2016-06-28 solar_prod.tex
@@
 
-\title{Lease or Buy? Quality differences and asymmetric information in California solar panels}
+\title{Are Solar Panels Commodities? Evidence of Quality Differences and Assymetric Information from California}
 
@@
 
-\author{Johannes Mauritzen \\ BI Norwegian Business School \\ Trondheim Campus \\ Department of Economics \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}
+\author{Johannes Mauritzen \\ BI Norwegian Business School \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}
 
@@
 \begin{abstract}
- The market for rooftop solar panel systems are vulnerable to information asymmetry issues of quality.
-In such a market, poor quality solar panel systems could be expected to push out high quality ones.
-However, solar systems that are leased can be expected to overcome this asymmetric information problem.
-Using production data on approximately 1000 solar panel systems in California and a Bayesian hierarchical regression model, I find some evidence that leased systems show a higher level of quality - in the form of degradation over time - than those bought out-right.
+ Solar panels should not be considered commodities.
+Considerable quality differences, as measured by average production over time, are found between manufacturers.
+This has implications for pricing and competition in the market for solar panel systems.
+I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with third party owners display higher quality than those owned directly.
+2.) Prices of solar panels are not related to quality for solar panels that are owned directly.
+I use multilevel models estimated by both maximum likelihood and Bayesian Markov Chain Monte-Carlo.
 \end{abstract}
@@
 \section{Introduction}
+
+Solar panels have been referred to colloquially in the power industry as being commodities.
+This is in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade.
+This has made solar power price competitive without subsidies in many sunny locations.
+But it has also meant upheaval in the panel manufacturing industry with cheaper Chinese produced panels replacing panels produced in Europe, North America and Japan.
+Most recently, this led to the imposition of trade tariffs on Chinese produced solar panels by the US government.
+
+However, a more thorough definition of a commodity generally includes a certain degree of fungibility.
+That is, the good from a certain producer will be largely interchangeable with the good of another producer.
+Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea is sold at the same price, as established by an exchange.
+In a similar manner, the electricity produced by a solar panel at a certain location and time can also be considered a commodity.
+
+However, an important differentiating characteristic between solar panels is the quality of these solar panels.
+This can be measured as both the failure rate of these panels as well as the degree of degradation of the panels over time.
+This can imply that certain manufacturers with superior quality could extract a premium.
+
+The combination of quality differences between panel manufacturers and the distributed nature of solar panel generation also opens up the issue of asymmetric information in the market for solar panels.
+
 In an earlier paper \citep{mauritzen_whats_2015} I show that the boom in rooftop solar panel investment in California coincided with an adoption of a leasing business model by contractors.
@@
 Lack of ownership could also be attractive to those concerned about the practicalities of maintaining and operating a small power station.
+
 In addition, I argue that an issue of information asymmetry of quality could exist in the purchase of rooftop solar panel systems, and that leasing could provide a mechanism for alleviating this asymmetry.

### 2017-12-10 data_clean.py

@@ -1,5 +1,6 @@
 #data_clean.py
-
+#last updated 2.2.2018
+#
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
@@ -8,86 +9,144 @@ import seaborn as sns
 pd.set_option('display.max_columns', None)
 pd.set_option("display.max_rows", None)
 
+#%%
 #import data solar initiative
 #Measured production data
 #from http://www.californiasolarstatistics.ca.gov/current_data_files/
 
-#"Projects that receive Performance-Based Incentives (PBI) 
-# also provide monthly system production data, which can be found in 
+#"Projects that receive Performance-Based Incentives (PBI)
+# also provide monthly system production data, which can be found in
 # the Measured Production Data Set below. ""
-prod=pd.read_csv("research/solar_prod/MeasuredProduction_2-25-2015.csv")
 
-prod.columns=["app_num", "prog_admin", "program", "tot_cost", "own_sect", 
-"host_cust_sect","cust_city", "cust_county", "cust_zip", "pbi_payment_num",
-"prod_end_date", "prod_kwh"]
+#%%
 
-#convert date to date format
-prod["prod_end_date"]=pd.to_datetime(prod["prod_end_date"], format="%Y-%m-%d")
-
-#convert production from int to float
-prod["prod_kwh"]=prod["prod_kwh"].astype(float)
+def format_production():
+    prod=pd.read_csv("/Users/johannesmauritzen/research/solar_prod_data/MeasuredProduction_1-24-2018.csv")
 
 
-#link to data on installations
-installations=pd.read_csv("/Users/johannesmauritzen/Google Drive/Research/CalSolar/multilevel_data.csv")
-installations.rename(columns={'Application.Number': 'app_num'}, inplace=True)
+    prod.columns.values
 
-interconnected = pd.read_csv("/Users/johannesmauritzen/research/solar_prod/interconnected_solar.csv")
-interconnected = interconnected[['Application Id', 'Utility', 
-		'Service City', 'Service Zip',
-     	 'Service County', 'Technology Type', 'System Size DC',
-     	  'System Size AC','Mounting Method', 'Tracking',
-       	 'Customer Sector', 'App Received Date', 'App Complete Date', 
-       	 'App Approved Date', 'Installer Name', 'Third Party Owned',
-       	  'Third Party Owned Type', 'Third Party Name', 
[…]

### 2017-12-10 descriptives.py

@@ -1,5 +1,5 @@
 #descriptives.py
-
+#%%
 import pandas as pd
 import numpy as np
 import matplotlib as mpt
@@ -14,73 +14,77 @@ pd.options.display.max_rows = 999
 pd.options.display.max_columns = 50
 
 
-#larger sample but fewer covariates
-prod_long=pd.read_csv("research/solar_prod/prod_long.csv")
-prod_long["prod_end_date"]=pd.to_datetime(prod_long["prod_end_date"])
+# %%
+def set_graphics_params():
+    plt.rcParams['xtick.labelsize'] = 14
+    plt.rcParams['ytick.labelsize'] = 14
+    plt.rcParams['axes.facecolor'] ="white"
+    plt.rcParams['grid.color'] ="grey"
+    plt.rcParams['grid.linestyle'] = "dotted"
+    plt.rcParams["axes.labelsize"]= 14
+    #plt.rcParams['figure.savefig.dpi'] = 100
+    plt.rcParams['savefig.edgecolor'] = "#f2f2f2"
+    plt.rcParams['savefig.facecolor'] ="white"
+    plt.rcParams["figure.figsize"] = [15,8]
+    plt.rcParams['savefig.bbox'] = "tight"
+    plt.rcParams['font.size'] = 14
+    greens = ['#66c2a4','#41ae76','#238b45','#006d2c','#00441b']
+    multi =['#66c2a4','#1f78b4','#a6cee3','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f']
 
+#%%
 
 
 #data file with many covariates, but limited sample
-prod_wide=pd.read_csv("research/solar_prod/prod_wide.csv")
+prod_wide=pd.read_csv("prod_wide.csv")
+prod_wide.head()
+
 prod_wide["prod_end_date"]=pd.to_datetime(prod_wide["prod_end_date"])
-prod_wide["complete_date"]=pd.to_datetime(prod_wide["complete_date"])
+#prod_wide["complete_date"]=pd.to_datetime(prod_wide["complete_date"])
+
 
 wide_cross = prod_wide.groupby("app_num").apply(lambda x: x.iloc[0])
 wide_cross.groupby("program").size()
 
-#just the individual cells
 
+#%%
 
+grouped_solar = solar_data.groupby("app_num")
 
-
-#create max-production variable
-prod_long.reset_index(inplace=True)
-prod_long["prod_max"]=prod_long.groupby("app_num")["prod_kwh"].transform(max)
[…]

### 2017-12-10 lmel_modelling.r

@@ -1,0 +1,251 @@
+#lmel_modelling.r
+#Modelling data using r package lmel
+
+install.packages("tidyverse")
+install.packages("gamm4")
+install.packages("stargazer")
+
+load_libraries = function(){
+  library(lme4)
+  library(tidyverse)
+  library(gamm4)
+  library(mgcv)
+  library(arm)
+  library(stargazer)
+
+}
+
+#initial functions
+load_data = function(){
+  solar_data = read.csv("data/prod_csi.csv")
+  return(solar_data)
+}
+
+show_production = function(solar_data){
+  prod_graph =
+  ggplot(solar_data, aes(x=months_operation, y=prod_scaled, group = app_num)) +
+  geom_line(alpha=.01)
+  return(prod_graph)
+}
+
+format_solar = function(solar_data){
+  solar_data["month"] = as.factor(solar_data$month)
+  solar_data = solar_data[solar_data$module_manufacturer!="Chaori", ]
+  #at least three years of data
+  solar_data = solar_data[solar_data$total_months_operation>=36,]
+  return(solar_data)
+}
+
+#helper functions
+extract_slopes = function(lme_model){
+  #test
+  #lme_model = solar_sys
+  #
+  slope_fe = fixef(lme_model)[13]
+
+  slope_re=data.frame(ranef(lme_model))
+
+  levels(slope_re$term) = c("intercept", "slope")
+  #colnames(slope_re) = c("intercept", "slope")
+  slopes = slope_fe + slope_re$condval[slope_re$term=="slope"]
+
+  se_re = se.ranef(lme_model)
+
+  if(length(se_re)>1){
+    i=1
+    slope_ses <- list(NA, length(se_re))
+
+    for (name in names(se_re)){
+      se_re_df=data.frame(se_re[name])
+      colnames(se_re_df) = c("intercept_se", "slope_se")
+      slope_ses[[i]] = se_re_df[2]
+      i=i+1
+    }
+    slope_ses_df = do.call("rbind", slope_ses)
+  } else{
+    se_re_df = data.frame(se_re[1])
+    colnames(se_re_df) = c("intercept_se", "slope_se")
+    slope_ses_df = se_re_df[2]
+  }
+
+  slopes_df = data.frame(slopes)
+  slopes_df["se"] = slope_ses_df["slope_se"]
+
+  slopes_df["pos_ci"] = slopes_df["slopes"] + 2*slopes_df["se"]
[…]

### 2017-12-10 solar_prod_MLM.py

@@ -36,7 +36,7 @@ def logit_to_prob(value):
 
 # In[5]:
 
-solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_wide_clean.csv")
+solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_csi.csv")
 
 solar_data = solar_data[solar_data.prod_index.notnull()]
 solar_data = solar_data[solar_data.months_operation.notnull()]
@@ -63,7 +63,7 @@ solar_data["log_prod_kwh"] = np.log(solar_data["prod_kwh"])
 # In[7]:
 
 #observation data
-log_prod_kwh = solar_data["log_prod_kwh"] 
+log_prod_kwh = solar_data["log_prod_kwh"]
 month = solar_data["month"]
 
 
@@ -133,7 +133,7 @@ solar_stan_data = {"log_prod_kwh":solar_data["log_prod_kwh"],
 # In[12]:
 
 # def initfun():
-# 	return dict(a=.9, b_1=0, sigma=sd_prod_index, nuMinusOne=1) 
+# 	return dict(a=.9, b_1=0, sigma=sd_prod_index, nuMinusOne=1)
 
 
 # In[19]:
@@ -148,36 +148,36 @@ data{
     int<lower = 0> C; //number of different counties
     int<lower = 0> S; //number of different sectors
     int<lower = 0> T; //number of different trackers
-    
+
 	int installation[N]; //Which of J installations does it belong too
 	vector[N] log_prod_kwh; //response variable
 	vector[N] months_operation; //main predictor
 	int month[N]; // which of 12 months does it belong too.
-    
+
 	int lease[J]; //indicator variable for lease
     int county[J];
     int tracking[J];
     int sector[J];
-    
+
     vector[J] sys_size; //size of system
     vector[J] installation_year; //year of installation
-    
+
 }
 
 transformed data{
     vector[N] st_log_prod_kwh;
     vector[N] st_months_operation;
-    
+
     vector[J] st_sys_size;
     vector[J] st_installation_year;
-    
+
[…]

### 2017-12-10 stan_analysis_new.py

@@ -1,0 +1,287 @@
+#stan_analysis_new.py
+#stan MCMC analysis from 2.5.2017
+
+
+# coding: utf-8
+
+# In[2]:
+
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+
+# In[3]:
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+#get_ipython().magic('matplotlib inline')
+
+
+# In[4]:
+
+def get_percentiles(ser):
+    return(np.percentile(ser, [2.5,15,50,85,97.5]))
+
+def logit_to_prob(value):
+    return(math.exp(value)/(1+math.exp(value)))
+
+
+# In[5]:
+
+solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_csi.csv")
+
+solar_data = solar_data[solar_data.prod_index.notnull()]
+solar_data = solar_data[solar_data.months_operation.notnull()]
+
+app_nums  = solar_data.app_num.unique()
+installation=[i + 1 for i in range(0,len(app_nums))]
+installations = pd.DataFrame({"app_num":app_nums, "installation":installation})
+solar_data = solar_data.merge(installations, how="left", on="app_num")
+
+
+# In[6]:
+
+#age of panel system in data
+N = solar_data.count(axis=0)[0]
+
+app_num = solar_data["app_num"]
+unique_app_num = app_num.unique()
+J = unique_app_num.size
+L = 2
+
+solar_data["log_prod_kwh"] = np.log(solar_data["prod_kwh"])
+
+
+# In[7]:
+
+#observation data
+log_prod_kwh = solar_data["log_prod_kwh"]
+month = solar_data["month"]
+
+
+# In[8]:
+
+#installation level variables
+installations = solar_data[["app_num", "installation", "cust_county", "sector", "tracking", "sys_size_dc", "third_party_owned", "installation_year"]][~solar_data.installation.duplicated()]
+
+
+# In[9]:
+
+#lease variables
[…]

### 2018-02-21 solar_commodities.tex
update

@@
 
-\title{Are Solar Panels Commodities? Evidence of Quality Differences and Assymetric Information from California}
+\title{Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California}
 
@@
  Solar panels should not be considered commodities.
-Considerable quality differences, as measured by average production over time, are found between manufacturers.
+Considerable quality differences, as measured by degradation of production over time, are found between manufacturers.
 This has implications for pricing and competition in the market for solar panel systems.
 I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with third party owners display higher quality than those owned directly.
-2.) Prices of solar panels are not related to quality for solar panels that are owned directly.
+2.) Prices of solar panels are less correlated to quality for solar panels that are owned directly.
 I use multilevel models estimated by both maximum likelihood and Bayesian Markov Chain Monte-Carlo.
@@
 This can be measured as both the failure rate of these panels as well as the degree of degradation of the panels over time.
-This can imply that certain manufacturers with superior quality could extract a premium.
+This can imply that certain manufacturers with superior quality could extract a premium in the market.
 
@@
 
-Testing for the degradation in production of nearly 1000 panels is however problematic using traditional econometric models based on frequentist approaches.
+Testing for the degradation in production of more than 3000 panels is however problematic using traditional econometric models based on frequentist approaches.
 The production patterns of the solar panels could be expected to vary greatly depending on location, equipment, size as well as other factors.
@@
 
-\section{California Solar Initiative, Data and Empirical Model}
+\section{California Solar Initiative, Data}
 
@@
 The installation data can be matched with the production data for those installations receiving a production subsidy.
-I include only those installations that have been producing for at least 4 years (48 monthly observations), where the maximum number of observations in this data is 5 years (60 months) of production data.
-The data then covers similar vintage of solar panels - those that were installed from 2007 through 2009 and which continued to produce through 2013.
+I include only those installations that have been producing for at least 3 years (36 monthly observations), where the maximum number of observations in this data is 5 years (60 months) of production data.
 
@@
 
-For later comparison, consider results from a traditional single-level regression of production over time estimated by maximum likelihood shown in table \ref{table:max_likelihood}.
-Fixed effects are included for month, host sector, county\footnote{coefficients for county are not shown}, and an indicator variable for leased systems.
-Variables for year of installation and size of the system are also included.
-The variable of interest is the interaction of the lease indicator with the time variable - which is transformed from monthly to fractional year units.
+\section{Linear Mixed Effects Models Estimated by Maximum Likelihood}
 
-The results indicate a good fit with an $R^2$ of 90 percent.
-The coefficient of interest indicates that leased systems show 4\% less degradation over time than those sold outright - with a p-value of .001.
-However, even with the relatively large number of fixed effects included in this regression, much of the idiosyncratic variation from one photovoltaic system to another is not accounted for.
-I will show that when the full idiosyncratic variation is better accounted for, the evidence for a substantial difference in performance over time between leased and systems sold outright is substantially weaker than the below regression results indicate.
+\subsection{Testing for differences between module manufacturer}
+Following the notation of \citet{bates_fitting_2015}, the conditional distribution of the response variables, $y$ in a multilevel model can be written compactly as in equation \ref{equation:lme_simple}.
+Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or "fixed effects") of the $n times p$ matrix of (p-1) explanatory variables $\mathbf{X}$ plus intercept term.
+$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or "random effects" $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of "random" refers to the fact that these groups are assumed to be specific to the sample.
+Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms, thus the "random" label.
+Still, interpreting coefficients as group varying or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.
+$\mathbf{W}$ is a diagonal matrix of known prior weights.
+$B$ is distributed as multivariable normal with zero mean and where $\Sigma$ is variance-covariance matrix.
+
+\begin{align}
+(y|\mathbf{B=b} &\sim \mathcal{N}(\mathbf{X\beta + Zb}, \sigma^2\mathbf{W^{-1}}))\\
+B &\sim \mathcal{N}(\mathbf{0,\Sigma})
+\label{equation:lme_simple}
+\end{align}
+
+Models of this form can be solved efficiently using a penalized maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.
+
+To test whether I can detect quality differences between module manufacturers, I need to first estimate a restricted comparison model.
+I write the restricted comparison model explicitly in equations \ref{eqn:observation_level_rest} and \ref{eqn:system_level_rest}.
+
+In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modelled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$.
+Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by the individual system.
+Thus in this model formulation I take into account the groupings by system, explicitly modelling varying relative initial production values (intercepts) and degradation over time (slope).
+Moving up a level, the estimated $a$ and $b$ parameters are in turn modelled as multivariate normal.
+
+\begin{align}
+prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+\label{eqn:observation_level_rest}
+\begin{pmatrix}
+  a\\
+  b
+\end{pmatrix}
+&\sim N
+\begin{pmatrix}
+  \mu_a\\
+  \mu_b
+\end{pmatrix},
+\begin{pmatrix}
+  \sigma_a^2 & \rho \sigma_a \sigma_b \\
+  \rho \sigma_a \sigma_b & \sigma_b^2
+\end{pmatrix}
+\label{eqn:system_level_rest}
+\end{align}
+
+The inclusion of the matrix of month dummies $\mathbf{month}$ controls for seasonal variation across all systems.
+This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and the seasonality is not of direct importance, this should suffice.
+As long as un-modeled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.
+
+To test whether quality varies among module manufacturers, I simply add panel manufacturer as an extra hierarchy to the model.
+In this model, the observation level regression is identical to the restricted comparison model as in equation \ref{eqn:observation_level_rest}.
+The system-level intercepts, $a$, and slope, $b$, terms are also modelled as a multivariate normal.
+But the mean values of the multivariate normal distribution, $mu_a$ and $mu_b$, are now indexed by $m$, representing that they are allowed to vary by module manufacturer.
+\begin{align}
+\begin{pmatrix}
+  a\\
+  b
+\end{pmatrix}
+&\sim N
+\begin{pmatrix}
+  \mu_{a[m]}\\
+  \mu_{b[m]}
+\end{pmatrix},
+\begin{pmatrix}
+  \sigma_{a[m]}^2 & \rho \sigma_{a[m]} \sigma_{b[m]} \\
+  \rho \sigma_a \sigma_b & \sigma_b^2
+\end{pmatrix} \\
+\mu_{a} &\sim N(\theta_a, \sigma_{mu_a}^2) \\
+\mu_{b} & \sim N(\theta_b), \sigma_{mu_b}^2)
+\end{align}
+
+If there is no significant variation between the module manufacturer groupings, this model collapses into the restricted comparison model, which is the null hypothosis.
+Figure \ref{sfig:sys_slope_fig} gives a visual summary of the system-level slope terms ordered from lowest to highest.
+The lines represent approximate 95\% confidence intervals around the point estimate.
+One anomaly we can notice is that a handful of the systems have slopes that are estimated to be positive.
+This is clearly problematic if we wish to interpret the slope coefficient as degradation over time.
+I address this in the full Bayesian model without resorting to excluding data.
+
+\begin{figure}
+\begin{minipage}{.48\textwidth}
+  \includegraphics[width=1\linewidth]{figures/sys_slope_fig.png}
+  \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, ordered from lowest to highest.}
+  \label{sfig:sys_slope_fig}
+\end{minipage}\hfill
+\begin{minipage}{.48\textwidth}
+  \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
+  \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level slope terms $\mu_b$, ordered from lowest to highest.}
+  \label{sfig:nested_manuf_fig}
+\end{minipage}
+\end{figure}
+
+\ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes.
+Again the lines represent 95\% confidence intervals around the point estimates.
+Visually, we can see substantial variation in quality between the manufacturer groups.
+We can more formally test the hypothesis that manufacturer groupings improve the explanatory power of the model by comparing to the restricted model with an LM test.
+The results are shown in table \ref{tbl:lm_commodity}.
+The information criteria measures (AIC, BIC) are both lower, indicating a better predictive power of the model with manufacturer groupings.
+The chis-square statistic also indicates that the model with manufacturer groupings substantially improves model performance.
 
 \begin{table}
-\begin{center}
-\begin{tabular}{l c }
-\hline
- & Maximum Likelihood Results\\
-\hline
-(Intercept)                                            & $8.62^{***}$  \\
-                                                       & $(0.08)$      \\
-I(months\_operation/12)                                & $-0.03^{***}$ \\
-                                                       & $(0.00)$      \\
-Leased                         & $0.19^{***}$  \\
-                                                       & $(0.01)$      \\
-sys\_size\_dc                                          & $0.00^{***}$  \\
-                                                       & $(0.00)$      \\        & $(0.08)$      \\
-February                                         & $0.14^{***}$  \\
-                                                       & $(0.02)$      \\
-March                                        & $0.27^{***}$  \\
-                                                       & $(0.02)$      \\
-April                                        & $0.59^{***}$  \\
-                                                       & $(0.02)$      \\
-May                                         & $0.70^{***}$  \\
-                                                       & $(0.02)$      \\
-June                                         & $0.79^{***}$  \\
-                                                       & $(0.02)$      \\
-July                                         & $0.75^{***}$  \\
-                                                       & $(0.02)$      \\
-August                                         & $0.70^{***}$  \\
-                                                       & $(0.02)$      \\
-September                                         & $0.69^{***}$  \\
-                                                       & $(0.02)$      \\
-October                                       & $0.54^{***}$  \\
-                                                       & $(0.01)$      \\
-November                                       & $0.41^{***}$  \\
-                                                       & $(0.01)$      \\
-December                                        & $0.15^{***}$  \\
-                                                       & $(0.01)$      \\
-Fixed Array                            & $-0.52^{***}$ \\
-                                                       & $(0.03)$      \\
-Mixed                                  & $-0.34^{***}$ \\
-                                                       & $(0.06)$      \\
-Single-Axis Tracker                    & $-0.31^{***}$ \\
-                                                       & $(0.04)$      \\
-Government                               & $0.07^{***}$  \\
-                                                       & $(0.01)$      \\
-Non-Profit                               & $-0.17^{***}$ \\
-                                                       & $(0.02)$      \\
-Residential                              & $-2.12^{***}$ \\
-                                                       & $(0.01)$      \\
-I(months\_operation/12):leased & $0.04^{***}$  \\
-                                                       & $(0.01)$      \\
-\hline
-R$^2$                                                  & 0.90          \\
-Adj. R$^2$                                             & 0.90          \\
-Num. obs.                                              & 32979         \\
-RMSE                                                   & 0.57          \\
-\hline
-\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$}}
+  \begin{tabular}{rllllllll}
+  \toprule
+    Model &  DF &   AIC &   BIC &  logLik &  deviance &  Chisq &  Chi df &  Pr(>Chisq) \\
+    \midrule
+             Null &  17 &  8805 &  8976 &   -4385 &      8771 &     -- &      -- &          -- \\
+    Grouped Manuf. &  20 &  8430 &  8631 &   -4195 &      8390 &    381 &       3 &     2.6e-82 \\
+  \bottomrule
   \end{tabular}
-\caption{Results from a standard linear regression model with fixed effects for }
-\label{table:max_likelihood}
-\end{center}
+  \label{tbl:lm_commodity}
 \end{table}
 
-Instead of standard single-level maximum likelihood estimation,  I use a hierarchical Bayesian model where parameters can be estimated through Markov Chain Monte Carlo (MCMC) simulation techniques.
-The main advantage of such models are the flexibility of taking into account the multi-level structure of the data and improved inference that comes from partial pooling.
+The first result, of which the evidence appears to be quite strong is that solar panel systems do appear to vary substantially in quality between module manufacturers.
+Given that module quality is an important differentiating characteristic of panels, it is then difficult to formally define solar panels as commodities.
 
-Consider first a no pooling model.
-The slope of production over time from each solar panel system would be estimated independently - allowing for large amounts of solar-power-system level variance, but excluding information from data on all the other plants from each of the individual system-level inferences.
-On the other hand, in a fully pooled model, the data from all of the observations of all the solar power systems in each group (leased, sold outright) are used in the estimation, but idiosyncratic variation between solar-power-systems are not taken into account.
-An outlying solar power system may, for instance, have too much influence and bias the results.
+\subsection{testing for differences in quality between leased and host-owned solar power systems}
 
-A hierarchical model with partial pooling is a compromise between these two extremes.
-Individual estimators of the solar-power-system slopes make use of the information in the full dataset with weights established by the relative variance between and within solar power systems.
-A full discussion of partial pooling can be found in \citet{gelman_bayesian_2013} or \citet{kruschke_doing_2014}.
+Provided the evidence that solar panel systems are not commodities and that substantial differences in quality do exist between manufacturers, we can now move on to the testable implication that comes from a combination of standard economic theory of asymmetric information and a descriptive understanding of the market that suggests the market for solar panels could plausibly be subject to issues of asymmetric information of quality.
+
+From our informal model of asymmetric information where we see host owners of solar systems as being low-information types, while owners of leased panels as being high information types, we would predict that on average leased panels will display higher quality over time.
+To test this directly with the data, I only need to do a slight modification of the model in the previous subsection.
+Instead of grouping the system-level coefficients by manufacturer, I now group into whether the solar power system was leased or not, as indexed by $l$.
+We can then compare the average slopes between these groups.
+For convenience the model is presented in full in equations \ref{eqn:observation_level_rest2}.
+
+\begin{align}
+prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+\label{eqn:observation_level_rest2}
+\begin{pmatrix}
+  a\\
+  b
+\end{pmatrix}
+&\sim N
+\begin{pmatrix}
+  \mu_{a[l]}\\
+  \mu_{b[l]}
+\end{pmatrix},
+\begin{pmatrix}
+  \sigma_{a[l]}^2 & \rho \sigma_{a[l]} \sigma_{b[l]} \\
+  \rho \sigma_a \sigma_b & \sigma_b^2
+\end{pmatrix} \\
+\mu_{a} &\sim N(\theta_a, \sigma_{mu_a}^2) \\
+\mu_{b} & \sim N(\theta_b), \sigma_{mu_b}^2)
+\label{eqn:lease_model}
+\end{align}
+
+A visual summary of the results are shown in figure \ref{lease_sys_fig}.
+Again, the vertical lines represent approximate 95\% confidence intervals over the point estimates for the system-level slope terms, $b_s$.
+But here the the slope estimates are grouped into host-owned (left) and leased (right).
+The estimated group means are shown as horizontal lines.
+At the scale of the figure, the effect appears small, but the average slopes are estimated to be significantly different from each other at the 95\% confidence interval.
+The difference is also economically significant.
+Extrapolating out from the difference in the point estimates between leased and host-owned suggests that on average a leased system can expect to be producing 11\% more power after 10 years compared to an otherwise identical solar panel systems.
+
+\begin{figure}
+	\includegraphics[width=.8\textwidth]{figures/lease_sys_fig.png}
+	\caption{A visual summary of the estimation results. The vertical lines again represent the estimated system-level slope coefficients.}
+	\label{lease_sys_fig}
+\end{figure}
+
+\section{Full Bayesian multilevel model estimated by Markov Chain Monte Carlo}
+
+Until now, I have used relatively parsimonious models with few fixed-effects covariates and only one or maximum two hierarchies ("random effects").
+Ideally, I would like to estimate a full probability model with all relevant and available covariates and hierarchies.
+First and foremost so that the parameters of interests to the testable implications can be estimated from one model.
+But also, citet{bar_random_2013} shows using monte-carlo simulations that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
+Estimating mixed effects model by maximum likelihood is efficient and straight-forward.
+However, convergence of the maximum likelihood routine is not guaranteed under richer models with more hierarchies and larger number of paramaters to be estimated.
+
+Specific to my model, I wish to estimate a model where the variable for cost per-kw is included as an interaction that varies by system and module manufacturer.
+I also wish to have a model where both system, module manufacturer and leasing hierarchies are estimated simultaneously.
+I am not able to obtain convergence of the maximum likelihood routines under this rich model.
+
+As a solution, I build a fully Bayesian multilevel model\footnote{I purposefully change my nommenklature here and refer to the Bayesian model as a "multilevel model" to distinquish it from the models estimated by maximum likelihood, which I refer to as a linear mixed effects model.
+This distinction is however mostly artificial.
+Both models resemble each other closely in structure.
+"Multilevel model" is however a broader term.
+There is no restriction to linearity, for example.
+Models with non-linear or non-parametric components can also be incorporated and estimated with MCMC.} estimated by Markov Chain Monte Carlo (MCMC) simulation techniques.
+
+The main advantage to using the Bayesian techniques here are practical rather than philosophical.
+Models up to levels of nearly arbitrary richness can be estimated, with computation time and available data being the main constraints.
+Inference is also well defined and exact,\footnote as opposed to the approximate inference from maximum likelihood.
+
+A Bayesian MCMC approach does also have some attractive properties.
+Results are presented estimates of marginal posterior distributions over coefficients of interest, which can be interpreted directly and intuitively as probabilities.
+Group-level variances are not constrained to be equal and assumptions of normality of the likelihood and posterior distributions can be loosened.
+Several excellent discussions of Bayesian techniques are available, see \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mc}
 
@@
 
-% \begin{align}
-% 		log(production_{ij}) &\sim N(\hat{y}, \sigma^2), \quad i=1,...,n \\
-% 		\hat{y} &= \beta_{0j} + \beta_{1j}*time_{ij} + month_{m}
-% 		\beta_{1,j} &= \mu_{\beta, l} + \zeta_{b1, j}, \quad j=1...j, \quad l = 0,1 \\
-% 		\mu_{\beta1, l} &\sim N(0,10) \\
-% 		\beta_{0,j} &\sim N(\mu_{b0}, \sigma_{b0}) \\
-% 		\month_{m} &\sim N(1,10)\\
-% 		\sigma &\sim uniform(0,100) \\
-% 		\zeta_{b1, j} &\sim N(0,\sigma_{b1})
-% 		\label{equation:model}
-% 		\caption{}
-% \end{align}
-
-\section{Results}
-
-I run 1000 iterations with 4 chains of the Hamiltonian MCMC algorithm in order to sample the joint posterior probability distribution of the model.
-I can then extract the the estimates of the marginal posterior distributions of the individual parameters in the form of a list of simulated parameter estimates.
-
-Summary statistics for the posterior distributions of the main parameters of the model are shown in table \ref{table_results}.
-The parameters of interest are $\mu_{own}, \mu_{lease}$ which are the average slope of production over time in, respectively, systems that were sold outright and systems that were leased.
-These coefficients as well as the other component coefficients of $\beta1$, should be interpreted for a marginal change of 1 standard deviation change in time.
-Intuitive illustrations of the magnitudes of the coefficients of interest are shown further below.
-
-\begin{table}
-\resizebox{12cm}{!} {
-\begin{tabular}{lrrrr}
-\toprule
-{} &  2.5\%&   97.5\% &   mean &  median \\
-params           &          &          &         &          \\
-\midrule
-apr              &   0.1452 &  -0.0838 &  0.0069 &  -0.0013 \\
-aug              &   0.2065 &  -0.0223 &  0.0681 &   0.0594 \\
-dec              &  -0.1005 &  -0.3295 & -0.2386 &  -0.2471 \\
-feb              &  -0.1037 &  -0.3346 & -0.2437 &  -0.2523 \\
-jan              &  -0.1935 &  -0.4234 & -0.3319 &  -0.3404 \\
-jul              &   0.2364 &   0.0058 &  0.0968 &   0.0886 \\
-jun              &   0.2584 &   0.0282 &  0.1191 &   0.1105 \\
-mar              &  -0.0257 &  -0.2535 & -0.1631 &  -0.1717 \\
-may              &   0.2067 &  -0.0225 &  0.0687 &   0.0602 \\
-nov              &   0.0546 &  -0.1748 & -0.0839 &  -0.0923 \\
-oct              &   0.1256 &  -0.1034 & -0.0124 &  -0.0211 \\
-sep              &   0.2094 &  -0.0205 &  0.0704 &   0.0619 \\
-mu1\_commercial   &   0.0014 &  -0.0041 & -0.0007 &  -0.0005 \\
-mu1\_government   &   0.0035 &  -0.0026 &  0.0004 &   0.0002 \\
-mu1\_non\_profit   &   0.0018 &  -0.0057 & -0.0008 &  -0.0003 \\
-mu1\_not\_tracking &   0.0044 &  -0.0074 & -0.0016 &  -0.0009 \\
-mu1\_residential  &   0.0039 &  -0.0023 &  0.0005 &   0.0003 \\
-mu1\_tracking     &   0.0039 &  -0.0092 & -0.0021 &  -0.0014 \\
-mu\_lease         &  -0.0001 &  -0.0115 & -0.0045 &  -0.0042 \\
-mu\_own           &  -0.0001 &  -0.0125 & -0.0054 &  -0.0054 \\
-beta0\_size       &   0.7501 &   0.6948 &  0.7173 &   0.7141 \\
-beta1\_ins\_year   &  -0.0007 &  -0.0032 & -0.0020 &  -0.0020 \\
-beta1\_size       &  51.1251 & -45.7583 &  0.1312 &   0.0594 \\
-sigma            &   0.0295 &   0.0287 &  0.0291 &   0.0291 \\
-\bottomrule
-\end{tabular}
-}
-\label{table_results}
-\caption{Summary statistics of the distributions over the main parameters in the model.
-Confidence intervals are estimated as quantiles at 2.5 and 97.5 percent of the estimated posterior distribution of the parameters.}
-\end{table}
-
-I display the histograms representing the posterior distributions of these two parameters in the top panel of figure \ref{diff_lease}.
-The histograms indicate that the posterior distributions appear to have means that are different from each other.
-
-In the lower panel, I show a histogram of the posterior of the difference between the two parameters, $\mu_{b1,l=1} - \mu_{b1,l=2} $ which is the direct test of the question of interest.
-95\% of the probability lies below the vertical line, which stands at approximately zero.
-A direct interpretation is then that there is a 95 \% probability that on average solar power systems that are sold outright have a higher rate of degradation than those that are leased.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{figures/diff_lease.png}
-	\caption{The figure shows the distribution of the difference of $\mu_{lease}$ and $\mu_{own} $.
-The black vertical line is placed at zero.
-The distribution of the difference of the two parameters suggests that leased panels had about a 80\% chance of being higher quality than those sold outright, as measured by the slope of production over time.}
-	\label{diff_lease}
-\end{figure}
-
-The magnitude of the effect can best be seen by plotting average predicted values over time, which are shown for systems that are leased and sold-outright in figure \ref{predicted_deg}.
-The solid lines represent the median value of the posterior distribution on the $mu_lease$ and $mu_old$ parameters, where the light lines represent random draws from the respective posterior distributions to represent uncertainty.
-The model suggests that a leased solar panel system will on average experience degradation of about 6\% after 20 years while those owned outright experience on average of 5\% of degradation.
-However, as the figure makes clear, the average values hide significant variation in the overall distribution.
-
-\begin{figure}
-	\includegraphics[width=1\textwidth]{figures/predicted_deg.png}
-	\caption{The figure shows the average predicted values of degradation over time of solar panel systems that are leased and sold out-right.
-The solid lines represent the mean value of the posterior where the lighter lines represent random draws from the posterior distribution}
-	\label{predicted_deg}
-\end{figure}
-
 \section{Discussion and Conclusion}
@@
 The results provide some weak evidence for the existence of information asymmetry issues of quality.
-
-The use of Bayesian MCMC simulation techniques in this article also provides an illustrative example of the use and usefullnes of a powerful emerging toolset.
-The Bayesian frameworks and recently developed software such as Stan provide major practical and theoretical advantages over many traditional econometric methods.
-These methods and tools are also increasingly accessible to the practicioner.
-\citet{kruschke_doing_2014} and \citet{stan_development_team_stan_2014} provide accessible and thorough introductions and guides.
 

### 2018-02-21 descriptives.py

@@ -33,6 +33,15 @@ def set_graphics_params():
 
 #%%
 
+set_graphics_params()
+solar_data = pd.read_csv("data/prod_csi.csv")
+
+comm_table = [["Model",	"DF",	"AIC",	"BIC",	"logLik",	"deviance",	"Chisq",	"Chi df",	"Pr(>Chisq)"],
+["Null",	"17",	"8805",	"8976",	"-4385",	"8771",	"--",	"--",	"--"],
+["Grouped Manuf.", 	"20",	"8430",	"8631",	"-4195",	"8390",	"381",	"3",	"2.6e-82"]]
+
+comm_table_df = pd.DataFrame(comm_table)
+comm_table_df.to_latex(buf="table3.tex")
 
 #data file with many covariates, but limited sample
 prod_wide=pd.read_csv("prod_wide.csv")

### 2018-02-21 lmel_modelling.r

@@ -137,9 +137,23 @@ lease_model = function(solar_data){
 }
 
 lease_model2 = function(solar_data){
-  solar_lme4 = lmer(prod_scaled ~ month + months_operation_scaled + cost_per_kw (months_operation_scaled + cost_per_kw |third_party_owned) + (months_operation_scaled|third_party_owned:app_num), solar_data)
+  #test if prices are more correlated with
+  solar_lme4 = lmer(prod_scaled ~ month + months_operation_scaled*third_party_owned + (months_operation_scaled*third_party_owned|app_num), solar_data)
   return(solar_lme4)
 }
+
+#useing GAM from GAMM4
+gamm_model_man = function(solar_data){
+
+  solar_data["month"] = as.numeric(as.character(solar_data$month))
+  solar_gamm4 = gamm4(prod_scaled ~ s(month) + months_operation_scaled, dat=solar_data, random=~(months_operation_scaled|module_manufacturer))
+  return(solar_gamm4)
+
+}
+
+
+
+
 
 main = function(){
   load_libraries()
@@ -150,9 +164,16 @@ main = function(){
   lm_model = lm(prod_scaled~months_operation, data=solar_data)
   lm_coef = coef(lm_model)
 
-  prod_fig + geom_abline(intercept=lm_coef[1], slope=lm_coef[2], color="red") + xlim(0,60) + ylim(-1.5,1.5) + labs(x="months",y="Production, de-scaled")
+  prod_fig = prod_fig + geom_abline(intercept=lm_coef[1], slope=lm_coef[2], color="red") + xlim(0,60) + ylim(-1.5,1.5) + labs(x="months",y="Production, de-scaled")
 
-  ggsave("figures/production.png", plot=prod_fig, dpi=300)
+  ggsave("figures/production.png", plot=prod_fig, dpi=100,width = 9, height = 5)
+
+  #sum summary statistics
+  solar_data[c("prod_kwh", "first_prod_year", "cost_per_kw", "csi_rating")]
+  table(solar_data["third_party_owned"])
+
+
+  59108/(111919+59108)
 
   #model with manufacturers
   solar_lme2 = lme2_model(solar_data)
@@ -160,7 +181,17 @@ main = function(){
   man_fig = slope_plot(slopes_df_man, no_xlabs=TRUE)
[…]

### 2018-02-21 stan_analysis_new.py

@@ -1,10 +1,5 @@
 #stan_analysis_new.py
-#stan MCMC analysis from 2.5.2017
-
-
-# coding: utf-8
-
-# In[2]:
+#stan MCMC analysis from 2.6.2017
 
 import pandas as pd
 import tabulate as tab
@@ -20,15 +15,13 @@ import pystan
 import scipy
 import pickle
 
-
-# In[3]:
-
 pd.options.display.max_rows = 999
 pd.options.display.max_columns = 50
+
+
+
 #get_ipython().magic('matplotlib inline')
 
-
-# In[4]:
 
 def get_percentiles(ser):
     return(np.percentile(ser, [2.5,15,50,85,97.5]))
@@ -36,252 +29,433 @@ def get_percentiles(ser):
 def logit_to_prob(value):
     return(math.exp(value)/(1+math.exp(value)))
 
+def load_data():
+    solar_data = pd.read_csv("data/prod_csi.csv")
+    return(solar_data)
 
-# In[5]:
+def format_data(solar_data):
+    solar_data = solar_data[solar_data.prod_scaled.notnull()]
+    solar_data = solar_data[solar_data.months_operation_scaled.notnull()]
+    #limit to data over three years
+    solar_data = solar_data[solar_data.total_months_operation>=36]
+    return(solar_data)
 
-solar_data = pd.read_csv("/Users/johannesmauritzen/research/solar_files/prod_csi.csv")
+def descale_data(solar_series):
+    return((solar_series - np.mean(solar_series))/(2*np.std(solar_series)))
 
-solar_data = solar_data[solar_data.prod_index.notnull()]
-solar_data = solar_data[solar_data.months_operation.notnull()]
+def transform_log(solar_data):
+    solar_data.loc[:,"log_prod_kwh"] = np.log(solar_data.prod_kwh+ 0.00001)
+    solar_data.loc[:,"log_cost_per_kw"] = np.log(solar_data.cost_per_kw + 0.00001)
+    solar_data.loc[:,"log_csi_rating"] = np.log(solar_data.csi_rating + 0.00001)
+    return(solar_data)
 
-app_nums  = solar_data.app_num.unique()
-installation=[i + 1 for i in range(0,len(app_nums))]
-installations = pd.DataFrame({"app_num":app_nums, "installation":installation})
-solar_data = solar_data.merge(installations, how="left", on="app_num")
+def create_num_dict(sseries):
+    skeys  = sseries.unique()
+    #numbered from 1 to N
[…]

### 2018-02-21 stan_helpers.py

@@ -9,12 +9,12 @@ def get_cent_interval(sample, conf=.95):
 	high_ci = sample[round(n_sample*(1-p/2))]
 	return([low_ci, high_ci])
 
-def summary_info(post_sample, lim=5):	
+def summary_info(post_sample, lim=5):
 	#post_sample = solar_params
 	#lim=5
 	"""
 	Input extracted STAN model sample: post_sample
-	lim - confidence interval ie lim = 5 => 95 central posterior interval  
+	lim - confidence interval ie lim = 5 => 95 central posterior interval
 	"""
 
 	params = []
@@ -34,7 +34,7 @@ def summary_info(post_sample, lim=5):
 			param_sim=np.sort(param_sim)
 			low_ci.append(param_sim[n_tail])
 			high_ci.append(param_sim[n_samples-n_tail])
-		else:	
+		else:
 			for par_case in param_sim.T:
 				params.append(i)
 				means.append(np.mean(par_case))

### 2018-02-21 stan_results_analysis.py

@@ -1,0 +1,187 @@
+#Stan_results_analysis
+#last updated 17.2.2018
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+pd.options.display.max_rows = 999
+pd.options.display.max_columns = 50
+
+plt.rcParams['xtick.labelsize'] = 14
+plt.rcParams['ytick.labelsize'] = 14
+plt.rcParams['axes.facecolor'] ="white"
+plt.rcParams['grid.color'] ="grey"
+plt.rcParams['grid.linestyle'] = "dotted"
+plt.rcParams["axes.labelsize"]= 14
+#plt.rcParams['figure.savefig.dpi'] = 100
+plt.rcParams['savefig.edgecolor'] = "#f2f2f2"
+plt.rcParams['savefig.facecolor'] ="white"
+plt.rcParams["figure.figsize"] = [15,8]
+plt.rcParams['savefig.bbox'] = "tight"
+plt.rcParams['font.size'] = 14
+greens = ['#66c2a4','#41ae76','#238b45','#006d2c','#00441b']
+multi =['#66c2a4','#1f78b4','#a6cee3','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f']
+#plt.rcParams["axes.color_cycle"] = multi
+
+
+
+solar_file = open("/Users/johannesmauritzen/research/solar_prod_data/stan_extracts/solar_extr_mac_small.pkl", "rb")
+solar_extr = pickle.load(solar_file)
+
+solar_extr.keys()
+
+mu1_leaseDF = pd.DataFrame(solar_extr["mu1_lease"])
+beta1_costDF = pd.DataFrame(solar_extr["beta1_cost"])
+
+
+mu1_leaseDF.head()
+beta1_costDF.head()
+
+beta1_costDF.hist()
+plt.show()
+
+mu1_leaseDF["mu_diff"] = mu1_leaseDF[1]-mu1_leaseDF[0]
+beta1_costDF["beta_diff"] = beta1_costDF[1]-beta1_costDF[0]
+
+beta1_costDF.head()
[…]

### 2018-03-22 solar_commodities.tex
update, first draft of results

@@
 \end{pmatrix}
-\label{eqn:system_level_rest}
+%\label{eqn:system_level_rest}
 \end{align}
@@
 \section{Full Bayesian multilevel model estimated by Markov Chain Monte Carlo}
-
 Until now, I have used relatively parsimonious models with few fixed-effects covariates and only one or maximum two hierarchies ("random effects").
@@
 First and foremost so that the parameters of interests to the testable implications can be estimated from one model.
-But also, citet{bar_random_2013} shows using monte-carlo simulations that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
+But also, \citet{bar_random_2013} shows, using Monte-Carlo simulations, that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
 Estimating mixed effects model by maximum likelihood is efficient and straight-forward.
-However, convergence of the maximum likelihood routine is not guaranteed under richer models with more hierarchies and larger number of paramaters to be estimated.
+However, convergence of the maximum likelihood routine is not guaranteed under richer models with more hierarchies and larger number of parameters to be estimated.
 
-Specific to my model, I wish to estimate a model where the variable for cost per-kw is included as an interaction that varies by system and module manufacturer.
+Specific to my model, I wish to estimate a model where the variable for cost per-kW is included as an interaction that varies by system and module manufacturer.
 I also wish to have a model where both system, module manufacturer and leasing hierarchies are estimated simultaneously.
@@
 
-As a solution, I build a fully Bayesian multilevel model\footnote{I purposefully change my nommenklature here and refer to the Bayesian model as a "multilevel model" to distinquish it from the models estimated by maximum likelihood, which I refer to as a linear mixed effects model.
+As a solution, I build a fully Bayesian multilevel model\footnote{I purposefully change my nomenclature here and refer to the Bayesian model as a "multilevel model" to distinguish it from the models estimated by maximum likelihood, which I refer to as a linear mixed effects model.
 This distinction is however mostly artificial.
@@
 Models up to levels of nearly arbitrary richness can be estimated, with computation time and available data being the main constraints.
-Inference is also well defined and exact,\footnote as opposed to the approximate inference from maximum likelihood.
+Inference is also well defined and exact.\footnote{Exact in the sense that given a correctly specified probability model, then an MCMC routine that is given infinite time to sample, will in theory converge to the correct posterior probability distribution.
+Maximum likelihood is approximate in the sense that inference is based on asymptotic theory and assumptions of normality.}
 
-A Bayesian MCMC approach does also have some attractive properties.
-Results are presented estimates of marginal posterior distributions over coefficients of interest, which can be interpreted directly and intuitively as probabilities.
+A Bayesian MCMC approach has some additional attractive properties.
+Results are presented as estimates of marginal posterior distributions over coefficients of interest, which can be interpreted directly and intuitively as probabilities.
 Group-level variances are not constrained to be equal and assumptions of normality of the likelihood and posterior distributions can be loosened.
-Several excellent discussions of Bayesian techniques are available, see \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mc}
+For further discussion of the Bayesian frameworks and techniques, I refer to the excellent discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2017}.
 
-Figure \ref{solar_prod_bayes_diag} shows the structure of the model.
-Starting from the bottom, the individual production data are log transformed and modeled as having a Cauchy distribution with mean $y_hat$ and variance $sigma^2$.
-$sigma^2$ is in turn given a half-Cauchy prior.
+The model is presented in equations x through y.
+The response variable is production data that has been log transformed and scaled as described earlier.
+This response variable is assumed to have a normal likelihood with mean, $\hat{y}$ and standard deviation, $\sigma$.
+At the observation level, the fitted values, $\hat{y}$, are modelled as in equation \ref{obslevel}.
+This observation level regression, as before, is a simple linear regression with an intercept $\alpha$, and a slope $\beta$ which are both allowed to vary by system.
+In addition a vector of month dummies, $\mathbf{month}$ is included in order to control for seasonality.
 
-The use of the Cauchy and half-Cauchy distribution reflects that a large in magnitude effect of around 5 or greater on the logit scale is highly unlikely in logistic regressions where all non-binary data has been transformed to have mean zero and standard deviation 1 - which I have done.
-In addition, the Cauchy distribution will give answers even under complete separation, and avoids computational problems inherent in assigning completely non-informative priors in multilevel models \citep{gelman_weakly_2008}.
+Equations \ref{eqn:sysa} and \ref{eqnsysb} describe the model at the system level.
+The intercept terms, $a_s$ are given a common higher level normal distribution with mean $\mu^a$ and standard deviation $sigma^a$.
+But because the intercept term is not of direct economic importance, it is not modelled further.
 
-Going up a level, $y_hat$ is modeled as having random group-level intercepts $\beta_{0,j}$, where j represents the j-th of $J$ solar panel system.
-The $\beta_{0,j}s$ are in turn modeled as function of system size, county, whether solar trackers are used, and a random effects term $re_j^0$.
+The slope terms, $b_s$ are again modelled directly.
+I decompose the variation into three mean terms that vary by lease, $\mu^{lease}_l$, host sector $\mu^{sect}_c$, and manufacturer $\mu^m_m$.
+In addition three continuous variables are included in the system-level equation: the year of first production, $first\_prod\_year_s$, log capacity, $log\_capacity_s$ and the log system cost per kilowatt, $log\_cost_s$ with corresponding coefficients $\beta^{fy}$, $\beta^{size}$, and $\beta^{cost}$.
 
-In a similar manner the J $B_j$ coefficients are modeled as a function of whether the system is leased, $\mu_{lease}^1$, the host sector of the installation, the county, use of tracking, installation year and a random effects term $re^1_j$.
-The question of interest - whether leased solar systems display a lower degree of degradation over time than those sold outright - can then be expressed as the distribution of the difference of the $mu_{lease}$ parameters: $\mu_{lease=1} - \mu_{lease=0}$.
+\begin{align}
+log_prod &\sim N(\hat{y}, \sigma)\\ %\label{eqn:likelihood}
+\hat{y} &= a_s + b_s months\_operation_i + \mathbf{month}\\ %\label{eqn:obslevel}
+a_s &\sim N(\mu^a, \sigma^a)\\ %\label{eqn:sysa}
+b_s & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \beta^{fy} first\_prod\_year_s + beta^{size} log\_csi\__rating_s + \beta^{cost} log\_cost_s %\label{eqn:sysb}
+\end{align}
 
-Finally, to take into account the seasonality of the monthly data, random effects for each of the 12 months are estimated with a $Cauchy(0,4)$ prior on each parameter.
-The remaining hyper-parameters are given Cauchy or half-Cauchy priors.
+The $\mu$ terms are all in term modelled as coming from a Cauchy distribution with a location meta-parameter $\mu^{meta}$ and scale parameter $\sigma^{mu}$, which are also estimated by the model.
+These meta-parameters are in turn given Cauchy prior distributions with location 0 and scale 5.
+
+The $cauchy(0,5)$ prior is what is referred to as a weakly-informative prior.
+It provides some information to the MCMC algorithm for where to explore for probability mass, but does not impose any strong prior knowledge that meaningfully effects the posterior distribution.
+The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional large coefficient in a lower level regression, but still provides some shrinkage towards zero.
+\citep{gelman_weakly_2008} The coefficients $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{month}$ are all also given $Cauchy(0,5)$ prior distributions.
+
+\begin{align}
+\mu^{lease} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ % \label{eqn:meta1}
+\mu^{sect} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ %\label{eqn:meta2}
+\mu^{manuf} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ %\label{eqn:meta3}
+\mu^{meta} &\sim Cauchy(0, 5) \\ %\label{eqn:meta4}
+\sigma^{mu} &\sim half-Cauchy(0,5) %\label{eqn:meta5}
+\end{align}
+
+\subsection{Computation and Results from Bayesian multilevel model}
+
+The model is estimated using the Stan probabilistic programming language and sampler, which implements Hamiltonian MCMC and a No U-turn sampler (NUTS) \citep{stan_development_team_stan_2014}.
+The sampler was run with four chains and 1000 iterations.
+R-statistics of 1 indicate convergence in the model.
+2000 samples are taken from the estimated posterior distribution, and model results for the marginal distributions of the parameters of interest are presented in the form of histograms of the samples.
+
+Recall, the testable implications from the theory of asymmetric information and quality were that 1.) average quality should be higher in the panels owned by the high-information type (leased panels).
+2.) Price should be more highly correlated with quality in the panels owned by the high-information type.
+
+The parameters of interest in our model are then the higher level parameters $\mu_{lease}$ and $\beta$.
+The first represents the mean value of the system level slope terms, $b$ grouped by whether the systems were leased or host-owned.
+The latter is a measure of the correlation between the cost of the system and the system slope terms.
+The $\beta$ parameter is also allowed to vary by leased or host-owned.
+It is important to note, that the estimated $\beta$ paramaters are estimated conditional on the installation date of the solar panels, thus they do not reflect changes in price over time.
+This could lead to a heavy bias in the results if, as seems plausible, both prices and quality changed over time.
+
+For the testable implications, the inference we are interested is in the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta{owned}$.
+Histograms representing the marginal distributions of these contrasts are presented in figure \ref{fig:contrasts}.
+The top panel shows the distribution of the contrast $\mu_{lease, yes} - \mu_{lease, no}$.
+Essentially all of the probability mass is located in the positive range.
+This reinforces the finding from the mixed effects model that leased panels due appear to have substantially higher quality, as the theory of asymmetric information of quality would suggest.
+
+The evidence for the second testable implication is less conclusive.
+The model estimates that there is approximately a 89\% probability that there is a higher correlation between quality and cost for the high information type owner of leased panels compared to the low information type owner of host-owned panels.
 
 \begin{figure}
-	\includegraphics[width=1\textwidth]{figures/solar_prod_bayes_diag.png}
-	\caption{The data and the research question combined suggest a hierarchical Bayesian model.
-The figure shows the structure of the model.
-Production data is grouped by individual solar panel systems (index i).}
-	\label{solar_prod_bayes_diag}
+  \includegraphics[width=1\linewidth]{figures/bayes_hypos.png}
+  \caption{Marginal posterior distributions for the contrasts of interest.
+The top panel shows that the distributions of the slope parameters, $b_s$, are estimated to have means that are significantly different from each other between leased and not leased systems as defined as the contrast, $\mu_{lease, yes} - \mu_{lease, no}$.
+The lower panel shows the marginal posterior distribution of the contrast between the $\beta$ parameter for leased and not-leased solar systems.
+The results indicate that there is an approximately 89\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned.}
+  \label{fig:contrasts}
 \end{figure}
 
-To estimate the parameters of the model, I use the Stan Bayesian programming language and simulator \citep{stan_development_team_stan_2014} which uses Hamiltonian MCMC to estimate a joint posterior distribution of the parameters.
-\citet{kruschke_doing_2014} provides an accessible explanation of the Hamiltonian MCMC algorithm, while more detailed technical descriptions are provided by \citet{gelman_bayesian_2013} and \citet{stan_development_team_stan_2014}.
+A direct comparison of prices and costs between these groups is however complicated.
+The prices of host-owned panels are market prices: simply the purchase price that the owner of the solar panel system paid to the contractor.
+On the other hand, the price of leased systems is a reported price from the contractors consisting of the component and installation costs plus an estimated mark-up.
+If these costs were reported correctly, then the theory should still hold.
+The reported cost should reflect the ability of the high-information owners to judge quality of the panels that they purchase.
+In practice, the reliability of reported cost is a source of added uncertainty, and should be taken into consideration when interpreting the results.
 
-Bayesian simulation methods are still emerging - especially in Economics.
-So it is worth spending a moment to motivate their use in this case over more common asymptotic methods.
-As mentioned, the main reason for using the Bayesian model is the flexible treatment it allows for modeling the inherent hierarchy of the data and allowing for partial pooling in estimation.
+When that is said, if we assume that misreporting of the true costs will tend to lead to less variation in reported costs, then that should have the affect of attenuating the magnitude of the contrast, $\beta_{cost, l=yes} - \beta_{cost, l=no}$.
+In that case, the estimated contrast can be seen as being a conservative estimate.
 
-The usefulness of Bayesian models is most apparent in the estimation of complex hierarchical models with many parameters.
-In a hierarchical model, lower-level parameters within a certain group are themselves modeled as coming from a distribution characterized by ``meta-parameters''.
-Thus, the estimated group-level parameters are a weighted function of both the observations within the group as well as the full set of observations in the dataset.
-For example, the estimated mean parameter on a group with a few, outlying observations would be pulled towards the full sample mean.
-This partial pooling also serves as a natural form of parameter shrinkage, which mostly eliminates the need for using corrections for multiple comparisons in inference \citep{gelman_data_2006}.
+Other higher level parameters are reported in table \ref{table:higher_level_params}.
+The sector groups can be of interest in a discussion of the role of asymmetric information.
+Figure \ref{mu_sectors} shows the distributions of the contrasts relative to the residential host sector.
+The top panel represents the contrast with commercial sector hosts, the middle with government sector host and the bottom panel with non-profit sector host.
+All three contrasts indicate relatively lower quality for installations for these sectors compared to the residential sector.
 
-With Bayesian multilevel model, I also have greater flexibility in specifying the model without having to rely on assumptions like constant group-level variances and a Gaussian noise distribution.
-Because Bayesian simulation techniques result in an estimate of the full joint probability distribution of the model, the inference has the potential to be more informative than the typical point estimates and p-values of of the standard hypothesis testing frameworks \citep{kruschke_doing_2014}.
+While it is not clear why this should be, an explanation consistent with the theory of asymmetric information of quality, is that there exists an agency problem.
+A homeowner will, in some form, have a direct financial incentive to acquire solar panels of good quality.
+On the other hand, A manager or employee of a firm, government agency or non-profit may not have any direct financial incentive in ensuring that a solar panel system is of high quality.
 
-Another advantage is that the posterior distributions of the parameters have a natural and direct interpretation as probabilities, as opposed to standard hypothesis testing where the concepts of p-values and significance are often ill-defined.
-See \citet{kruschke_doing_2014} or \citet{gelman_bayesian_2013} for further discussion.
+\begin{table}
+  \begin{tabular}{lrrrrrrrrr}
+  \toprule
+  {} &    mean &    std &     min &    2.5\% &     25\% &     50\% &     75\% &   97.5\% &     max \\
+  \midrule
+  $\mu^{meta}$     &  0.004' & 0.022' & -0.018' & -0.010' & -0.007' & -0.005' & -0.002' &  0.069' &  0.084' \\
+  $\beta^{meta}$   &  0.003' & 0.838' & -8.120' & -1.707' & -0.074' & -0.007' &  0.075' &  1.738' &  6.787' \\
+  $\mu^{meta,mon}$ &  0.005' & 0.036' & -0.138' & -0.065' & -0.019' &  0.006' &  0.030' &  0.075' &  0.146' \\
+  $\mu^{b_s}$      & -0.000' & 0.015' & -0.043' & -0.028' & -0.012' & -0.002' &  0.011' &  0.031' &  0.044' \\
+  $\mu^{jan}$      & -0.164' & 0.013' & -0.188' & -0.186' & -0.173' & -0.163' & -0.154' & -0.144' & -0.142' \\
+  $\mu^{feb}$      & -0.121' & 0.013' & -0.146' & -0.143' & -0.131' & -0.120' & -0.111' & -0.102' & -0.100' \\
+  $\mu^{mar}$      & -0.054' & 0.013' & -0.078' & -0.076' & -0.064' & -0.053' & -0.044' & -0.035' & -0.033' \\
+  $\mu^{apr}$     &  0.040' & 0.013' &  0.017' &  0.018' &  0.031' &  0.042' &  0.051' &  0.060' &  0.062' \\
+  $\mu^{may}$      &  0.087' & 0.013' &  0.063' &  0.064' &  0.077' &  0.088' &  0.097' &  0.106' &  0.108' \\
+  $\mu^{jun}$      &  0.111' & 0.013' &  0.087' &  0.089' &  0.102' &  0.113' &  0.121' &  0.131' &  0.133' \\
+  $\mu^{jul}$      &  0.108' & 0.013' &  0.084' &  0.086' &  0.099' &  0.110' &  0.118' &  0.128' &  0.130' \\
+  $\mu^{aug}$      &  0.101' & 0.013' &  0.077' &  0.079' &  0.092' &  0.103' &  0.112' &  0.121' &  0.123' \\
+  $\mu^{sep}$     &  0.082' & 0.013' &  0.058' &  0.060' &  0.072' &  0.083' &  0.092' &  0.102' &  0.103' \\
+  $\mu^{oct}$      &  0.026' & 0.013' &  0.002' &  0.004' &  0.016' &  0.027' &  0.036' &  0.045' &  0.047' \\
+  $\mu^{nov}$     & -0.031' & 0.013' & -0.056' & -0.053' & -0.041' & -0.030' & -0.021' & -0.011' & -0.010' \\
+  $\mu^{dec}$     & -0.123' & 0.013' & -0.147' & -0.146' & -0.133' & -0.122' & -0.113' & -0.104' & -0.101' \\
+  $\mu^{lease, no}$   & -0.030' & 0.045' & -0.193' & -0.159' & -0.015' & -0.011' & -0.008' & -0.002' &  0.015' \\
+  $\mu^{lease, yes}$   & -0.022' & 0.045' & -0.186' & -0.151' & -0.007' & -0.003' &  0.000' &  0.006' &  0.023' \\
+  $\mu^{sect, com}$      &  0.002' & 0.022' & -0.024' & -0.015' & -0.010' & -0.007' & -0.004' &  0.066' &  0.087' \\
+  $\mu^{sect, res}$       &  0.012' & 0.022' & -0.014' & -0.005' & -0.000' &  0.003' &  0.006' &  0.075' &  0.096' \\
+  $\mu^{sect, gov}$       &  0.006' & 0.022' & -0.017' & -0.010' & -0.005' & -0.003' &  0.001' &  0.071' &  0.092' \\
+  $\mu^{sect, npr}$      & -0.001' & 0.022' & -0.026' & -0.018' & -0.012' & -0.010' & -0.006' &  0.063' &  0.084' \\
+  $\beta^{cost, l=no}$  & -0.002' & 0.002' & -0.007' & -0.005' & -0.003' & -0.002' & -0.001' &  0.001' &  0.004' \\
+  $\beta^{cost, l=yes}$  &  0.001' & 0.001' & -0.002' & -0.001' &  0.000' &  0.001' &  0.001' &  0.002' &  0.003' \\
+  $\beta^{size}$  &  0.002' & 0.001' & -0.001' &  0.000' &  0.002' &  0.002' &  0.003' &  0.004' &  0.006' \\
+  $\beta^{fy}$    & -0.018' & 0.001' & -0.021' & -0.020' & -0.018' & -0.018' & -0.017' & -0.016' & -0.014' \\
+  $\sigma$      &  0.065' & 0.000' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  0.066' &  0.066' \\
+  $\sigma^{b_s}$    &  0.476' & 0.006' &  0.454' &  0.465' &  0.472' &  0.476' &  0.480' &  0.488' &  0.495' \\
+  $\sigma^{mon}$   &  0.114' & 0.030' &  0.059' &  0.074' &  0.094' &  0.108' &  0.128' &  0.185' &  0.341' \\
+  $\sigma^{\beta}$  &  0.833' & 1.773' &  0.001' &  0.006' &  0.047' &  0.195' &  0.836' &  5.675' & 19.879' \\
+  $\sigma^{mu}$   &  0.006' & 0.001' &  0.003' &  0.004' &  0.005' &  0.006' &  0.006' &  0.008' &  0.011' \\
+  \bottomrule
+  \end{tabular}
+\label{table:higher_level_params}
+\caption{Summary statistics of the estimated posterior distributions of higher level variables}
+\end{table}
+
+\begin{figure}
+  \includegraphics[width=1\linewidth]{figures/mu_sectors.png}
+  \caption{Distribution of the sector contrasts relative to residential sector.
+Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
+  \label{mu_sectors}
+\end{figure}
+
+Summary statistics for the parameters on the manufacturer group means as well as the lower-level intercept and slope paramaters $a_s$ and $b_s$ are too numerous to present in a table.
+Visual summaries are presented in figures \ref{sfig:BayManPlot}, \ref{sfig:Bay_as} and \ref{sfig:Bay_bs}.
+Figure \ref{sfig:BayManPlot} shows box plots of the manufacturer distributions, with the box representing 70 percent of probability and the lines representing 95 percent of the probability.
+Mirroring earlier results, the plot shows most manufacturers with roughly equivalent quality, with handful with meaningfully lower quality.
+Figures \ref{sfig:Bay_as} and \ref{sfig:Bay_bs} show 95\% intervals of the distribution of the $a_s$ and $b_s$ parameters.
+The intercept terms are shown to have considerable variation between systems, as would be expected, but are individually estimated precisely.
+
+The slope parameters, $b_s$ are modelled explicitly and are also allowed to have varying variance parameters.
+This is evident in the plot.
+A similar pattern to the higher-level manufacturer group means is apparent: modest differences in quality are seen between most systems, with the exception of a lower tail of systems with particularly poor quality.
+
+\begin{figure}[!htb]
+\begin{minipage}{.48\textwidth}
+  \includegraphics[width=1\linewidth]{figures/BayManPlot.png}
+  \caption{Box plot of the the parameters $\mu^{manuf}$, representing the distributions of the mean parameter for manufacturer groupings.}
+  \label{sfig:BayManPlot}
+\end{minipage}\hfill
+\begin{minipage}{.48\textwidth}
+  \includegraphics[width=1\linewidth]{figures/Bay_as.png}
+  \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$.}
+  \label{sfig:Bay_as}
+\end{minipage}\hfill
+\begin{minipage}{.48\textwidth}
+  \includegraphics[width=1\linewidth]{figures/Bay_bs.png}
+  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
+  \label{sfig:Bay_bs}
+\end{minipage}\hfill
+%\caption{Simulated distribution of coefficients}
+%\label{fig:coeff_sim}
+\end{figure}
 
@@
 The theoretical literature on information asymmetry and quality is established and deep.
-The empirical literature for consumer durable investments is, however, more sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability.
+The empirical literature for durable investments is, however, more sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability.
 This article provides one of few case studies that both identifies a market that could be expected to have issues of information asymmetry of quality, and provides an indirect empirical test for the presence of information asymmetry of quality.

### 2018-03-22 stan_analysis_new.py

@@ -93,8 +93,10 @@ def create_stan_data(solar_data):
 
     #sector categories
     owner_sect_dict = create_num_dict(system_level_df["owner_sector"])
+    print("owner_sect_dict", owner_sect_dict)
     #for varying effects
     owner_sect_id = system_level_df["owner_sector"].map(owner_sect_dict)
+
     #for dummy variables
     owner_sect_dummies = pd.get_dummies(system_level_df.owner_sector)
     S=4
@@ -421,7 +423,6 @@ def solar_stan_model2():
     """
     return(solar_stan_model1)
 
-
 #run:
 solar_data = load_data()
 solar_data = format_data(solar_data)
@@ -433,6 +434,8 @@ np.std(solar_data["months_operation"])
 
 
 solar_data[["log_prod_kwh", "months_operation", "month"]].apply(np.mean, axis=0)
+
+solar_data.head()
 
 solar_stan_data = create_stan_data(solar_data);
 

### 2018-03-27 solar_commodities.tex
first draft

@@
  Solar panels should not be considered commodities.
-Considerable quality differences, as measured by degradation of production over time, are found between manufacturers.
+Considerable quality differences, as measured directly by degradation of production over time, are found between manufacturers.
 This has implications for pricing and competition in the market for solar panel systems.
-I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with third party owners display higher quality than those owned directly.
-2.) Prices of solar panels are less correlated to quality for solar panels that are owned directly.
-I use multilevel models estimated by both maximum likelihood and Bayesian Markov Chain Monte-Carlo.
+I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts.
+2.) Furthermore, prices of solar panels are more correlated to quality for solar panels that are owned by high-information owners.
+I use linear mixed effects models estimated by maximum likelihood and multilevel models estimated by Bayesian Markov Chain Monte-Carlo.
 \end{abstract}
@@
 
-Solar panels have been referred to colloquially in the power industry as being commodities.
+Solar panels have been referred to colloquially as being commodities.
 This is in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade.
@@
 
-However, a more thorough definition of a commodity generally includes a certain degree of fungibility.
+More formally, a commodity is generally defined by a certain degree of fungibility.
 That is, the good from a certain producer will be largely interchangeable with the good of another producer.
@@
 
-However, an important differentiating characteristic between solar panels is the quality of these solar panels.
+An important differentiating characteristic between solar panels is the quality of these solar panels.
 This can be measured as both the failure rate of these panels as well as the degree of degradation of the panels over time.
-This can imply that certain manufacturers with superior quality could extract a premium in the market.
+If there were significant differences in quality between manufacturers, then this has implications for the structure and development of the solar panel market.
+Manufacturers with superior quality could differentiate themselves and extract a premium in the market.
+A reputation for quality among established manufacturers could also act as a barrier to entry, thereby reducing long-term pricing pressure.
 
-The combination of quality differences between panel manufacturers and the distributed nature of solar panel generation also opens up the issue of asymmetric information in the market for solar panels.
+Furthermore, if significant quality differences between manufacturers exists, then the characteristics of solar panel investments suggest the possibility of asymmetric information of quality in the market.
 
-In an earlier paper \citep{mauritzen_whats_2015} I show that the boom in rooftop solar panel investment in California coincided with an adoption of a leasing business model by contractors.
+%In an earlier paper \citep{mauritzen_cost_2017} I show that the boom in rooftop solar panel investment in California coincided with an adoption of a leasing business model by contractors.
 With leased panels, homeowners and small businesses do not pay upfront or own the panels on their roof.
@@
 
-Leased panels could be attractive for homeowners and businesses for several reasons.
-Not having to invest anything upfront would be attractive to those that are cash constrained.
-Lack of ownership could also be attractive to those concerned about the practicalities of maintaining and operating a small power station.
-
-In addition, I argue that an issue of information asymmetry of quality could exist in the purchase of rooftop solar panel systems, and that leasing could provide a mechanism for alleviating this asymmetry.
+%In addition, I argue that an issue of information asymmetry of quality could exist in the purchase of rooftop solar panel systems, and that leasing could provide a mechanism for alleviating this asymmetry.
 Homeowners and small business owners can not generally be expected to have the expertise to judge the quality of panels on their own.
@@
 
-Generally, the market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality.
-Solar panels can be characterized as an ``experience'' good, where an investor needs to learn about the quality through use.
+The market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality.
+Solar panels can be characterised as an ``experience'' good, where an investor/consumer needs to learn about the quality through use.
 In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels.
@@
 
-Given the difficulty of judging quality and lack of market mechanisms to signal quality, the established economic theory on the subject would suggest that the market may tend to provide low-quality panels \citet{tirole_theory_1988}.
+Given the difficulty of judging quality and lack of market mechanisms to signal quality, the established economic theory on the subject would suggest that the market may tend to provide low-quality panels at relatively undifferentiated prices \citep{tirole_theory_1988}.
 
-However, over the period from 2010 to 2014, many of the larger solar contractors moved to leasing solar power systems to homeowners.
-The introduction of a leasing model could potentially get over this information asymmetry problem by shifting ownership to the large contractors that install and finance the solar systems.
+Over the period from 2010 to 2016, many of the larger solar contractors moved to leasing solar power systems to homeowners.
+The introduction of a leasing model could potentially get over this information asymmetry problem by shifting ownership to the large contractors that install and finance the solar systems \citep{mauritzen_cost_2017}.
 These contractors can then in turn take steps such as testing panels and visiting manufacturing sites to ensure quality of suppliers.
+In other words, a descriptive understanding of the market suggests that it is possible to divide the market into two types of solar panel owners: high-information owners who lease the solar panel systems to individual home owners, businesses and other organisations, and low-information owners who buy and own the solar panel systems directly.
 
-A testable implication that emerges is then that the quality of panels - as measured by degradation of output over time - is better in systems that are leased.
-In this paper, I use a data set of California solar power systems that includes monthly production data.
-I measure the average degradation of production over time as a proxy for the inherent quality of solar panel system.
+From this informal model of information asymmetry in the market for solar panels, a testable implication arises.
+The quality of panels, as measured by degradation of output over time, would be predicted to be higher in the systems owned by the high-information owners (third-party-owner) who can discern between low and high quality solar panels.
+Conversely, the basic theory of asymmetric information suggests that lower quality panels will tend to push out higher quality panels in the presence of low-information buyers who can not discern quality, and therefor make purchasing decision solely on price.
 
-Testing for the degradation in production of more than 3000 panels is however problematic using traditional econometric models based on frequentist approaches.
-The production patterns of the solar panels could be expected to vary greatly depending on location, equipment, size as well as other factors.
-The inclusion of many fixed effects to control for these factors would likely lead to over-fitting.
-That is, the model may have a good fit to the data, but will tend to have poor out-of-sample predictive power \citep{gelman_bayesian_2013}.
+A related prediction from the theory of asymmetric information of quality is that the cost of solar panel systems should have a higher degree of correlation with quality when the owner is a high-information type.
+Because a high-information owner can, by definition, differentiate between high and low quality panels.
 
-As a solution, I use a Bayesian hierarchical model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
-The main benefit of using a Bayesian multilevel model is that we can fit a complex model with many hundreds of parameters, while avoiding problems of over-fitting and issues of multiple comparison in inference by relying on partial pooling of the parameters.
-The main finding is that solar panel systems that were leased tend to show significantly less degradation over time than those that were sold outright.
+In this paper, I use a data set of California solar power systems that includes monthly production data from over 3000 solar panel systems.
+Where many studies of asymmetric information of quality must use proxies for quality, I measure it directly as the average degradation of production over time.
+
+In summary, I have three nested hypothesis that I wish to test:
+
+\begin{enumerate}
+  \item Solar panels are not pure commodities. Significant quality differences can be detected between manufacturers of panels.
+  \item High information owners (third-party-owners) will purchase higher quality panels than low-information owners (host-owners)
+  \item Price and quality will be more highly correlated in solar panel systems owned by high information types.
+\end{enumerate}
+
+The data has a natural hierarchy of between 36 and 60 monthly observations within the 3000-plus solar panel systems, which in turn can be grouped according to relevant characteristics such as whether they are host-owned or third-party owned.
+Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model them directly.
+For relatively parsimonious models with one or two hierarchies and few covariates, I can use linear mixed effects models (LMEM) estimated by maximum likelihood.
+However, for richer models  with all relevant hierarchies, maximum likelihood becomes impractical.
+
+As a solution, I use a Bayesian multilevel\footnote{Sometimes also called hierarchical} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
+The main benefit of using a Bayesian multilevel model is that I can fit a complex model with many hundreds of parameters, while avoiding problems of over-fitting and issues of multiple comparison in inference by relying on regularisation of parameters through the use of higher level parameters.
 
@@
 
-Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller - in other words the seller can choose the level of quality of its product.
-Under these assumptions, the authors predict several equilibrium - essentially markets for both high and low quality products depending on the extent to which the buyer is informed.
+Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller -- in other words the seller can choose the level of quality of its product.
+Under these assumptions, the authors predict several equilibrium -- essentially markets for both high and low quality products depending on the extent to which the buyer is informed.
 
-Reputation - or lack thereof - plays a particularly important role in this market.
-Most of both the solar installers and panel manufacturers are relatively new firms - especially the increasingly dominant Chinese solar panel producers,  \citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.
+Reputation, or lack thereof, plays a particularly important role in this market.
+Most of both the solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers.
+\citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.
 
-Expanding empirical literatures also exist on asymmetric information of quality - especially in the marketing, finance and accounting fields.
+An expanding empirical literatures exist on asymmetric information of quality, especially in the marketing, finance and accounting fields.
 In accounting, the issue of firm quality and voluntary disclosure in regulatory filings as a signal has received broad attention by empirical researchers.
 \citet{healy_information_2001} review the early literature.
-Empirical studies in finance have, for example, looked at issues of adverse-selection in Initial Public Offerings \citep{michaely_pricing_1994} and small business lending \citep{petersen_benefits_1994} among many other topics.
+Empirical studies in finance have, for example, looked at issues of adverse-selection and quality in Initial Public Offerings \citep{michaely_pricing_1994} and small business lending \citep{petersen_benefits_1994} among many other topics.
 
 The literature in the marketing field may be more relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables.
-The effects of signaling on quality - usually in the form of advertising - is particularly important in the field.
+The effects of signalling on quality, usually in the form of advertising, is particularly important in the field.
 \citet{kirmani_no_2000} provide an overview of this literature.
 
-The use of leasing, discussed in this article, can be seen as a form of costly signaling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of poor quality to the contractor.
-
-Traditionally, energy generation has been an activity undertaken by large, specialized firms with specialized knowledge and resources.
+The use of leasing, discussed in this article, can be seen as a form of costly signalling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of poor quality to the contractor.
+Traditionally, energy generation has been an activity undertaken by large, specialised firms with specialised knowledge and resources.
 The role of information asymmetry in investment and purchase decisions has likely been seen as a secondary issue, and to my knowledge, no literature exists on the role of such informational issues in the industry.
 
-However, the growth of distributed energy technologies, like solar panels, has added new focus on behavioral and information issues related to energy investments by homeowners and small businesses.
-In turn, a growing literature is forming around solar power investment behavior of homeowners and small businesses.
+However, the growth of distributed energy technologies, like solar panels, has added new focus on informational and behavioural issues related to energy investments by homeowners and small businesses.
+In turn, a growing literature is forming around solar power investment behaviour of homeowners and small businesses.
 
@@
 
-Despite the growth of the literature on investments in distributed energy, to my knowledge this is the first paper to identify and directly test implications of information asymmetry in the expanding industry.
-The paper has implications for regulation and subsidy policy for the expanding distributed energy industry.
+Despite the growth of the literature on investments in distributed energy, to my knowledge this is the first paper to identify and directly test for quality differences and related role of information asymmetry in the expanding market.
+The article has implications for an understanding of the market for distributed energy technologies, subsidy policies and regulation of the market.
 
@@
 
-The data used in this paper is an intersection of two datasets from the California Solar Initiative (CSI).
+The data used in this article is an intersection of two datasets from the California Solar Initiative (CSI).
 The California Solar Initiative is a state-wide program that gives incentives for installing grid-connected solar panels systems.
 The incentives are based on performance.
-For smaller systems, typical of most residential and small business installations, this incentive was given in the form of an upfront payout based on the capacity and the expected performance of the system.
+For smaller systems, typical of most residential and small business installations, this incentive could be given in the form of an upfront payout based on the capacity and the expected performance of the system.
 
@@
 
-Below are the key variables present in the CSI installation data:
-
-\begin{itemize}
-\item Installation date
-\item Location: address, zip code, county
-\item System capacity
-\item Name of system owner, host, and contractor
-\item Panel and inverter manufacturer
-\item Third party owner (leased)
-\item Use of single or double axis trackers
-\item Year of installation
-\end{itemize}
-
 The installation data can be matched with the production data for those installations receiving a production subsidy.
@@
 
-Figure \ref{tot_production} shows the production for the included solar panel systems over time.
-The seasonality of the systems over the calendar year is clear, and needs to be accounted for in the statistical analysis.
+Table \ref{tbl:variables} gives an overview of the variables in the dataset that are used in the analysis.
+Notice that variables can vary by monthly observation or by solar power system.
+I will take account of this natural data hierarchy in my modelling.
+There are a total of 171,027 monthly observations used in the analysis.
+These observations come from 3,145 unique solar power systems.
+From these solar panel systems, panels from a total of 107 unique manufacturers are used.
+
+\begin{table}
+  \begin{tabular}{rllll}
+  \toprule
+    Variable &  Label &   Level & Values &   Mean Value  \\
+    \midrule
+    Production (kWh) &  prod\_kwh &  Observation & Real & 35,908   \\
+    Months of operation &  months\_operation &  Observation & Integer & --   \\
+    Month of year & month & Observeration & Categ.$^{1}$& -- \\
+    Cost per kw & cost\_per\_kw & System & Real & 7,026 \\
+    Capacity rating (kW) & csi\_rating & System & Real & 239 \\
+    First year of production & first\_prod\_year & System & Integer & 2011 \\
+    Third party owned & lease & System & Categ.$^{2}$  &  35\% leased \\
+    Owner sector & own\_sector\_id & System & Categ.$^{3}$& -- \\
+    \midrule
+    \# Observations && 171,027 && \\
+    \# Solar panel systems && 3,145 && \\
+    \# Module manufacturers && 107 && \\
+     Min. months of data && 36 && \\
+     Max. months of data && 60 && \\
+  \bottomrule
+    \multicolumn{5}{l}{\scriptsize{$^{1}$January - December. $^{2}$Leased 1, Owned 0.$^{3}$ Residential (497), Commercial (1759), Governmental (777), Non-profit (112).}}
+  \end{tabular}
+  \label{tbl:variables}
+  \caption{}
+\end{table}
+
+Production is measured in kilowatt hours (kWh).
+In the analysis, production data is first log transformed and then descaled, as shown in equation \ref{eqn:descaled}.
+For production data for each system, $s$, the log production data is demeaned and then divided by 2 times the standard deviation of the production data for that solar power system.
+Figure \ref{fig:production} gives a visualisation of the transformed production data.
+The red line shows the average degradation over time of all the systems.
+The strong seasonality of production is also clearly visible.
+
+Other continuous variables are transformed in a similar manner, with observation level variables descaled at the system level, and system-level variables descaled across systems.
+These transformations avoid confusion that may come when interpreting coefficients on variables with different units and scales.
+There are also technical reasons for the transformations relating to the MCMC routine.
+Conversion in the MCMC algorithm is aided when all variables are on the same scale.
+Setting appropriate weakly informative prior distributions on the parameters is also simplified when continuous variables are free from units.
+
+\begin{equation}
+prod\_scaled_{s[i]} = \frac{log(prod\_kwh_{s[i])} - mean(log(prod\_kwh_{s[i]}))}{2 std(log(prod\_kwh_{s[i]}))}
+\label{eqn:descaled}
+\end{equation}
 
 \begin{figure}
-	\includegraphics[width=1\textwidth]{tot_production.png}
-	\caption{Production over time from California solar panels, Host owned and leased.
-Production index reflects monthly production in kWh normalized by capacity multiplied by 30.5*12 to reflect a rough estimate of maximum day-light hours in a month.}
-	\label{tot_production}
+  \includegraphics[width=1\linewidth]{figures/production.png}
+  \caption{The figure shows the de-scaled production over time from the sample of solar panel systems. The red line represents the average degradation over time for all panels.}
+  \label{fig:production}
 \end{figure}
 
-The main methodological goal of this paper is to make an unbiased estimate of the slope of the average degradation of solar panels - comparing the slopes between groups of systems that are leased and bought outright.
-These estimates should also control for seasonality and idiosyncratic factors between solar power systems.
-A simple example of fitting a straight lines through two production series is shown in figure \ref{example_prod} - here as a simple single variable OLS estimate.
+Month of year dummy variables are used to control for the seasonality in the data.
+The capacity of the solar panel system is the actual maximum capacity of a solar power system as reported to and audited by the CSI.
+This can be somewhat lower than the total nameplate capacity of the solar panels, and can depend on not only the technical capacity of the panels, but also on supporting equipment, such as the inverter, and factors such as the angle of the roof and orientation of the solar panel system.
 
-\begin{figure}
-	\includegraphics[width=1\textwidth]{example_prod.png}
-	\caption{The main methodological task in this paper is to make an unbiased estimate of the slope of the average degradation of production over time.
-The figure demonstrates a straight line fitted to a time series from a single solar power system.}
-	\label{example_prod}
-\end{figure}
+The cost per kilowatt (\$/kW) variable is the reported total cost of the solar panel system divided by the CSI capacity rating.
+For solar panel systems that are sold outright, the total cost is simply the transaction price of the entire system.
+For systems with third party owners, the total cost is an estimate reported by the system owner based on the sum of component costs, installation costs and a mark-up.
+I will discuss how this may effect the results.
 
-The question of interest and the structure of the available data suggests a hierarchical structure to the empirical model.
-The individual production data are grouped by the different solar panel systems.
-Each system is then in turn grouped into categories of leased or host-owned.
-As mentioned, seasonality must also be adequately accounted for as well as other factors such as geographic location, use of tracking and idiosyncratic variation from one solar panel system to another.
+In the analysis I distinguish between solar panels that are host-owned and those that are third-party owned, which I refer to as being leased.
+What I refer to as leased can however take two main forms.
+The most common is a Power Purchase Agreement (PPA) where the host agrees to buy the electricity from the solar panel system for an agreed upon price, (which may increase at an agreed upon rate over time).
+Alternatively, a host may pay a flat monthly fee for the electricity produced by the solar panel.
+
+Finally, among the 3000 plus solar power systems, 497 had residential owners, 1759 had commercial owners, 777 had governmental owners and 112 had non-profit owners.
+A portion of the commercial owners operate as third-party-owners, with hosts in the other sectors.
+Alternatively, I could include host sector in my analysis: 1241 solar panel systems had a commercial host, 501 had a residential host, 1199 had a governmental host and 204 had a non-profit host.
+However, In a study of asymmetric information, it makes most sense to focus on the sector of the owner.
 
@@
 Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or "fixed effects") of the $n times p$ matrix of (p-1) explanatory variables $\mathbf{X}$ plus intercept term.
-$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or "random effects" $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of "random" refers to the fact that these groups are assumed to be specific to the sample.
+$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or "random effects" $\mathbf{B}$ that are fixed at values $\mathbf{b}$.
+\footnote{The use of "random" refers to the fact that these groups are assumed to be specific to the sample.
 Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms, thus the "random" label.
@@
 \begin{align}
-(y|\mathbf{B=b} &\sim \mathcal{N}(\mathbf{X\beta + Zb}, \sigma^2\mathbf{W^{-1}}))\\
+(y|\mathbf{B=b}) &\sim \mathcal{N}(\mathbf{X\beta + Zb}, \sigma^2\mathbf{W^{-1}})\\
 B &\sim \mathcal{N}(\mathbf{0,\Sigma})
@@
 
-Models of this form can be solved efficiently using a penalized maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.
+Models of this form can be solved efficiently using a penalised maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.
 
 To test whether I can detect quality differences between module manufacturers, I need to first estimate a restricted comparison model.
-I write the restricted comparison model explicitly in equations \ref{eqn:observation_level_rest} and \ref{eqn:system_level_rest}.
+I write the restricted comparison model explicitly in equation \ref{eqn:observation_level_rest}.
 
@@
 
-\begin{align}
+\begin{equation}
+\begin{aligned}
 prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
-\label{eqn:observation_level_rest}
 \begin{pmatrix}
@@
   \rho \sigma_a \sigma_b & \sigma_b^2
-\end{pmatrix}
-%\label{eqn:system_level_rest}
-\end{align}
+\end{pmatrix}\label{eqn:observation_level_rest}
+\end{aligned}
+\end{equation}
 
@@
 The system-level intercepts, $a$, and slope, $b$, terms are also modelled as a multivariate normal.
-But the mean values of the multivariate normal distribution, $mu_a$ and $mu_b$, are now indexed by $m$, representing that they are allowed to vary by module manufacturer.
-\begin{align}
+But the mean values of the multivariate normal distribution, $\mu_a$ and $\mu_b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.
+
+\begin{equation}
+\begin{aligned}
 \begin{pmatrix}
@@
 \end{pmatrix} \\
-\mu_{a} &\sim N(\theta_a, \sigma_{mu_a}^2) \\
-\mu_{b} & \sim N(\theta_b), \sigma_{mu_b}^2)
-\end{align}
+\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
+\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:unrest}
+\end{aligned}
+\end{equation}
 
-If there is no significant variation between the module manufacturer groupings, this model collapses into the restricted comparison model, which is the null hypothosis.
+If there is no significant variation between the module manufacturer groupings, this model collapses into the restricted comparison model, which is the null hypothesis.
 Figure \ref{sfig:sys_slope_fig} gives a visual summary of the system-level slope terms ordered from lowest to highest.
@@
   \toprule
-    Model &  DF &   AIC &   BIC &  logLik &  deviance &  Chisq &  Chi df &  Pr(>Chisq) \\
+    Model &  DF &   AIC &   BIC &  logLik &  deviance &  Chisq &  Chi df &  $Pr(>Chisq)$ \\
     \midrule
@@
   \label{tbl:lm_commodity}
+  \caption{A likelihood ratio test indicates that the addition of manufacterer as a explanatory variable significantly improves the predictive power of the model}
 \end{table}
@@
 We can then compare the average slopes between these groups.
-For convenience the model is presented in full in equations \ref{eqn:observation_level_rest2}.
+For convenience the model is presented in full in equations \ref{eqn:fullmodel}.
 
-\begin{align}
+\begin{equation}
+\begin{aligned}
 prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
-\label{eqn:observation_level_rest2}
 \begin{pmatrix}
@@
 \end{pmatrix} \\
-\mu_{a} &\sim N(\theta_a, \sigma_{mu_a}^2) \\
-\mu_{b} & \sim N(\theta_b), \sigma_{mu_b}^2)
-\label{eqn:lease_model}
-\end{align}
+\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
+\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:fullmodel}
+\end{aligned}
+\end{equation}
 
@@
 \begin{figure}
+  \centering
 	\includegraphics[width=.8\textwidth]{figures/lease_sys_fig.png}
@@
 First and foremost so that the parameters of interests to the testable implications can be estimated from one model.
-But also, \citet{bar_random_2013} shows, using Monte-Carlo simulations, that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
+But also, \citet{barr_random_2013} shows, using Monte-Carlo simulations, that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
 Estimating mixed effects model by maximum likelihood is efficient and straight-forward.
@@
 Group-level variances are not constrained to be equal and assumptions of normality of the likelihood and posterior distributions can be loosened.
-For further discussion of the Bayesian frameworks and techniques, I refer to the excellent discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2017}.
+For further discussion of the Bayesian frameworks and techniques, I refer to the excellent discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2015}.
 
@@
 
-Equations \ref{eqn:sysa} and \ref{eqnsysb} describe the model at the system level.
+Equations \ref{eqn:bayesMod} describe the model at the system level.
 The intercept terms, $a_s$ are given a common higher level normal distribution with mean $\mu^a$ and standard deviation $sigma^a$.
@@
 
-\begin{align}
-log_prod &\sim N(\hat{y}, \sigma)\\ %\label{eqn:likelihood}
+\begin{equation}
+\begin{aligned}
+prod\_scaled &\sim N(\hat{y}, \sigma)\\ %\label{eqn:likelihood}
 \hat{y} &= a_s + b_s months\_operation_i + \mathbf{month}\\ %\label{eqn:obslevel}
-a_s &\sim N(\mu^a, \sigma^a)\\ %\label{eqn:sysa}
-b_s & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \beta^{fy} first\_prod\_year_s + beta^{size} log\_csi\__rating_s + \beta^{cost} log\_cost_s %\label{eqn:sysb}
-\end{align}
+a_s &\sim N(\mu^a, \sigma^a) \label{eqn:bayesMod}\\ %\label{eqn:sysa}
+b_s & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \beta^{fy} first\_prod\_year_s + \beta^{size} log\_csi\_rating_s + \beta^{cost} log\_cost_s
+\end{aligned}
+\end{equation}
 
@@
 
-The $cauchy(0,5)$ prior is what is referred to as a weakly-informative prior.
+The $Cauchy(0,5)$ prior is what is referred to as a weakly-informative prior.
 It provides some information to the MCMC algorithm for where to explore for probability mass, but does not impose any strong prior knowledge that meaningfully effects the posterior distribution.
-The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional large coefficient in a lower level regression, but still provides some shrinkage towards zero.
-\citep{gelman_weakly_2008} The coefficients $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{month}$ are all also given $Cauchy(0,5)$ prior distributions.
+The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional large coefficient in a lower level regression, but still provides some shrinkage towards zero \citep{gelman_weakly_2008}.
+The coefficients $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{month}$ are all also given $Cauchy(0,5)$ prior distributions.
 
-\begin{align}
+\begin{equation}
+\begin{aligned}
 \mu^{lease} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ % \label{eqn:meta1}
 \mu^{sect} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ %\label{eqn:meta2}
-\mu^{manuf} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ %\label{eqn:meta3}
+\mu^{manuf} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) label{eqn:meta} \\ %\label{eqn:meta3}
 \mu^{meta} &\sim Cauchy(0, 5) \\ %\label{eqn:meta4}
 \sigma^{mu} &\sim half-Cauchy(0,5) %\label{eqn:meta5}
-\end{align}
+\end{aligned}
+\end{equation}
 
@@
 
-For the testable implications, the inference we are interested is in the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta{owned}$.
+For the testable implications, the inference we are interested is in the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta_{owned}$.
 Histograms representing the marginal distributions of these contrasts are presented in figure \ref{fig:contrasts}.
@@
 \begin{figure}
-  \includegraphics[width=1\linewidth]{figures/bayes_hypos.png}
+  \centering
+  \includegraphics[width=.6\linewidth]{figures/bayes_hypos.png}
   \caption{Marginal posterior distributions for the contrasts of interest.
@@
 \begin{figure}
-  \includegraphics[width=1\linewidth]{figures/mu_sectors.png}
+  \centering
+  \includegraphics[width=.6\linewidth]{figures/mu_sectors.png}
   \caption{Distribution of the sector contrasts relative to residential sector.
@@
 \section{Discussion and Conclusion}
-The theoretical literature on information asymmetry and quality is established and deep.
-The empirical literature for durable investments is, however, more sparse, reflecting the difficulty of getting reliable data on purchases and long-term reliability.
-This article provides one of few case studies that both identifies a market that could be expected to have issues of information asymmetry of quality, and provides an indirect empirical test for the presence of information asymmetry of quality.
 
-The structure of the emerging solar industry in the US, as well as other parts of the world suggests that issues of information asymmetry may play an important role.
-The established theory on the subject suggests that under information asymmetry of quality, poor quality products may push out good quality products, at least in some sub-markets.
-In this article I have tested an implication of that theory for the case of solar panels in California.
-The results provide some weak evidence for the existence of information asymmetry issues of quality.
+Because of the dramatic price declines in solar panels and the competitive pressure that western producers have faced, it has become common to hear in the industry that solar panels have become commoditised.
+But in finance and economics, commodity has a specific meaning.
+Commodities of a certain grade need to be fungible: goods from different producers are interchangeable.
 
-This article should also be considered as part of a broader literature on the special characteristics of new distributed energy generation technologies.
-The energy investment behavior of home owners, farmers, and small cooperatives are bound to be substantially different than those of large, specialized energy companies that have traditionally done most of the investment in electricity generation.
-Understanding the quickly evolving energy and power industry requires taking into account informational and behavioral factors, and this is a rich field for further empirical research
+Historically, the commoditisation of goods was a necessary condition for the establishment of sophisticated financial markets for those goods.
+The invention of grain elevators for mixing and then loading of grain onto trains in Chicago, was a necessary precursor to the establishment of a futures markets with receipts of delivery traded on the Chicago Board of Trade.\citep{cronon_natures_1992} The abstraction between producer and end purchaser of goods was necessary for the functioning of these markets.
+A trader could fulfil a promise to deliver grain of a certain grade by simply buying it on the market or with cash settlement, rather than having produced it oneself.
+
+ The point of this discussion is that if solar panels truly have become commodities, it would have meaningful and substantial implications for the solar power industry.
+For example, financial instruments could arise providing developers of solar power plants or solar panel retailers the ability to hedge price uncertainty in panels and guarantee supplies for projects with long lead times.
+Particularly relevant to the current environment would be trade policy.
+Trade restrictions on panels from specific countries, like the current US tariffs on Chinese produced panels, would presumably be less effective.
+Fungible panels from China could easily be sold elsewhere, while similar panels from other countries supply the US market.
+
+ Yet the findings from this article strongly suggest that solar panels are not commodities.
+In particular, I find that the salient property of quality varies substantially between producers.
+Because solar power is a distributed technology, where assets can be owned by individual homeowners, businesses and other other organisations that do not have the resources to judge quality, the issue of asymmetric information may then enter the market.
+
+ From a descriptive understanding of the market, I identify a high information type owner (third-party owners) and low information type owner (host-owners) of solar power assets.
+Then, from the basic theory of asymmetric information of quality, I explore two testable implications.
+I find that, as the theory would predict, high information owners are able to attain higher-quality panels.
+With about 90 percent probability, I also find that the cost of solar panel systems owned by high-information types is more highly correlated with quality, as theory would also suggest.
+
+ The market for solar panels and solar panel systems is however still young and dynamic.
+One of the advantages of the linear mixed-effects models and the multilevel models is that we can also get an idea of the degree of differentiation in quality between different manufacturers.
+It appears that a relatively small tail of producers accounted for the worst quality panels, while the majority of panels had
 
 \begin{spacing}{1}
-\bibliographystyle{plainnat}
+\bibliographystyle{chicago}
 \bibliography{solar_prod}
 
-\FloatBarrier
+%\FloatBarrier
 
-\appendix
-\section{Appendix: Tables}
+%\appendix
+%\section{Appendix: Tables}
 

### 2018-03-27 stan_analysis_new.py

@@ -73,7 +73,10 @@ def create_stan_data(solar_data):
     sys_dict = create_num_dict(solar_data["app_num"])
     system_id =app_num.map(sys_dict)
 
-    system_level_df = solar_data[["app_num", "owner_sector", "log_csi_rating", "third_party_owned","module_manufacturer", "first_prod_year", "incentive_step", "log_cost_per_kw"]][~solar_data.app_num.duplicated()]
+    system_level_df = solar_data[["app_num", "host_sector", "owner_sector", "log_csi_rating", "third_party_owned","module_manufacturer", "first_prod_year", "incentive_step", "log_cost_per_kw"]][~solar_data.app_num.duplicated()]
+
+    system_level_df["owner_sector"].value_counts()
+    system_level_df["host_sector"].value_counts()
 
     unique_system_id = system_level_df["app_num"].map(sys_dict)
     J = unique_system_id.size
@@ -128,6 +131,7 @@ def create_stan_data(solar_data):
     "leased":leased
     }
     return(solar_stan_data)
+
 
 def solar_stan_model():
     solar_stan_model1 = """

### 2018-06-01 solar_prod_220518.tex
final mcmc version

--- 2018-03-27 solar_commodities.tex
@@
-% solar_prod.tex
+\documentclass[a4paper]{article}
 
-\documentclass[12pt]{article}
-\usepackage{setspace}
-\usepackage[margin=1in]{geometry}
-\usepackage{mathtools}
-\usepackage{natbib} %for citet and citep
-\usepackage{syntonly}
-\usepackage{esdiff} %for writing partial derivatives
-\usepackage{url} %for inserting urls
-\usepackage{placeins}
-\usepackage{textcomp}
+%% Language and font encodings
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
+
+%% Sets page size and margins
+\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+
+%% Useful packages
 \usepackage{amsmath}
@@
 \usepackage{booktabs}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorlinks=true, allcolors=black]{hyperref}
+\usepackage{natbib}
+\usepackage{url}
+\usepackage{placeins} %for float barrier
 
-\doublespacing % from package setspacs
+%spacing
+\usepackage{setspace}
+\doublespacing
+%\onehalfspacing
 
 \title{Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California}
-
-\date{\today}
-
-\author{Johannes Mauritzen \\ BI Norwegian Business School \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}
+\author{Johannes Mauritzen\thanks{BI Norwegian Business School, johannes.mauritzen@bi.no, \url{http://jmaurit.github.io}}}
 
 \begin{document}
- \begin{spacing}{1} %sets spacing to single for title page
 \maketitle
@@
 I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts.
-2.) Furthermore, prices of solar panels are more correlated to quality for solar panels that are owned by high-information owners.
-I use linear mixed effects models estimated by maximum likelihood and multilevel models estimated by Bayesian Markov Chain Monte-Carlo.
+2.) Furthermore, the price of solar panels that are owned by high-information owners are more highly correlated to quality.
+I use random effects models estimated by maximum likelihood and hierarchical models estimated by Bayesian Markov Chain Monte-Carlo.
 \end{abstract}
 
-% \thanks{*I would like to thank...}
-% JEL Codes: Q4, L71
- \end{spacing}
+\newpage{}
 
@@
 
-Solar panels have been referred to colloquially as being commodities.
-This is in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade.
+Solar panels have been referred to colloquially as being commodities.\footnote{See for example \url{https://www.greentechmedia.com/articles/read/are-pv-modules-a-commodity\#gs.lx3tMbY}} This is often in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade.
 This has made solar power price competitive without subsidies in many sunny locations.
@@
 That is, the good from a certain producer will be largely interchangeable with the good of another producer.
-Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea is sold at the same price, as established by an exchange.
+Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea are sold at the same price, as established by an exchange.
 In a similar manner, the electricity produced by a solar panel at a certain location and time can also be considered a commodity.
@@
 The market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality.
-Solar panels can be characterised as an ``experience'' good, where an investor/consumer needs to learn about the quality through use.
+Solar panels can be characterized as an ``experience'' good, where an investor/consumer needs to learn about the quality through use.
 In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels.
@@
 
-Over the period from 2010 to 2016, many of the larger solar contractors moved to leasing solar power systems to homeowners.
-The introduction of a leasing model could potentially get over this information asymmetry problem by shifting ownership to the large contractors that install and finance the solar systems \citep{mauritzen_cost_2017}.
-These contractors can then in turn take steps such as testing panels and visiting manufacturing sites to ensure quality of suppliers.
-In other words, a descriptive understanding of the market suggests that it is possible to divide the market into two types of solar panel owners: high-information owners who lease the solar panel systems to individual home owners, businesses and other organisations, and low-information owners who buy and own the solar panel systems directly.
+Over the period from 2010 to 2016, many of the larger solar contractors moved to leasing solar power systems to home- and business-owners \citep{mauritzen_cost_2017}.
+These contractors, who buy in scale, have the resources and expertise to judge quality in solar panels.
+They can take steps such as getting panels tested at specialized engineering firms and visiting manufacturing sites to ensure quality of suppliers.
+This descriptive understanding of the market suggests that it is possible to divide the market into two types of solar panel owners: high-information owners who lease the solar panel systems to individual home owners, businesses and other organizations, and low-information owners who buy and own the solar panel systems directly.
 
@@
 In this paper, I use a data set of California solar power systems that includes monthly production data from over 3000 solar panel systems.
-Where many studies of asymmetric information of quality must use proxies for quality, I measure it directly as the average degradation of production over time.
+Where many studies of asymmetric information of quality use proxies for quality, I measure it directly as the average degradation of production over time.
 
-In summary, I have three nested hypothesis that I wish to test:
+I have three nested predictions that I wish to test:
 
@@
   \item Solar panels are not pure commodities. Significant quality differences can be detected between manufacturers of panels.
-  \item High information owners (third-party-owners) will purchase higher quality panels than low-information owners (host-owners)
-  \item Price and quality will be more highly correlated in solar panel systems owned by high information types.
+  \item Given that quality differences exist between manufacturers, high information owners (third-party owners) will purchase higher quality panels than low-information owners (host owners).
+  \item Price and quality will be more highly correlated in solar panel systems owned by high information owners.
 \end{enumerate}
@@
 The data has a natural hierarchy of between 36 and 60 monthly observations within the 3000-plus solar panel systems, which in turn can be grouped according to relevant characteristics such as whether they are host-owned or third-party owned.
-Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model them directly.
+Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model the relevant structure of the data directly.
 For relatively parsimonious models with one or two hierarchies and few covariates, I can use linear mixed effects models (LMEM) estimated by maximum likelihood.
@@
 
-As a solution, I use a Bayesian multilevel\footnote{Sometimes also called hierarchical} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
-The main benefit of using a Bayesian multilevel model is that I can fit a complex model with many hundreds of parameters, while avoiding problems of over-fitting and issues of multiple comparison in inference by relying on regularisation of parameters through the use of higher level parameters.
+As a solution, I use a Bayesian hierarchical\footnote{Sometimes also called multilevel modeling or random effects modeling} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
+This form of modeling has long been established in multivariate dynamic modeling in macroeconomics \citep{doan_forecasting_1984, litterman_forecasting_1986,sims_bayesian_1998, canova_forecasting_2004},  but is also increasingly common in microeconomics and corporate finance.
+One of the main advantages is superior out of sample predictive properties with data with distinct groupings.
+For example \citet{dehejia_was_2003} uses a Bayesian hierarchical model to evaluate multiple programs at different sites.
+By incorporating ``predictive uncertainty'' by way of a hierarchical model, the impact of the sites, which otherwise is estimated to be significant with a fixed-effects model, disappears.
+\citet{meager_understanding_2018} presents a more recent application in the microeconomics litterature.
+She uses a hierarchical model to aggregate studies of microcredit expansion in developing countries in order to estimate an overall treatment effect with ``external validity,'' that is the degree to which an estimated treatment effect generalizes to other populations.
 
-The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organisation literature.
+Within the field of corporate finance, \citet{jones_mutual_2005} uses a hierarchical model to study returns of mutual funds, arguing that estimates of long-run returns of any single fund should be informed, by way of a higher-level distribution, of the returns of the population of other mutual funds.
+In a study with parallels to my own, \citet{korteweg_skill_2017} uses a Bayesian hierarchical model to study persistence of returns in the private equity sector both between and within firms, where persistence is measured by excess variation between firms as captured by firm-level parameters (or random effects) that are drawn from a higher-level distribution.
+If returns are thought of as the converse of production degradation, then the existence of persistence of returns can be thought of as the converse of quality differences.
+
+%In another early application, \citep{chamberlain_hierarchical_1996} used a Bayesian hierarchical model to aggregate multiple weak instruments.
+
+%Corporate finance: Korteweg_markov_2011, Korteweg_skill_2017
+
+The property of hierarchical models that allows for improved predictive performance with the presence of groupings in the data is regularization, or alternatively called partial pooling \citep{gelman_bayesian_2013}.
+As opposed to fixed-effects estimation where a fixed-effects coefficient is determined exclusively with data from within a particular group, in a hierarchical model each of the group parameters are regularized towards a higher-level parameter estimated with the full data in proportion to the relative number of data points in a particular group.
+
+The linear mixed effects models estimated by maximum likelihood that I use in the first section of this article also have have the regularization property.
+Bayesian estimation provides several theoretic advantages, such as correct inference in small samples and generalization to distributions other than the normal.
+However, for my purproses, the main benefit of using a Bayesian methodology is practical.
+I can fit a rich model with several hierarchies and many thousands of parameters, that simply can not be handled by maximum likelihood procedures.
+
+The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organization literature.
 Ch.
@@
 
-Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller -- in other words the seller can choose the level of quality of its product.
-Under these assumptions, the authors predict several equilibrium -- essentially markets for both high and low quality products depending on the extent to which the buyer is informed.
+Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller--in other words the seller can choose the level of quality of its product.
+Under these assumptions, the authors predict several equilibrium--essentially markets for both high and low quality products depending on the extent to which the buyer is informed.
 
 Reputation, or lack thereof, plays a particularly important role in this market.
-Most of both the solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers.
+Most solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers.
 \citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.
@@
 \citet{healy_information_2001} review the early literature.
-Empirical studies in finance have, for example, looked at issues of adverse-selection and quality in Initial Public Offerings \citep{michaely_pricing_1994} and small business lending \citep{petersen_benefits_1994} among many other topics.
+Empirical studies in finance have, for example, looked at issues of adverse-selection and quality in Initial Public Offerings \citep{michaely_pricing_1994}, small business lending \citep{petersen_benefits_1994}, and in a recent application for consumer credit markets \citep{dobbie_information_2013}.
 
-The literature in the marketing field may be more relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables.
-The effects of signalling on quality, usually in the form of advertising, is particularly important in the field.
+The literature in the marketing field is also relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables.
+The effects of signaling on quality, usually in the form of advertising, is particularly important in the field.
 \citet{kirmani_no_2000} provide an overview of this literature.
 
-The use of leasing, discussed in this article, can be seen as a form of costly signalling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of poor quality to the contractor.
-Traditionally, energy generation has been an activity undertaken by large, specialised firms with specialised knowledge and resources.
+The use of leasing, discussed in this article, can be seen as a form of costly signaling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of quality to the contractor.
+Traditionally, energy generation has been an activity undertaken by large, specialized firms with specialized knowledge and resources.
 The role of information asymmetry in investment and purchase decisions has likely been seen as a secondary issue, and to my knowledge, no literature exists on the role of such informational issues in the industry.
 
-However, the growth of distributed energy technologies, like solar panels, has added new focus on informational and behavioural issues related to energy investments by homeowners and small businesses.
-In turn, a growing literature is forming around solar power investment behaviour of homeowners and small businesses.
+However, the growth of distributed energy technologies, like solar panels, has added new focus on informational and behavioral issues related to energy investments by homeowners and small businesses.
+In turn, a growing literature is forming around solar power investment behavior of homeowners and small businesses.
 
-For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption - showing evidence for a ``solar price premium'' in homes with solar panels.
+For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption, showing evidence for a ``solar price premium'' in homes with solar panels.
 \citet{bollinger_peer_2012} study the the role of peer effects in solar power adoption.
@@
 Notice that variables can vary by monthly observation or by solar power system.
-I will take account of this natural data hierarchy in my modelling.
+I will take account of this natural data hierarchy in my modeling.
 There are a total of 171,027 monthly observations used in the analysis.
@@
 \begin{table}
+  \caption{Summary of data.
+Information is provided on whether whether the variables are at the observation (system-month) level or at the system level, Whether the variable values are real, integer or categorical in nature and the mean value of observations is also provided if relevant.
+The parenthesis after the sector categories in the footnote indicate the number of systems for each sector.}
   \begin{tabular}{rllll}
@@
     Months of operation &  months\_operation &  Observation & Integer & --   \\
-    Month of year & month & Observeration & Categ.$^{1}$& -- \\
+    Month of year & month & Observation & Categ.$^{1}$& -- \\
     Cost per kw & cost\_per\_kw & System & Real & 7,026 \\
@@
     Owner sector & own\_sector\_id & System & Categ.$^{3}$& -- \\
+    Host county & county\_id & System & Categ. & -- \\
     \midrule
@@
   \label{tbl:variables}
-  \caption{}
 \end{table}
@@
 Production is measured in kilowatt hours (kWh).
-In the analysis, production data is first log transformed and then descaled, as shown in equation \ref{eqn:descaled}.
+In the analysis, production data is first log transformed and then de-scaled, as shown in equation \ref{eqn:descaled}.
 For production data for each system, $s$, the log production data is demeaned and then divided by 2 times the standard deviation of the production data for that solar power system.
-Figure \ref{fig:production} gives a visualisation of the transformed production data.
+Figure \ref{fig:production} gives a visualization of the transformed production data.
 The red line shows the average degradation over time of all the systems.
@@
 
-Other continuous variables are transformed in a similar manner, with observation level variables descaled at the system level, and system-level variables descaled across systems.
+Other continuous variables are transformed in a similar manner, with observation level variables de-scaled at the system level, and system-level variables de-scaled across systems.
 These transformations avoid confusion that may come when interpreting coefficients on variables with different units and scales.
 There are also technical reasons for the transformations relating to the MCMC routine.
-Conversion in the MCMC algorithm is aided when all variables are on the same scale.
+Convergence in the MCMC routine is aided when all variables are on the same scale.
 Setting appropriate weakly informative prior distributions on the parameters is also simplified when continuous variables are free from units.
@@
 \begin{equation}
-prod\_scaled_{s[i]} = \frac{log(prod\_kwh_{s[i])} - mean(log(prod\_kwh_{s[i]}))}{2 std(log(prod\_kwh_{s[i]}))}
+prod_{s[i]} = \frac{prod\_kwh_{s[i]} - mean(prod\_kwh_{s[i]})}{std(prod\_kwh_{s[i]})}
 \label{eqn:descaled}
@@
 
-Month of year dummy variables are used to control for the seasonality in the data.
+Month-of-year dummy variables are used to control for the seasonality in the data.
 The capacity of the solar panel system is the actual maximum capacity of a solar power system as reported to and audited by the CSI.
@@
 \subsection{Testing for differences between module manufacturer}
-Following the notation of \citet{bates_fitting_2015}, the conditional distribution of the response variables, $y$ in a multilevel model can be written compactly as in equation \ref{equation:lme_simple}.
-Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or "fixed effects") of the $n times p$ matrix of (p-1) explanatory variables $\mathbf{X}$ plus intercept term.
-$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or "random effects" $\mathbf{B}$ that are fixed at values $\mathbf{b}$.
-\footnote{The use of "random" refers to the fact that these groups are assumed to be specific to the sample.
-Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms, thus the "random" label.
-Still, interpreting coefficients as group varying or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.
+Following the notation of \citet{bates_fitting_2015}, the conditional distribution of the response variables, $y$ in a linear mixed effects model can be written compactly as in equation \ref{equation:lme_simple}.
+Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or  ``fixed effects'') of the $n \times p$ matrix $\mathbf{X}$ of (p-1) explanatory variables plus intercept term.
+$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modeled directly with corresponding vector of coefficients, or ``random effects'' $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of ``random'' refers to the fact that these groups are assumed to be specific to the sample.
+Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms.
+Interpreting coefficients as group varying coefficients or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.
 $\mathbf{W}$ is a diagonal matrix of known prior weights.
-$B$ is distributed as multivariable normal with zero mean and where $\Sigma$ is variance-covariance matrix.
+$B$ is distributed as multinomial normal with zero mean and where $\Sigma$ is variance-covariance matrix.
 
@@
 
-Models of this form can be solved efficiently using a penalised maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.
+Models of this form can be solved efficiently using a penalized maximum likelihood routine. For further details of both the model and routine, I refer to \citet{bates_fitting_2015}.
 
@@
 
-In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modelled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$.
+In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modeled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$.
 Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by the individual system.
-Thus in this model formulation I take into account the groupings by system, explicitly modelling varying relative initial production values (intercepts) and degradation over time (slope).
-Moving up a level, the estimated $a$ and $b$ parameters are in turn modelled as multivariate normal.
+Thus in this model formulation I take into account the groupings by system, explicitly modeling varying relative initial production values (intercepts) and degradation over time (slope).
+Moving up a level, the estimated $a$ and $b$ parameters are in turn modeled as multivariate normal.
 
@@
 \begin{aligned}
-prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
 \begin{pmatrix}
@@
 
-The inclusion of the matrix of month dummies $\mathbf{month}$ controls for seasonal variation across all systems.
-This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and the seasonality is not of direct importance, this should suffice.
+The inclusion of the matrix of month dummies, $\mathbf{\mu^{month}}$, controls for seasonal variation across all systems.
+This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and a more direct modeling of seasonality is not of direct importance, this should suffice.
 As long as un-modeled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.
@@
 To test whether quality varies among module manufacturers, I simply add panel manufacturer as an extra hierarchy to the model.
-In this model, the observation level regression is identical to the restricted comparison model as in equation \ref{eqn:observation_level_rest}.
-The system-level intercepts, $a$, and slope, $b$, terms are also modelled as a multivariate normal.
+In this model, the observation-level regression is identical to the restricted comparison model as in equations \ref{eqn:observation_level_rest}.
+The system-level intercepts, $a$, and slopes, $b$, terms are also modeled as multivariate normal.
 But the mean values of the multivariate normal distribution, $\mu_a$ and $\mu_b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.
@@
 \end{pmatrix}
-&\sim N
+&\sim N \Bigg(
 \begin{pmatrix}
@@
 \begin{pmatrix}
-  \sigma_{a[m]}^2 & \rho \sigma_{a[m]} \sigma_{b[m]} \\
+  \sigma_{a[m]}^2 & \rho \sigma_{a} \sigma_{b} \\
   \rho \sigma_a \sigma_b & \sigma_b^2
-\end{pmatrix} \\
+\end{pmatrix} \Bigg) \\
 \mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
@@
 Figure \ref{sfig:sys_slope_fig} gives a visual summary of the system-level slope terms ordered from lowest to highest.
-The lines represent approximate 95\% confidence intervals around the point estimate.
+The lines represent approximate 95\% confidence intervals around the point estimates.
 One anomaly we can notice is that a handful of the systems have slopes that are estimated to be positive.
@@
 \begin{figure}
-\begin{minipage}{.48\textwidth}
+  \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, $b_i$, ordered from lowest to highest.
+Visually, we can see significant differences in the average degradation over time across solar power systems.}
   \includegraphics[width=1\linewidth]{figures/sys_slope_fig.png}
-  \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, ordered from lowest to highest.}
   \label{sfig:sys_slope_fig}
-\end{minipage}\hfill
-\begin{minipage}{.48\textwidth}
-  \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
-  \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level slope terms $\mu_b$, ordered from lowest to highest.}
-  \label{sfig:nested_manuf_fig}
-\end{minipage}
 \end{figure}
 
-\ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes.
+Figure \ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes.
 Again the lines represent 95\% confidence intervals around the point estimates.
 Visually, we can see substantial variation in quality between the manufacturer groups.
-We can more formally test the hypothesis that manufacturer groupings improve the explanatory power of the model by comparing to the restricted model with an LM test.
+We can formally test the hypothesis that manufacturer groupings improve the explanatory power of the model with a comparison to the restricted model with a likelihood ratio test.
 The results are shown in table \ref{tbl:lm_commodity}.
-The information criteria measures (AIC, BIC) are both lower, indicating a better predictive power of the model with manufacturer groupings.
-The chis-square statistic also indicates that the model with manufacturer groupings substantially improves model performance.
+The information criteria measures (AIC, BIC) are both lower, indicating better predictive power of the model with manufacturer groupings.
+The Chi-square statistic indicates that the model with manufacturer groupings substantially improves model performance.
+
+\begin{figure}
+  \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level mean slope terms $\mu_b$, ordered from lowest to highest.
+These are the means of the higher level normal distributions that vary by manufacturer that the system-level parameters are assumed to be drawn from.
+The visual impression is of substantial variation between manufacturer groupings.}
+  \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
+  \label{sfig:nested_manuf_fig}
+\end{figure}
 
@@
     \midrule
-             Null &  17 &  8805 &  8976 &   -4385 &      8771 &     -- &      -- &          -- \\
+             Restricted &  17 &  8805 &  8976 &   -4385 &      8771 &     -- &      -- &          -- \\
     Grouped Manuf. &  20 &  8430 &  8631 &   -4195 &      8390 &    381 &       3 &     2.6e-82 \\
@@
   \label{tbl:lm_commodity}
-  \caption{A likelihood ratio test indicates that the addition of manufacterer as a explanatory variable significantly improves the predictive power of the model}
+  \caption{A likelihood ratio test indicates that the addition of panel manufacturer as a hierarchy significantly improves the predictive power of the model.
+Akaike Information Criterion (AIC) and Bayesian Information Criterion are lower in the model with manufacturer groupings, indicating better out-of-sample predictive fit.}
 \end{table}
 
-The first result, of which the evidence appears to be quite strong is that solar panel systems do appear to vary substantially in quality between module manufacturers.
-Given that module quality is an important differentiating characteristic of panels, it is then difficult to formally define solar panels as commodities.
+The first result, of which the evidence appears to be quite strong, is that solar panel systems do appear to vary substantially in quality between module manufacturers.
+Given that module quality is a salient differentiating characteristic between panels, it is then difficult to view solar panels as commodities.
 
-\subsection{testing for differences in quality between leased and host-owned solar power systems}
+\subsection{Testing for asymmetric information: differences in quality between leased and host-owned solar power systems}
 
-Provided the evidence that solar panel systems are not commodities and that substantial differences in quality do exist between manufacturers, we can now move on to the testable implication that comes from a combination of standard economic theory of asymmetric information and a descriptive understanding of the market that suggests the market for solar panels could plausibly be subject to issues of asymmetric information of quality.
+Provided the evidence that solar panel systems are not commodities and that substantial differences in quality do exist between manufacturers, we can now move on to the testable implication that comes from a combination of standard economic theory of asymmetric information and a descriptive understanding of the market that suggests solar panels could be subject to issues of asymmetric information of quality.
 
-From our informal model of asymmetric information where we see host owners of solar systems as being low-information types, while owners of leased panels as being high information types, we would predict that on average leased panels will display higher quality over time.
+From our informal model of asymmetric information where we see host owners of solar systems as being low-information types, while owners of leased panels as being high-information types, we would predict that on average leased panels will display higher quality over time.
 To test this directly with the data, I only need to do a slight modification of the model in the previous subsection.
@@
 \begin{aligned}
-prod\_scaled_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+log_prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
 \begin{pmatrix}
@@
 \begin{pmatrix}
-  \sigma_{a[l]}^2 & \rho \sigma_{a[l]} \sigma_{b[l]} \\
+  \sigma_{a}^2 & \rho \sigma_{a} \sigma_{b} \\
   \rho \sigma_a \sigma_b & \sigma_b^2
-\end{pmatrix} \\
+\end{pmatrix} \Bigg) \\
 \mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
@@
 The difference is also economically significant.
-Extrapolating out from the difference in the point estimates between leased and host-owned suggests that on average a leased system can expect to be producing 11\% more power after 10 years compared to an otherwise identical solar panel systems.
+Extrapolating out from the difference in the point estimates between leased and host-owned panels suggests that on average a leased system can expect to be producing 11\% more power after 10 years compared to an otherwise similar solar panel system that is host-owned.
 
@@
   \centering
-	\includegraphics[width=.8\textwidth]{figures/lease_sys_fig.png}
-	\caption{A visual summary of the estimation results. The vertical lines again represent the estimated system-level slope coefficients.}
+	\includegraphics[width=1\textwidth]{figures/lease_sys_fig.png}
+	\caption{A visual summary of the estimation results where panels are grouped by owner type: leased (high information third-party owner) or host-owned (low information owner).
+The vertical lines again represent 95\% confidence interval on the estimated system-level slope coefficients, which are sorted first by ownership type (host-owned on the left, leased on the right), and then from low to high of the maximum likelihood point estimate of the slope estimates.
+The red line represents the zero threshold.
+The horizontal black line on the left represents the mean of the higher level distribution on the slopes for host-owned systems, $\mu_b[l=0]$, while the dotted black line on the right represents the mean of the higher level distribution on the slopes for third-party owned systems, $\mu_b[l=1]$ }
 	\label{lease_sys_fig}
@@
 
-\section{Full Bayesian multilevel model estimated by Markov Chain Monte Carlo}
-Until now, I have used relatively parsimonious models with few fixed-effects covariates and only one or maximum two hierarchies ("random effects").
+\section{Full Bayesian hierarchical model}
+Up until now, I have used relatively parsimonious models with few fixed-effects and only one or two hierarchies ("random effects").
 Ideally, I would like to estimate a full probability model with all relevant and available covariates and hierarchies.
-First and foremost so that the parameters of interests to the testable implications can be estimated from one model.
-But also, \citet{barr_random_2013} shows, using Monte-Carlo simulations, that mixed effects models that have the maximum random effects structure justified by the model and data, provide the best out-of-sample prediction.
-Estimating mixed effects model by maximum likelihood is efficient and straight-forward.
+First and foremost so that the parameters of interests to the testable implications can be estimated simultaneously from one model.
+But also, as \citet{barr_random_2013} shows, mixed-effects models that have the maximum random effects structure justified by the model, provide the best out-of-sample prediction.
+Estimating mixed-effects models by maximum likelihood is efficient and straight-forward.
 However, convergence of the maximum likelihood routine is not guaranteed under richer models with more hierarchies and larger number of parameters to be estimated.
 
-Specific to my model, I wish to estimate a model where the variable for cost per-kW is included as an interaction that varies by system and module manufacturer.
-I also wish to have a model where both system, module manufacturer and leasing hierarchies are estimated simultaneously.
+I wish to estimate a model where the variable for cost per-kW is included as an interaction that varies by system and module manufacturer.
+I also wish to have a model where both system, module manufacturer, and leasing hierarchies are estimated simultaneously.
+Geographic information on the location of solar panels could also conceivably be correlated with degradation of solar panels, and should be accounted for in the model.
 I am not able to obtain convergence of the maximum likelihood routines under this rich model.
 
-As a solution, I build a fully Bayesian multilevel model\footnote{I purposefully change my nomenclature here and refer to the Bayesian model as a "multilevel model" to distinguish it from the models estimated by maximum likelihood, which I refer to as a linear mixed effects model.
-This distinction is however mostly artificial.
-Both models resemble each other closely in structure.
-"Multilevel model" is however a broader term.
-There is no restriction to linearity, for example.
-Models with non-linear or non-parametric components can also be incorporated and estimated with MCMC.} estimated by Markov Chain Monte Carlo (MCMC) simulation techniques.
+As a solution, I build a fully Bayesian hierarchical model\footnote{I purposefully change my nomenclature here and refer to the Bayesian model as a ``hierarchical mode'' to distinguish it from the models estimated by maximum likelihood, which I refer to as a linear mixed effects model.
+This distinction is mostly semantic.
+Both models resemble each other in structure.
+"hierarchical model" is however a broader term.
+There is no restriction to linearity.
+Models with non-linear or non-parametric components can also be incorporated and estimated.} estimated by Markov Chain Monte Carlo (MCMC) simulation techniques.
 
-The main advantage to using the Bayesian techniques here are practical rather than philosophical.
-Models up to levels of nearly arbitrary richness can be estimated, with computation time and available data being the main constraints.
-Inference is also well defined and exact.\footnote{Exact in the sense that given a correctly specified probability model, then an MCMC routine that is given infinite time to sample, will in theory converge to the correct posterior probability distribution.
+Multilevel models estimated using Bayesian techniques have some nice theoretical properties, as I discussed earlier.
+However, the main advantages to using the Bayesian techniques here are practical.
+Models of nearly arbitrary richness can be estimated, with computational time and available data being the main constraints.
+Inference is also well-defined and exact.\footnote{Exact in the sense that given a correctly specified probability model, then an MCMC routine that is given infinite time to sample, will in theory converge to the correct posterior probability distribution.
 Maximum likelihood is approximate in the sense that inference is based on asymptotic theory and assumptions of normality.}
@@
 Group-level variances are not constrained to be equal and assumptions of normality of the likelihood and posterior distributions can be loosened.
-For further discussion of the Bayesian frameworks and techniques, I refer to the excellent discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2015}.
+For further discussion of the Bayesian frameworks and techniques, I refer to discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2015}.
 
-The model is presented in equations x through y.
-The response variable is production data that has been log transformed and scaled as described earlier.
-This response variable is assumed to have a normal likelihood with mean, $\hat{y}$ and standard deviation, $\sigma$.
-At the observation level, the fitted values, $\hat{y}$, are modelled as in equation \ref{obslevel}.
-This observation level regression, as before, is a simple linear regression with an intercept $\alpha$, and a slope $\beta$ which are both allowed to vary by system.
-In addition a vector of month dummies, $\mathbf{month}$ is included in order to control for seasonality.
+The model is described by the equations in \ref{eqn:bayesMod}.
+The response variable is production data that has been de-scaled as described earlier.
+This response variable is given a normal likelihood with mean, $\hat{y}$ and standard deviation, $\sigma$.
+At the observation level, the fitted values, $\hat{y}$, are modeled as a simple linear regression, with an intercept $a_s$, and a slope $b_s$ which are both allowed to vary by system, $s$.
+In addition a vector of month dummies, $\mathbf{\mu^{month}}$ is included in order to control for seasonality.
 
-Equations \ref{eqn:bayesMod} describe the model at the system level.
-The intercept terms, $a_s$ are given a common higher level normal distribution with mean $\mu^a$ and standard deviation $sigma^a$.
-But because the intercept term is not of direct economic importance, it is not modelled further.
+At the system level, the intercept terms, $a_s$ are given a common higher level Cauchy distribution with location parameter $\mu^a$ and scale parameter $\sigma^a$.
+But because the intercept term is not of direct economic importance, it is not modeled further.
 
-The slope terms, $b_s$ are again modelled directly.
-I decompose the variation into three mean terms that vary by lease, $\mu^{lease}_l$, host sector $\mu^{sect}_c$, and manufacturer $\mu^m_m$.
-In addition three continuous variables are included in the system-level equation: the year of first production, $first\_prod\_year_s$, log capacity, $log\_capacity_s$ and the log system cost per kilowatt, $log\_cost_s$ with corresponding coefficients $\beta^{fy}$, $\beta^{size}$, and $\beta^{cost}$.
+The slope terms, $b_s$ are also modeled as Normal distributions with mean $\mu^{b}_s$, that is allowed to vary by system, and standard deviation $\sigma^{b}$.
+I decompose the variation of $\mu^{b}_s$ into three terms that vary by lease, $\mu^{lease}_l$, host sector $\mu^{sect}_c$, and manufacturer $\mu^m_m$.
+In addition three continuous variables are included in the system-level equation: the year of first production, $first\_prod\_year_s$, capacity, $capacity_s$ and the system cost per kilowatt, $cost_s$ with corresponding coefficients $\beta^{fy}$, $\beta^{size}$, and $\beta^{cost}$.
 
@@
 \begin{aligned}
-prod\_scaled &\sim N(\hat{y}, \sigma)\\ %\label{eqn:likelihood}
-\hat{y} &= a_s + b_s months\_operation_i + \mathbf{month}\\ %\label{eqn:obslevel}
-a_s &\sim N(\mu^a, \sigma^a) \label{eqn:bayesMod}\\ %\label{eqn:sysa}
-b_s & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \beta^{fy} first\_prod\_year_s + \beta^{size} log\_csi\_rating_s + \beta^{cost} log\_cost_s
+prod_{st} &\sim N(\hat{y}_{st}, \sigma)\\ %\label{eqn:likelihood}
+\hat{y}_{st} &= a_s + b_s months\_operation_{st} + \mathbf{\mu^{month}}\\
+a_s &\sim N(\mu^a, \sigma^a)  \label{eqn:bayesMod} \\
+b_s &\sim N(\mu_s^b, \sigma^b) \\
+\mu_{s}^{b} & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \mu^{county}_c  \\
+& +\beta^{fy} first\_prod\_year_s + \beta^{size} log\_csi\_rating_s + \beta^{cost} log\_cost_s
 \end{aligned}
@@
 
-The $\mu$ terms are all in term modelled as coming from a Cauchy distribution with a location meta-parameter $\mu^{meta}$ and scale parameter $\sigma^{mu}$, which are also estimated by the model.
-These meta-parameters are in turn given Cauchy prior distributions with location 0 and scale 5.
+Higher level terms are all given Cauchy prior distributions as summarized in the equations in \ref{eqn:meta}.
+The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional outlying coefficient in a lower level regression, while still provides some regularization towards zero \citep{gelman_weakly_2008}.
+
+The $\mu^{lease}_l$, $\mu^{sect}_s$, and $\mu^{manuf}$ terms are all
+modeled as coming from a Cauchy distribution with a location meta-parameter $\mu^{meta}$ and scale parameter $\sigma^{mu}$.
+Similarly, $\beta^{fy}$, $\beta^{size}$ and $\beta^{cost}_l$ are modeled as coming from a $Cauchy(\mu^{\beta}, \sigma^{\beta})$ distribution.
+The location parameters $\mu^{meta}$ and $\mu^{\beta}$ are in turn given a Cauchy(0,5) prior, while the corresponding $\sigma^{meta}$ and $\sigma^{beta}$ terms are given half-Cauchy(0,5) priors--essentially a Cauchy distribution constrained to the positive values.
 
 The $Cauchy(0,5)$ prior is what is referred to as a weakly-informative prior.
-It provides some information to the MCMC algorithm for where to explore for probability mass, but does not impose any strong prior knowledge that meaningfully effects the posterior distribution.
-The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional large coefficient in a lower level regression, but still provides some shrinkage towards zero \citep{gelman_weakly_2008}.
-The coefficients $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{month}$ are all also given $Cauchy(0,5)$ prior distributions.
+It imposes some weak regularization on the parameters, while also allowing for some outlying parameter estimates.
+At the same time, the prior does not impose any strong prior knowledge that meaningfully effects the posterior distribution, especially in the presence of substantial data.
+The parameters $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{\mu^{month}}$ are all also given $Cauchy(0,5)$ prior distributions.
 
@@
 \begin{aligned}
-\mu^{lease} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ % \label{eqn:meta1}
-\mu^{sect} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\ %\label{eqn:meta2}
-\mu^{manuf} &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) label{eqn:meta} \\ %\label{eqn:meta3}
-\mu^{meta} &\sim Cauchy(0, 5) \\ %\label{eqn:meta4}
-\sigma^{mu} &\sim half-Cauchy(0,5) %\label{eqn:meta5}
+\mu^{lease}_l &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\
+\mu^{sect}_s &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\
+\mu^{manuf}_m &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \label{eqn:meta} \\
+\mu^{meta} &\sim Cauchy(0, 5) \\
+\mu^{a} &\sim Cauchy(0,5) \\
+\mathbf{\mu^{month}} &\sim Cauchy(0,5) \\
+\beta^{fy} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta}) \\
+\beta^{size} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta})\\
+\beta_l^{cost} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta})\\
+\mu^{\beta} &\sim Cauchy(0, 5) \\
+\sigma &\sim half-Cauchy(0,5) \\
+\sigma^{a} &\sim half-Cauchy(0,5) \\
+\sigma^{b} &\sim half-Cauchy(0,5) \\
+\sigma^{\mu} &\sim half-Cauchy(0,5)\\
+\sigma^{\beta} &\sim half-Cauchy(0,5)\\
 \end{aligned}
@@
 
-\subsection{Computation and Results from Bayesian multilevel model}
+\begin{figure}
+  \centering
+	\includegraphics[width=1\textwidth]{figures/solar_prod_hierarchy.png}
+	\caption{An overview of the structure of the model.
+Starting from the top in panel I, the monthly production data is modeled with a normally distributed likelihood.
+Moving down to II, the fitted values, $\hat{y}_{st}$, are in turn decomposed into system-varying slope and intercept terms, as well as month-dummies that control for seasonality.
+Further down to III, the slope terms, $b_s$ are modeled as normal distributions with varying mean parameters, $\mu_b^s$ that decompose further into terms representing variables of interest and control variables.
+Finally, in IV, all parameters are given higher-level meta-parameters that serve to regularize (or "shrink") the lower-level parameter estimates.}
+	\label{solar_prod_hierarchy}
+\end{figure}
+
+In the earlier section, we noticed that some of the system slope parameters were estimated to be positive.
+There are intuitive and plausible reasons for observing higher average production some time after the initial installation.
+The placement of solar panels could have been optimized after the fact.
+Malfunctioning components could also have been replaced.
+However, when we wish to interpret the $b_s$ parameters as degradation over time, we should ideally discount positive values.
+Several possibilities present themselves.
+I could discard the systems with positive slope parameters from the data used.
+But discarding data and the associated information is generally not a desirable.
+Alternatively, I could keep in all data, allowing for the slope terms to be positive.
+After all, one of the advantages of the hierarchical model with weakly regularizing priors is that it allows for outlying parameter estimates, without unduly effecting the overall results.
+A third solution is to include all the data, but constrain the model by having priors that place zero probability on positive slope parameter values.
+$b_s$ parameters would be given half-Cauchy distributions with support between 0 and negative infinity.
+This has the desired effect of constraining slope parameters to be negative by imposing prior knowledge.
+The downside is that such strong priors can sometimes lead to unintended effects on the posterior distribution and can lead to faulty inference.
+In the proceeding analysis I therefor present results from both constrained and unconstrained models.
+
+\subsection{Computation and Results from Bayesian hierarchical model}
 
@@
 The sampler was run with four chains and 1000 iterations.
-R-statistics of 1 indicate convergence in the model.
-2000 samples are taken from the estimated posterior distribution, and model results for the marginal distributions of the parameters of interest are presented in the form of histograms of the samples.
+A Gelman-Rubin convergence statistic ($\hat{R}$ of 1 indicated convergence in the model.
+2000 samples are taken from the simulated posterior distribution, and model results for the marginal distributions of the parameters of interest are presented in the form of kernel densities of the samples.
 
-Recall, the testable implications from the theory of asymmetric information and quality were that 1.) average quality should be higher in the panels owned by the high-information type (leased panels).
+Recall that the testable implications from the theory of asymmetric information and quality were that 1.) average quality should be higher in the panels owned by the high-information type (leased panels).
 2.) Price should be more highly correlated with quality in the panels owned by the high-information type.
 
-The parameters of interest in our model are then the higher level parameters $\mu_{lease}$ and $\beta$.
-The first represents the mean value of the system level slope terms, $b$ grouped by whether the systems were leased or host-owned.
+The parameters of interest in our model are then the higher level parameters $\mu_l^{lease}$ and $\beta_l$.
+The first represents the mean value of the system level slope terms, $b_s$ grouped by whether the systems were leased or host-owned.
 The latter is a measure of the correlation between the cost of the system and the system slope terms.
-The $\beta$ parameter is also allowed to vary by leased or host-owned.
-It is important to note, that the estimated $\beta$ paramaters are estimated conditional on the installation date of the solar panels, thus they do not reflect changes in price over time.
+The $\beta_l$ parameter is also allowed to vary by leased or host-owned.
+It is important to note, that the estimated $\beta_l$ parameters are estimated conditional on the installation date of the solar panels, thus they do not reflect changes in price over time.
 This could lead to a heavy bias in the results if, as seems plausible, both prices and quality changed over time.
@@
 For the testable implications, the inference we are interested is in the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta_{owned}$.
-Histograms representing the marginal distributions of these contrasts are presented in figure \ref{fig:contrasts}.
+Kernel densities representing the marginal distributions of these contrasts are presented in figure \ref{fig:contrasts}.
 The top panel shows the distribution of the contrast $\mu_{lease, yes} - \mu_{lease, no}$.
 Essentially all of the probability mass is located in the positive range.
-This reinforces the finding from the mixed effects model that leased panels due appear to have substantially higher quality, as the theory of asymmetric information of quality would suggest.
+This reinforces the finding from the mixed effects model that leased panels do appear to have substantially higher quality, as the theory of asymmetric information of quality would suggest.
 
@@
 
+Figure \ref{fig:contrasts_const} show the corresponding distributions under the constrained model.
+$\mu_{lease, yes} - \mu_{lease, no}$ is pulled somewhat towards zero in magnitude, but nearly all of the probability mass is still in the positive range.
+The distribution of  $\beta_{leased}- \beta_{owned}$ is also pulled towards zero, and now 72\% of the probability mass is in the positive range.
+
 \begin{figure}
+\begin{minipage}{.48\textwidth}
   \centering
-  \includegraphics[width=.6\linewidth]{figures/bayes_hypos.png}
+  \includegraphics[width=1\linewidth]{figures/bayes_hypos.png}
+  \label{fig:contrasts}
+ \end{minipage}\qquad
+\begin{minipage}{.48\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/bayes_hypos_const.png}
+  \label{fig:contrasts_const}
+ \end{minipage}
+\bigskip
+\begin{minipage}[t]{.48\textwidth}
+\centering
 \caption{Marginal posterior distributions for the contrasts of interest.
-The top panel shows that the distributions of the slope parameters, $b_s$, are estimated to have means that are significantly different from each other between leased and not leased systems as defined as the contrast, $\mu_{lease, yes} - \mu_{lease, no}$.
-The lower panel shows the marginal posterior distribution of the contrast between the $\beta$ parameter for leased and not-leased solar systems.
+The top panel shows that the distributions of the slope parameters, $b_s$, are estimated to have means that are significantly different from each other between leased and host-owned systems as defined by the contrast, $\mu_{lease, yes} - \mu_{lease, no}$.
+The lower panel shows the marginal posterior distribution of the contrast between the $\beta$ parameter for leased and host-owned solar systems.
 The results indicate that there is an approximately 89\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned.}
-  \label{fig:contrasts}
+\end{minipage}\qquad
+\begin{minipage}[t]{.48\textwidth}
+\centering
+ \caption{Marginal posterior distributions for the contrasts when the $b_s$ paramaters and corresponding higher-level mean parameters are constrained to have a maximum of zero.
+The contrast, $\mu_{lease, yes} - \mu_{lease, no}$ is materially unchanged and continues to have nearly its entire distribution in a positive range, though it is pulled somewhat towards zero.
+The contrast $\beta_{lease, yes} - \beta_{lease, no}$ is also pulled towards zero, and now only approximately 72 \% of the probability is in a positive range.
+The results indicate that there is an approximately 72\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned when all higher level mean parameters are constrained to have a maximum of zero.}
+\end{minipage}
 \end{figure}
 
-A direct comparison of prices and costs between these groups is however complicated.
+A direct comparison of costs between leased and host-owned solar power systems is somewhat complicated.
 The prices of host-owned panels are market prices: simply the purchase price that the owner of the solar panel system paid to the contractor.
@@
 
-When that is said, if we assume that misreporting of the true costs will tend to lead to less variation in reported costs, then that should have the affect of attenuating the magnitude of the contrast, $\beta_{cost, l=yes} - \beta_{cost, l=no}$.
+However, if we assume that misreporting of the true costs will tend to lead to less variation in reported costs, then that should have the effect of attenuating the magnitude of the contrast, $\beta_{cost, l=yes} - \beta_{cost, l=no}$.
 In that case, the estimated contrast can be seen as being a conservative estimate.
 
-Other higher level parameters are reported in table \ref{table:higher_level_params}.
-The sector groups can be of interest in a discussion of the role of asymmetric information.
-Figure \ref{mu_sectors} shows the distributions of the contrasts relative to the residential host sector.
+Summary statistics of the distributions of higher level parameters are reported in table \ref{table:higher_level_params} and table \ref{table:higher_level_params_const} for the constrained model.
+The sector groupings, $\mu^{sect}$ can be of interest in a discussion of the role of asymmetric information.
+Figure \ref{fig:mu_sectors} shows the distributions of the contrasts relative to the residential host sector.
 The top panel represents the contrast with commercial sector hosts, the middle with government sector host and the bottom panel with non-profit sector host.
 All three contrasts indicate relatively lower quality for installations for these sectors compared to the residential sector.
+Results of the constrained model are shown in figure \ref{fig:mu_sectors_const}.
+The results are similar, though the difference in average quality between the government and residential sector is now centered closely around zero.
 
-While it is not clear why this should be, an explanation consistent with the theory of asymmetric information of quality, is that there exists an agency problem.
+While it is not clear why this should be, an explanation consistent with the theory of asymmetric information of quality is that there exists an agency problem.
 A homeowner will, in some form, have a direct financial incentive to acquire solar panels of good quality.
@@
 \label{table:higher_level_params}
-\caption{Summary statistics of the estimated posterior distributions of higher level variables}
+\caption{Summary statistics of the estimated posterior distributions of higher level parameters}
 \end{table}
 
-\begin{figure}
+% constrained table
+
+\begin{table}
+\begin{tabular}{lrrrrrrrrr}
+\toprule
+{} &    mean &    std &     min &    2.5\% &     25\% &     50\% &     75\% &   97.5\% &           max \\
+\midrule
+beta1\_fy0    & -0.017' & 0.001' & -0.021' & -0.019' & -0.018' & -0.017' & -0.017' & -0.016' & -1.454096e-02 \\
+meta\_mu0     & -0.006' & 0.001' & -0.009' & -0.008' & -0.007' & -0.006' & -0.006' & -0.005' & -3.700574e-03 \\
+meta\_beta0   &  0.012' & 0.798' & -4.910' & -1.877' & -0.083' & -0.009' &  0.080' &  1.989' &  6.342196e+00 \\
+meta\_mu\_mon0 & -0.114' & 0.034' & -0.280' & -0.180' & -0.137' & -0.115' & -0.093' & -0.044' &  3.362246e-02 \\
+sigma\_mu10   &  0.003' & 0.001' &  0.002' &  0.002' &  0.003' &  0.003' &  0.003' &  0.004' &  5.319401e-03 \\
+mu\_b00       &  0.121' & 0.011' &  0.088' &  0.101' &  0.113' &  0.121' &  0.129' &  0.142' &  1.531909e-01 \\
+sigma\_b00    &  0.476' & 0.006' &  0.459' &  0.464' &  0.472' &  0.476' &  0.480' &  0.487' &  4.995173e-01 \\
+sigma\_mon0   &  0.114' & 0.029' &  0.055' &  0.075' &  0.094' &  0.108' &  0.127' &  0.188' &  2.955012e-01 \\
+sigma\_beta0  &  0.800' & 1.495' &  0.006' &  0.010' &  0.051' &  0.201' &  0.814' &  5.056' &  1.976847e+01 \\
+sigma0       &  0.065' & 0.000' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  6.572528e-02 \\
+mu\_mon0      & -0.285' & 0.007' & -0.298' & -0.297' & -0.290' & -0.286' & -0.278' & -0.275' & -2.736025e-01 \\
+mu\_mon1      & -0.243' & 0.007' & -0.256' & -0.254' & -0.248' & -0.243' & -0.235' & -0.233' & -2.310124e-01 \\
+mu\_mon2      & -0.176' & 0.007' & -0.189' & -0.187' & -0.181' & -0.176' & -0.168' & -0.166' & -1.642462e-01 \\
+mu\_mon3      & -0.081' & 0.007' & -0.094' & -0.093' & -0.086' & -0.082' & -0.074' & -0.071' & -6.978557e-02 \\
+mu\_mon4      & -0.035' & 0.007' & -0.048' & -0.047' & -0.040' & -0.035' & -0.028' & -0.025' & -2.350794e-02 \\
+mu\_mon5      & -0.010' & 0.007' & -0.023' & -0.022' & -0.015' & -0.011' & -0.003' & -0.000' & -1.024029e-05 \\
+mu\_mon6      & -0.013' & 0.007' & -0.026' & -0.025' & -0.018' & -0.014' & -0.006' & -0.003' & -1.735978e-03 \\
+mu\_mon7      & -0.020' & 0.007' & -0.034' & -0.031' & -0.025' & -0.021' & -0.013' & -0.010' & -8.326959e-03 \\
+mu\_mon8      & -0.039' & 0.007' & -0.053' & -0.051' & -0.045' & -0.040' & -0.032' & -0.030' & -2.838155e-02 \\
+mu\_mon9      & -0.096' & 0.007' & -0.109' & -0.107' & -0.101' & -0.096' & -0.089' & -0.086' & -8.434625e-02 \\
+mu\_mon10     & -0.152' & 0.007' & -0.166' & -0.164' & -0.158' & -0.153' & -0.145' & -0.143' & -1.412385e-01 \\
+mu\_mon11     & -0.245' & 0.007' & -0.258' & -0.256' & -0.250' & -0.245' & -0.237' & -0.235' & -2.336429e-01 \\
+mu1\_lease0   & -0.005' & 0.001' & -0.008' & -0.007' & -0.006' & -0.005' & -0.005' & -0.004' & -2.602403e-03 \\
+mu1\_lease1   & -0.000' & 0.000' & -0.002' & -0.001' & -0.000' & -0.000' & -0.000' & -0.000' & -5.518242e-08 \\
+mu1\_s0       & -0.002' & 0.001' & -0.005' & -0.004' & -0.003' & -0.002' & -0.002' & -0.001' & -1.790787e-04 \\
+mu1\_s1       & -0.001' & 0.000' & -0.003' & -0.002' & -0.001' & -0.000' & -0.000' & -0.000' & -2.355766e-06 \\
+mu1\_s2       & -0.001' & 0.001' & -0.004' & -0.002' & -0.001' & -0.001' & -0.000' & -0.000' & -1.807113e-06 \\
+mu1\_s3       & -0.007' & 0.002' & -0.014' & -0.010' & -0.008' & -0.007' & -0.006' & -0.004' & -2.390480e-03 \\
+beta1\_cost0  & -0.000' & 0.002' & -0.005' & -0.003' & -0.001' & -0.000' &  0.001' &  0.003' &  4.742583e-03 \\
+beta1\_cost1  &  0.001' & 0.001' & -0.002' & -0.001' &  0.000' &  0.001' &  0.001' &  0.002' &  3.464386e-03 \\
+beta1\_size0  &  0.001' & 0.001' & -0.002' & -0.000' &  0.001' &  0.001' &  0.002' &  0.003' &  3.930082e-03 \\
+\bottomrule
+\end{tabular}
+\label{table:higher_level_params_const}
+\caption{Summary statistics of the estimated posterior distributions of higher level parameters when higher level mean parameters are constrained to have a maximum of zero.}
+\end{table}
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=1\linewidth]{figures/mu_sectors.png}
+%  \caption{Distribution of the sector contrasts relative to residential sector.
+Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
+%  \label{mu_sectors}
+%\end{figure}
+
+\begin{figure} [!htb]
+\begin{minipage}{.48\textwidth}
   \centering
-  \includegraphics[width=.6\linewidth]{figures/mu_sectors.png}
+  \includegraphics[width=1\linewidth]{figures/mu_sectors.png}
+  \label{fig:mu_sectors}
+ \end{minipage}\qquad
+\begin{minipage}{.48\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/mu_sectors_const.png}
+  \label{fig:mu_sectors_const}
+ \end{minipage}
+
+\bigskip
+
+\begin{minipage}[t]{.48\textwidth}
+\centering
  \caption{Distribution of the sector contrasts relative to residential sector.
 Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
-  \label{mu_sectors}
+\end{minipage}\qquad
+\begin{minipage}[t]{.48\textwidth}
+\centering
+ \caption{Distribution of the sector contrasts relative to residential sector when all higher level mean parameters are constrained to have a maximum of zero.
+Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
+\end{minipage}
 \end{figure}
 
-Summary statistics for the parameters on the manufacturer group means as well as the lower-level intercept and slope paramaters $a_s$ and $b_s$ are too numerous to present in a table.
-Visual summaries are presented in figures \ref{sfig:BayManPlot}, \ref{sfig:Bay_as} and \ref{sfig:Bay_bs}.
-Figure \ref{sfig:BayManPlot} shows box plots of the manufacturer distributions, with the box representing 70 percent of probability and the lines representing 95 percent of the probability.
-Mirroring earlier results, the plot shows most manufacturers with roughly equivalent quality, with handful with meaningfully lower quality.
-Figures \ref{sfig:Bay_as} and \ref{sfig:Bay_bs} show 95\% intervals of the distribution of the $a_s$ and $b_s$ parameters.
+Summary statistics for the parameters on the manufacturer group means as well as the lower-level intercept and slope parameters $a_s$ and $b_s$ are too numerous to present in a table.
+Visual summaries are presented in figures \ref{fig:BayManPlot}, \ref{fig:Bay_as} and \ref{fig:Bay_bs}.
+Figure \ref{fig:BayManPlot} shows box plots of the manufacturer distributions, with the box representing 70 percent of probability and the lines representing 95 percent of the probability.
+Mirroring earlier results, the plot shows most manufacturers with roughly equivalent quality, with a handful with meaningfully lower quality.
+Figures \ref{fig:Bay_as} and \ref{fig:Bay_bs} show 95\% intervals of the distribution of the $a_s$ and $b_s$ parameters.
 The intercept terms are shown to have considerable variation between systems, as would be expected, but are individually estimated precisely.
 
-The slope parameters, $b_s$ are modelled explicitly and are also allowed to have varying variance parameters.
+The slope parameters, $b_s$ are modeled explicitly and are also allowed to have varying variance parameters.
 This is evident in the plot.
@@
 
-\begin{figure}[!htb]
-\begin{minipage}{.48\textwidth}
+\begin{figure}
+\begin{minipage}{.45\textwidth}
+  \centering
   \includegraphics[width=1\linewidth]{figures/BayManPlot.png}
+  \label{fig:BayManPlot}
+ \end{minipage}\qquad
+\begin{minipage}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/BayManPlot_const.png}
+  \label{fig:BayManPlot_const}
+ \end{minipage}
+
+\bigskip
+
+\begin{minipage}[t]{.45\textwidth}
+\centering
   \caption{Box plot of the the parameters $\mu^{manuf}$, representing the distributions of the mean parameter for manufacturer groupings.}
-  \label{sfig:BayManPlot}
-\end{minipage}\hfill
-\begin{minipage}{.48\textwidth}
+\end{minipage}\qquad
+\begin{minipage}[t]{.45\textwidth}
+\centering
+  \caption{Box plot of the the parameters $\mu^{manuf}$, when all higher level mean parameters have a maximum of zero.}
+\end{minipage}
+\end{figure}
+
+%new image
+
+\begin{figure}
+\begin{minipage}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/BayCountyPlot.png}
+	\label{fig:BayCountyPlot}
+ \end{minipage}\qquad
+\begin{minipage}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/BayCountyPlot_const.png}
+  \label{fig:BayCountyPlot_const}
+ \end{minipage}
+
+\bigskip
+
+\begin{minipage}[t]{.45\textwidth}
+\centering
+  \caption{Box plot of the the parameters $\mu^{county}$, representing the distributions of the mean parameter for county groupings.}
+\end{minipage}\qquad
+\begin{minipage}[t]{.45\textwidth}
+\centering
+  \caption{Box plot of the the parameters $\mu^{county}$, when all higher level mean parameters have a maximum of zero.}
+\end{minipage}
+\end{figure}
+
+% new figure
+
+\begin{figure}
+\begin{minipage}{.45\textwidth}
+  \centering
   \includegraphics[width=1\linewidth]{figures/Bay_as.png}
+  \label{sfig:Bay_as}
+ \end{minipage}\qquad
+\begin{minipage}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/Bay_as_const.png}
+  \label{sfig:Bay_as_const}
+ \end{minipage}
+
+\bigskip
+
+\begin{minipage}[t]{.45\textwidth}
+\centering
   \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$.}
-  \label{sfig:Bay_as}
-\end{minipage}\hfill
-\begin{minipage}{.48\textwidth}
+\end{minipage}\qquad
+\begin{minipage}[t]{.45\textwidth}
+\centering
+  \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$ when higher level mean parameters are constrained to have a maximum of zero.}
+\end{minipage}
+\end{figure}
+
+% new figure
+
+\begin{figure}
+\begin{minipage}{.45\textwidth}
+  \centering
   \includegraphics[width=1\linewidth]{figures/Bay_bs.png}
+  \label{sfig:Bay_bs}
+ \end{minipage}\qquad
+\begin{minipage}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{figures/Bay_bs_const.png}
+  \label{sfig:Bay_bs_const}
+ \end{minipage}
+
+\bigskip
+
+\begin{minipage}[t]{.45\textwidth}
+\centering
  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
-  \label{sfig:Bay_bs}
-\end{minipage}\hfill
+\end{minipage}\qquad
+\begin{minipage}[t]{.45\textwidth}
+\centering
+  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
+\end{minipage}
+\end{figure}
+
+%\begin{figure}[!htb]
+%\begin{minipage}{.48\textwidth}
+%  \includegraphics[width=1\linewidth]{figures/BayManPlot.png}
+%  \caption{Box plot of the the parameters $\mu^{manuf}$, representing the distributions of %the mean parameter for manufacturer groupings.}
+%  \label{sfig:BayManPlot}
+%\end{minipage}\hfill
+%\begin{minipage}{.48\textwidth}
+%  \includegraphics[width=1\linewidth]{figures/BayCountyPlot.png}
+%  \caption{Box plot of the the parameters $\mu^{county}$, representing the distributions of the mean parameter for county groupings.}
+%  \label{sfig:BayCountyPlot}
+%\end{minipage}\hfill
+%\begin{minipage}{.48\textwidth}
+%  \includegraphics[width=1\linewidth]{figures/Bay_as.png}
+%  \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$.}
+%  \label{sfig:Bay_as}
+%\end{minipage}\hfill
+%\begin{minipage}{.48\textwidth}
+%  \includegraphics[width=1\linewidth]{figures/Bay_bs.png}
+%  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
+%  \label{sfig:Bay_bs}
+%\end{minipage}\hfill
 %\caption{Simulated distribution of coefficients}
 %\label{fig:coeff_sim}
-\end{figure}
+%\end{figure}
 
@@
 
-Because of the dramatic price declines in solar panels and the competitive pressure that western producers have faced, it has become common to hear in the industry that solar panels have become commoditised.
+Because of the dramatic price declines in solar panels and the competitive pressure that western producers have faced, it has become common to hear in the industry that solar panels have become commodities.
 But in finance and economics, commodity has a specific meaning.
@@
 
-Historically, the commoditisation of goods was a necessary condition for the establishment of sophisticated financial markets for those goods.
-The invention of grain elevators for mixing and then loading of grain onto trains in Chicago, was a necessary precursor to the establishment of a futures markets with receipts of delivery traded on the Chicago Board of Trade.\citep{cronon_natures_1992} The abstraction between producer and end purchaser of goods was necessary for the functioning of these markets.
-A trader could fulfil a promise to deliver grain of a certain grade by simply buying it on the market or with cash settlement, rather than having produced it oneself.
+Historically, the commoditization of goods was a necessary condition for the establishment of sophisticated financial markets for those goods.
+The invention of grain elevators for mixing and thereby standardizing grain was a necessary precursor to the establishment of a futures markets with receipts of delivery traded on the Chicago Board of Trade \citep{cronon_natures_1992}.
+The abstraction between producer and end purchaser of goods was necessary for the functioning of these markets.
+A trader could fulfill a promise to deliver grain of a certain grade by simply buying it on the market or with cash settlement, rather than having produced it oneself.
 
- The point of this discussion is that if solar panels truly have become commodities, it would have meaningful and substantial implications for the solar power industry.
+If solar panels truly have become commodities, it would have meaningful and substantial implications for the solar power industry.
 For example, financial instruments could arise providing developers of solar power plants or solar panel retailers the ability to hedge price uncertainty in panels and guarantee supplies for projects with long lead times.
@@
 In particular, I find that the salient property of quality varies substantially between producers.
-Because solar power is a distributed technology, where assets can be owned by individual homeowners, businesses and other other organisations that do not have the resources to judge quality, the issue of asymmetric information may then enter the market.
+Because solar power is a distributed technology, where assets can be owned by individual homeowners, businesses and other other organizations that do not have the resources to judge quality, the issue of asymmetric information may then enter the market.
 
@@
 The market for solar panels and solar panel systems is however still young and dynamic.
-One of the advantages of the linear mixed-effects models and the multilevel models is that we can also get an idea of the degree of differentiation in quality between different manufacturers.
-It appears that a relatively small tail of producers accounted for the worst quality panels, while the majority of panels had
+One of the advantages of the linear mixed-effects models and the hierarchical models is that we can also get an idea of the degree of differentiation in quality between different manufacturers.
+It appears that a relatively small tail of producers accounted for the worst quality panels, while the majority of panels had similar quality.
+The market for solar panels is still relatively young, and if low-quality producers either are forced to exit the market or improve their quality to match standards in the industry, then it is possible that solar panels could indeed become commodities in the future, with associated implications for pricing, competition and financing.
 
-\begin{spacing}{1}
 \bibliographystyle{chicago}
@@
 
-\end{spacing}
 \end{document}

### 2018-06-01 data_checks.py

@@ -1,0 +1,65 @@
+#data_checks.py
+#check which series have positive slopes
+
+import pandas as pd
+import tabulate as tab
+import numpy as np
+import matplotlib as mpt
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import seaborn as sns
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+import sys
+import pystan
+import scipy
+import pickle
+
+plt.rcParams['xtick.labelsize'] = 14
+plt.rcParams['ytick.labelsize'] = 14
+plt.rcParams['axes.facecolor'] ="white"
+plt.rcParams['grid.color'] ="grey"
+plt.rcParams['grid.linestyle'] = "dotted"
+plt.rcParams["axes.labelsize"]= 14
+#plt.rcParams['figure.savefig.dpi'] = 100
+plt.rcParams['savefig.edgecolor'] = "#f2f2f2"
+plt.rcParams['savefig.facecolor'] ="white"
+plt.rcParams["figure.figsize"] = [15,8]
+plt.rcParams['savefig.bbox'] = "tight"
+plt.rcParams['font.size'] = 14
+greens = ['#66c2a4','#41ae76','#238b45','#006d2c','#00441b']
+multi =['#66c2a4','#1f78b4','#a6cee3','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f']
+
+
+#check with systems have positive estimated slope
+solar_data = pd.read_csv("data/prod_csi.csv")
+
+solar_data.columns.values
+
+solar_data.total_months_operation.head(100)
+test_ser = solar_data.loc[solar_data.app_num=="SD-CSI-00001", :]
+
+
+def slope_est(df):
+    #df= test_ser
+    res = smf.ols("prod_scaled~months_operation", data=df).fit()
+    b = res.params[1]
+    return(b)
+
+b_df = solar_data.groupby("app_num").apply(slope_est)
+
+
+b_df.head()
+
+
+b_df[b_df>0.01].size
+
+posit_b = b_df.index[b_df>0.04]
+
+posit_df = solar_data.loc[solar_data.app_num.isin(posit_b),:]
+
+grouped_df = posit_df.groupby("app_num")
+fig, ax = plt.subplots()
+for df in grouped_df:
+    ax.plot(df[1].months_operation, df[1].prod_scaled)
+plt.show()

### 2018-06-01 lmel_modelling.r

@@ -12,7 +12,6 @@ load_libraries = function(){
   library(mgcv)
   library(arm)
   library(stargazer)
-
 }
 
 #initial functions
@@ -167,11 +166,19 @@ main = function(){
   prod_fig = prod_fig + geom_abline(intercept=lm_coef[1], slope=lm_coef[2], color="red") + xlim(0,60) + ylim(-1.5,1.5) + labs(x="months",y="Production, de-scaled")
 
   ggsave("figures/production.png", plot=prod_fig, dpi=100,width = 9, height = 5)
+<<<<<<< HEAD
 
   #sum summary statistics
   solar_data[c("prod_kwh", "first_prod_year", "cost_per_kw", "csi_rating")]
   table(solar_data["third_party_owned"])
 
+=======
+
+  #sum summary statistics
+  solar_data[c("prod_kwh", "first_prod_year", "cost_per_kw", "csi_rating")]
+  table(solar_data["third_party_owned"])
+
+>>>>>>> df0de029a7615344b408ad311170086016b230a0
 
   59108/(111919+59108)
 

### 2018-06-01 stan_analysis_new.py

@@ -1,5 +1,9 @@
 #stan_analysis_new.py
+<<<<<<< HEAD
+#stan MCMC analysis from 6.2.2018
+=======
 #stan MCMC analysis from 2.6.2017
+>>>>>>> df0de029a7615344b408ad311170086016b230a0
 
 import pandas as pd
 import tabulate as tab

### 2019-03-29 ejor_submission/R&R/replyToEditor.txt
update RR

Reply to editor

As you can see from the reports enclosed, the referees hold conflicting views on your paper. In these circumstances, it has been decided that your paper should be reconsidered for publication in the EUROPEAN JOURNAL OF OPERATIONAL RESEARCH if you will revise it to meet the enclosed referees' comments.

The most critical comments made me doubt whether I should reconsider your paper, so I need convincing arguments from you and a thorough revision to counterbalance the critical comments.

When resubmitting the revised paper, please provide your answers to the referees, separately for each referee.  Please also ensure that you have cited recent and relevant publications in EJOR and other OR journals.

I am most grateful for your interest and hope that you will not be deterred from resubmitting this paper in the very near future.

You are advised to revise and resubmit your manuscript by the Sep 19, 2019. After this time the review process of this submission will be terminated, by withdrawing your paper from the Elsevier Editorial System, and the revised paper will have to be submitted as a new submission.

### 2019-03-29 solar_prod_220518.tex
update RR

@@
 Solar panels should not be considered commodities.
-Considerable quality differences, as measured directly by degradation of production over time, are found between manufacturers.
-This has implications for pricing and competition in the market for solar panel systems.
-I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts.
-2.) Furthermore, the price of solar panels that are owned by high-information owners are more highly correlated to quality.
-I use random effects models estimated by maximum likelihood and hierarchical models estimated by Bayesian Markov Chain Monte-Carlo.
+Quality differences, as measured directly by degradation of production over time, are found between manufacturers.
+I test implications from the theory of asymmetric information of quality and find evidence that: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts.
+2.) The price of solar panels that are owned by high-information owners are more highly correlated to quality.
+I use random effects models estimated by maximum likelihood and Bayesian hierarchical models estimated by Markov Chain Monte-Carlo.
 \end{abstract}

### 2019-03-29 solar_prod.tex
update

@@
 The energy investment behavior of home owners, farmers, and small cooperatives are bound to be substantially different than those of large, specialized energy companies that have traditionally done most of the investment in electricity generation.
-Understanding the quickly evolving energy and power industry requires taking into account informational and behavioral factors, and this is a rich field for further empirical research
+Understanding the quickly evolving energy and power industry requires taking into account informational and behavioral factors, and this is a rich field for further research
 

### 2019-04-10 ejor_submission/R&R/replyToEditor.txt
revision

@@
-Reply to editor
+Dear Prof. Borgonovo,
 
-As you can see from the reports enclosed, the referees hold conflicting views on your paper.
-In these circumstances, it has been decided that your paper should be reconsidered for publication in the EUROPEAN JOURNAL OF OPERATIONAL RESEARCH if you will revise it to meet the enclosed referees' comments.
+Thank you for the opportunity to revise and resubmit my article, "Are Solar Panels Commodities?" to the European Journal of Operational Research.
 
-The most critical comments made me doubt whether I should reconsider your paper, so I need convincing arguments from you and a thorough revision to counterbalance the critical comments.
+I have done a thorough revision based on the comments of the three referees.
+Attached you will find detailed, point-by-point replies to each of the reviewer's suggestions and critiques in addition to the revised article.
 
-When resubmitting the revised paper, please provide your answers to the referees, separately for each referee.
-Please also ensure that you have cited recent and relevant publications in EJOR and other OR journals.
+I wish to highlight a few of the main changes to the article.
+As you mentioned in your letter to me, referee 2 was particularly critical and felt that the article lacked a decision-making perspective.
+In addition, the referee felt a lack of methodological novelty.
+I have taken the referee's critique to heart, and agree that the initial submission did not do an adequate job of conveying the article's relevance to decision makers nor did it sufficiently convey the articles methodological novelty.
+A substantial part of the revision consists of righting these weaknesses.
 
-I am most grateful for your interest and hope that you will not be deterred from resubmitting this paper in the very near future.
+Referee 2 is correct in suggesting that the article could, thematically, also fit into a pure economics or finance journal.
+But I chose to submit to a top operations research journal, in part, because of the practical implications of the research to decision makers.
+As an example of how I have tried to convey this in the revision, I write on p.2 :
 
-You are advised to revise and resubmit your manuscript by the Sep 19, 2019.
-After this time the review process of this submission will be terminated, by withdrawing your paper from the Elsevier Editorial System, and the revised paper will have to be submitted as a new submission.
+  From the perspective of a purchasing manager responsible for procuring solar panels, whether panels can be considered a commodity is of great importance.
+If panels are known to be commodities with standardised quality and similar pricing, then the decision for purchasing manager is substantially simplified.
+They need only to decide on the observable form factor, for example dimension and capacity.
+In addition, purchasing managers responsible for large projects with long lead times could make use of forward markets to hedge price movements.
+On the other hand, if solar panels are known to vary significantly by quality, then the decision becomes substantially more complex and costly.
+Purchasing managers would need to weigh the price of panels against expected quality.
+They may also need to make use of specialised engineering firms that test and certify panel quality.
+Since the purchasing decision is likely to be linked to a specific manufacturer, the purchasing manager is subject to risks such as bankruptcy or other supply disruption associated with that firm.
+
+The methodological focus of operations research is another reason I would prefer EJOR as an outlet over a finance or economics journal.
+Bayesian methods are generally rare in economics and finance, apart from a few small subfields.
+Hierarchical Bayesian models are even more sparse.
+On the other hand, hierarchical Bayesian models have recently become prominent in the operations research field, and in EJOR in particular.
+In the revision I have worked at highlighting the relative novelty of the hierarchical Bayesian model I use, as well as presenting a more extensive review of the recent applications of hierarchical Bayes models in the operations research field.
+For example, on pages 4-5 I write:
+
+  ...
+I use a Bayesian hierarchical model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
+Bayesian hierarchical models have long been one of the primary tools in statistical decision theory (Berger, 1985).
+This form of modelling is increasingly common in operations research and management science.
+One of the main advantages is superior out-of-sample predictive properties with data with distinct groupings.
+Recent advances in computation have also allowed for the construction and simulation of increasingly complex models.
+In particular, I use the Stan probabilistic programming language (Stan Development Team, 2014) that makes use of Hamiltonian Markov Chain Monte Carlo (MCMC) (MacKay, 2003)and a No-U-Turn sampler (Homan and Gelman, 2013).
+Unlike earlier MCMC simulation routines and software, notably Gibbs sampler, Hamiltonian MCMC is able to reach convergence for models, like the one I present in this article, with several hierarchies.
+Importantly, it is not necessary to use conjugate distributions, opening up to a wider variety choice of prior distributions (Kruschke, 2014).
+
+Later I write:
+
+  In part due to increasing computational tractability, Bayesian hierarchical models have recently become a common and useful tool in the operations research field.
+A common application is to incorporate expert opinion into quantitative decision models (Rufo et al., 2010; Szwed et al., 2006; Utkin, 2006).
+In this application, the Bayesian hierarchical model is flexible enough to aggregate the information of the expert opinions while also being able to explicitly handle and propagate the uncertainty in these opinions.
+Hierarchical models have also been used to identify default risk among lenders and improve credit scoring models (Liu et al., 2015; Bhattacharya et al., 2019).
+In this context, the hierarchical Bayesian models are found to provide superior out-of-sample prediction for the censored sample of the data.
+Other operations research related applications include a fund manager's portfolio selection problem (Soyer and Tanyeri, 2006; Bodnar et al., 2017) and conjoint analysis in operations management (Karniouchina et al., 2009; Gensler et al., 2012).
+A commonality among all these applications is the need to aggregate information from heterogenous groups without over-fitting the data with many fixed effects.
+
+I am happy to address any further questions or issues you or the referees should have with the revised paper. I thank you for your time and consideration.
+
+Best wishes,
+
+Johannes Mauritzen

### 2019-07-02 ejor_submission/R&R/revision_mauritzenApril19.tex
Final push after acceptance

--- 2018-06-01 solar_prod_220518.tex
@@
 
-\title{Are Solar Panels Commodities? Evidence of Quality Differences and Asymmetric Information from California}
+%All floats at end.
+% \usepackage[nomarkers]{endfloat}
+
+%keywords
+\providecommand{\keywords}[1]
+{
+  \small
+  \textbf{\textit{Keywords---}} #1
+}
+
+\title{Are Solar Panels Commodities? A Bayesian Hierarchical Approach to Detecting Quality Differences and Asymmetric Information.}
 \author{Johannes Mauritzen\thanks{BI Norwegian Business School, johannes.mauritzen@bi.no, \url{http://jmaurit.github.io}}}
@@
 Considerable quality differences, as measured directly by degradation of production over time, are found between manufacturers.
-This has implications for pricing and competition in the market for solar panel systems.
-I test two implications from the the theory of asymmetric information of quality and find: 1.) Solar systems with high-information third party owners display higher quality than those owned by low-information hosts.
-2.) Furthermore, the price of solar panels that are owned by high-information owners are more highly correlated to quality.
-I use random effects models estimated by maximum likelihood and hierarchical models estimated by Bayesian Markov Chain Monte-Carlo.
+I test two implications from the theory of asymmetric information of quality and find: 1.) Solar power systems with high-information third-party owners display higher quality than host-owned systems.
+2.) Furthermore, with a 85\% probability, the price of solar panels that are owned by high-information owners are more highly correlated to quality.
+Methodologically, the article demonstrates a novel application of Bayesian hierarchical regression models that are increasingly popular in operations research and the decision sciences.
 \end{abstract}
+
+%TC:ignore
+\keywords{Solar power, commodities, asymmetric information, quality, Bayesian hierarchical model}
 
@@
 That is, the good from a certain producer will be largely interchangeable with the good of another producer.
-Brent crude oil produced by Statoil from the Norwegian North Sea and that produced by BP from the UK North Sea are sold at the same price, as established by an exchange.
-In a similar manner, the electricity produced by a solar panel at a certain location and time can also be considered a commodity.
+Already in 1873 the German economist A.E.F.
+Schäffle distinguished between a popular understanding of a commodity as a material input in manufacturing, to the scientific understanding as "exchangeable material goods" \citep{menger_principles_2007}.
+Despite the simplicity of this definition, \citet{gordon_what_1999} show the difficulty in practice of determining whether a product is in fact a commodity.
+The root problem is that a good can be differentiated in a host of ways, some of which are difficult to observe at the time of transaction.
+
+Determining whether a product is a commodity is of more than just theoretical importance.
+Historically, the commodification of goods was a necessary condition for the establishment of sophisticated financial markets for those goods.
+A well-documented example: The invention of grain elevators for mixing and standardising grain was a necessary precursor to the establishment of a futures markets with receipts of delivery traded on the Chicago Board of Trade \citep{cronon_natures_1992}.
+The abstraction between producer and end purchaser of goods was necessary for the functioning of these markets.
+A trader could fulfil a promise to deliver grain of a certain grade by simply buying it on the market or with cash settlement, rather than having produced it oneself.
 
@@
 
+From the perspective of a purchasing manager responsible for procuring solar panels, whether panels can be considered a commodity is of great importance.
+If panels are known to be commodities with standardised quality and similar pricing, then the decision for purchasing manager is substantially simplified.
+They need only to decide on the observable form factor, for example dimension and capacity.
+In addition, purchasing managers responsible for large projects with long lead times could make use of forward markets to hedge price movements.
+On the other hand, if solar panels are known to vary significantly by quality, then the decision becomes substantially more complex and costly.
+Purchasing managers would need to weigh the price of panels against expected quality.
+They may also need to make use of specialised engineering firms that test and certify panel quality.
+Since the purchasing decision is likely to be linked to a specific manufacturer, the purchasing manager is subject to risks such as bankruptcy or other supply disruption associated with that firm.
+
 Furthermore, if significant quality differences between manufacturers exists, then the characteristics of solar panel investments suggest the possibility of asymmetric information of quality in the market.
+This is important in practice, since the presence of information asymmetry in a market tends to lead to the persistence of low quality suppliers.
 
@@
 The market for rooftop solar panels can be expected to be particularly vulnerable to issues of asymmetric information on quality.
-Solar panels can be characterized as an ``experience'' good, where an investor/consumer needs to learn about the quality through use.
+Solar panels can be characterised as an ``experience'' good, where an investor/consumer needs to learn about the quality through use \citep{caulkins_interaction_2017}.
 In particular, poor quality panels will tend to show a higher degradation of output over time than high quality panels.
@@
 The quality of panels, as measured by degradation of output over time, would be predicted to be higher in the systems owned by the high-information owners (third-party-owner) who can discern between low and high quality solar panels.
-Conversely, the basic theory of asymmetric information suggests that lower quality panels will tend to push out higher quality panels in the presence of low-information buyers who can not discern quality, and therefor make purchasing decision solely on price.
+Conversely, the basic theory of asymmetric information suggests that lower quality panels will tend to push out higher quality panels in the presence of low-information buyers who cannot discern quality, and therefor make purchasing decision solely on price.
 
 A related prediction from the theory of asymmetric information of quality is that the cost of solar panel systems should have a higher degree of correlation with quality when the owner is a high-information type.
-Because a high-information owner can, by definition, differentiate between high and low quality panels.
+Because a high-information owner can, by definition, differentiate between high- and low-quality panels.
 
@@
 
-I have three nested predictions that I wish to test:
+In summary, I have three nested research hypothesis that I wish to test empirically.
 
@@
 Because these groupings display considerable heterogeneity in sample size, mean values and variance, it is natural to model the relevant structure of the data directly.
-For relatively parsimonious models with one or two hierarchies and few covariates, I can use linear mixed effects models (LMEM) estimated by maximum likelihood.
+For relatively parsimonious models with one or two hierarchies and few covariates, I can use standard linear mixed effects models (LMEM) estimated by maximum likelihood.
 However, for richer models with all relevant hierarchies, maximum likelihood becomes impractical.
 
-As a solution, I use a Bayesian hierarchical\footnote{Sometimes also called multilevel modeling or random effects modeling} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
-This form of modeling has long been established in multivariate dynamic modeling in macroeconomics \citep{doan_forecasting_1984, litterman_forecasting_1986,sims_bayesian_1998, canova_forecasting_2004},  but is also increasingly common in microeconomics and corporate finance.
-One of the main advantages is superior out of sample predictive properties with data with distinct groupings.
-For example \citet{dehejia_was_2003} uses a Bayesian hierarchical model to evaluate multiple programs at different sites.
-By incorporating ``predictive uncertainty'' by way of a hierarchical model, the impact of the sites, which otherwise is estimated to be significant with a fixed-effects model, disappears.
-\citet{meager_understanding_2018} presents a more recent application in the microeconomics litterature.
-She uses a hierarchical model to aggregate studies of microcredit expansion in developing countries in order to estimate an overall treatment effect with ``external validity,'' that is the degree to which an estimated treatment effect generalizes to other populations.
+%Berger - statistical decision theory.
 
-Within the field of corporate finance, \citet{jones_mutual_2005} uses a hierarchical model to study returns of mutual funds, arguing that estimates of long-run returns of any single fund should be informed, by way of a higher-level distribution, of the returns of the population of other mutual funds.
+As a solution, I use a Bayesian hierarchical\footnote{Sometimes also called multilevel modell} model estimated using Markov Chain Monte Carlo (MCMC) simulation techniques.
+Bayesian hierarchical models have long been one of the primary tools in statistical decision theory \citep{berger_statistical_1985} and this form of modelling is increasingly common in operations research and management science.
+One of the main advantages is superior out-of-sample predictive properties with data with distinct groupings.
+Recent advances in computation have also allowed for the construction and simulation of increasingly complex models.
+In particular, I use the Stan probabilistic programming language \citep{stan_development_team_stan_2014} that makes use of Hamiltonian Markov Chain Monte Carlo (MCMC) \citep{mackay_information_2003} and a No-U-Turn sampler \citep{homan_no-u-turn_2014}.
+Unlike earlier MCMC simulation routines, notably Gibbs sampler, Hamiltonian MCMC is able to reach convergence even with models, like the one I present in this article, with multiple hierarchies.
+Importantly, with the Hamiltonian MCMC routine, it is not necessary to use conjugate distributions, opening up to a wider variety of choice of prior distributions \citep{kruschke_doing_2014}.
+
+In part due to increasing computational tractability, Bayesian hierarchical models have recently become a common and useful tool in the operations research field.
+A common application is to incorporate expert opinion into quantitative decision models \citep{rufo_merging_2010, szwed_bayesian_2006, utkin_method_2006}.
+In this application, the Bayesian hierarchical model is flexible enough to aggregate the information of the expert opinions while also being able to explicitly handle and propagate the uncertainty in these opinions.
+Hierarchical models have also been used to identify default risk among lenders and improve credit scoring models \citep{liu_identifying_2015, bhattacharya_bayesian_2019}.
+In this context, the hierarchical Bayesian models are found to provide superior out-of-sample prediction for the censored sample of the data.
+Other operations research related applications include a fund manager's portfolio selection problem \citep{soyer_bayesian_2006, bodnar_bayesian_2017} and conjoint analysis in operations management \citep{karniouchina_issues_2009, gensler_willingness--pay_2012}.
+A commonality among all these applications is the need to aggregate information from heterogenous groups without over-fitting the data with many fixed effects.
+
+Despite the usefulness of hierarchical Bayesian models, examples of hierarchical models in the economics and finance literature are sparse.
+\citet{dehejia_was_2003} uses a Bayesian hierarchical model to evaluate the effectiveness at multiple programs at different sites.
+By incorporating ``predictive uncertainty'' by way of a hierarchical model, the impact of the sites, which otherwise is estimated to be significant with a fixed effects model, disappears.
+\citet{meager_understanding_2018} uses a hierarchical model to aggregate studies of micro-credit expansion in developing countries in order to estimate an overall treatment effect with ``external validity,'' that is the degree to which an estimated treatment effect generalizes to other populations.
+\citet{jones_mutual_2005} uses a hierarchical model to study returns of mutual funds, arguing that estimates of long-run returns of any single fund should be informed, by way of a higher-level distribution, of the returns of the population of other mutual funds.
 In a study with parallels to my own, \citet{korteweg_skill_2017} uses a Bayesian hierarchical model to study persistence of returns in the private equity sector both between and within firms, where persistence is measured by excess variation between firms as captured by firm-level parameters (or random effects) that are drawn from a higher-level distribution.
@@
 The property of hierarchical models that allows for improved predictive performance with the presence of groupings in the data is regularization, or alternatively called partial pooling \citep{gelman_bayesian_2013}.
-As opposed to fixed-effects estimation where a fixed-effects coefficient is determined exclusively with data from within a particular group, in a hierarchical model each of the group parameters are regularized towards a higher-level parameter estimated with the full data in proportion to the relative number of data points in a particular group.
+As opposed to fixed effects estimation where a fixed effects coefficient is determined exclusively with data from within a particular group, in a hierarchical model each of the group parameters are regularized towards a higher-level parameter estimated with the full data in proportion to the relative number of data points in a particular group.
 
-The linear mixed effects models estimated by maximum likelihood that I use in the first section of this article also have have the regularization property.
+The linear mixed effects models estimated by maximum likelihood that I use in the first section of this article also have the regularization property.
 Bayesian estimation provides several theoretic advantages, such as correct inference in small samples and generalization to distributions other than the normal.
-However, for my purproses, the main benefit of using a Bayesian methodology is practical.
-I can fit a rich model with several hierarchies and many thousands of parameters, that simply can not be handled by maximum likelihood procedures.
+However, for my purposes, the main benefit of using a Bayesian methodology is practical.
+I can fit a rich model with several hierarchies and many thousands of parameters, that simply cannot be handled by maximum likelihood procedures.
 
-The issue of how information asymmetries of quality affects market outcomes, first introduced by \citet{akerlof_market_1970} is one of the main topics in the modern industrial organization literature.
+The issue of how information asymmetries of quality affects market outcomes was first introduced by \citet{akerlof_market_1970}.
 Ch.
@@
 Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller--in other words the seller can choose the level of quality of its product.
-Under these assumptions, the authors predict several equilibrium--essentially markets for both high and low quality products depending on the extent to which the buyer is informed.
+Under these assumptions, the authors predict several equilibrium--essentially markets for both high- and low-quality products depending on the extent to which the buyer is informed.
 
@@
 Most solar installers and panel manufacturers are relatively new firms, especially the increasingly dominant Chinese solar panel producers.
-\citet{shapiro_consumer_1982} explores the role of reputation, and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.
+\citet{shapiro_consumer_1982} explores the role of reputation and finds that because reputation is effective only with a lag, the "steady-state" quality level chosen by a seller should be lower than in the scenario with perfect information.
 
-An expanding empirical literatures exist on asymmetric information of quality, especially in the marketing, finance and accounting fields.
+An expanding empirical literature exists on asymmetric information of quality, especially in the marketing, finance and accounting fields.
 In accounting, the issue of firm quality and voluntary disclosure in regulatory filings as a signal has received broad attention by empirical researchers.
@@
 The literature in the marketing field is also relevant to this paper, as it is particularly concerned with the effects of information asymmetry on purchases of consumer durables.
-The effects of signaling on quality, usually in the form of advertising, is particularly important in the field.
+The effects of signalling on quality, usually in the form of advertising, is particularly important in the field.
 \citet{kirmani_no_2000} provide an overview of this literature.
 
-The use of leasing, discussed in this article, can be seen as a form of costly signaling about quality, however leasing plays more than just an informational role since it also transfers much of the risk of quality to the contractor.
+The use of leasing, discussed in this article, can be seen as a form of costly signalling about quality.
+However leasing plays more than just an informational role since it also transfers much of the risk of quality to the contractor.
 Traditionally, energy generation has been an activity undertaken by large, specialized firms with specialized knowledge and resources.
@@
 
-For example \citet{dastrup_understanding_2012} argue that solar panels can not be considered a pure investment good, but are also bundled as a type of green conspicuous consumption, showing evidence for a ``solar price premium'' in homes with solar panels.
+For example \citet{dastrup_understanding_2012} argue that solar panels cannot be considered a pure investment good, but are also bundled as a type of green conspicuous consumption, showing evidence for a ``solar price premium'' in homes with solar panels.
 \citet{bollinger_peer_2012} study the the role of peer effects in solar power adoption.
 They find evidence that the adoption of solar panels by homeowners in a certain zip-code will increase the probability that other households in that zip-code will install solar panels.
+A focus in the operations research field has been on designing systems to manage and aggregate such distributed energy resources on behalf of un-sophisticated owners \citep{velik_energy_2016, le_cadre_gametheoretic_2019}.
 
@@
 
-The data used in this article is an intersection of two datasets from the California Solar Initiative (CSI).
-The California Solar Initiative is a state-wide program that gives incentives for installing grid-connected solar panels systems.
+The data used in this article is an intersection of two data-sets from the California Solar Initiative (CSI).
+The California Solar Initiative is a state-wide program that gives incentives for installing grid-connected solar panel systems.
 The incentives are based on performance.
-For smaller systems, typical of most residential and small business installations, this incentive could be given in the form of an upfront payout based on the capacity and the expected performance of the system.
+For smaller systems, typical of most residential and small business installations, this incentive could be given in the form of an upfront pay-out based on the capacity and the expected performance of the system.
 
-Larger systems were required to accept a performance based incentive based on actual production over 5 years (60 months).
-From the beginning of the program in 2007 this was defined as those over 100kW, but was lowered to 50kW from January 2008 and 30kW from January 2010.
+Larger systems were required to accept a performance-based incentive based on actual production over 5 years (60 months).
+From the beginning of the program in 2007 this was defined as those over 100 kW, but was lowered to 50 kW from January 2008 and 30 kW from January 2010.
 Solar panel installations of all sizes have the option of getting an incentive based on actual performance.
@@
 \footnote{\url{http://www.californiasolarstatistics.ca.gov/current_data_files/}}.
-A cleaned and merged data set that I use in the following analysis can be found on my website at \url{jmaurit.github.io#buy_or_lease}.
+%A cleaned and merged data set that I use in the following analysis can be found on my website at \url{jmaurit.github.io#buy_or_lease}.
 
@@
 
-I removed systems from the dataset that may have had reporting errors.
-for example I removed systems where production was reported to be higher than what would be theoretically possible from a solar power system with a given nameplate capacity.
+I removed systems from the data that may have had reporting errors.
+For example, I removed systems where production was reported to be higher than what would be theoretically possible from a solar power system with a given nameplate capacity.
 I also removed systems that reported 0 production in a period.
 This could also be an indication of quality if zero production indicates a malfunction in the system.
-However it is not possible to identify which zero observations reflect malfunctions versus reporting issues or other non-quality issues.
+However, it is not possible to identify which zero observations reflect malfunctions versus reporting issues or other non-quality issues.
 
-Table \ref{tbl:variables} gives an overview of the variables in the dataset that are used in the analysis.
+Table \ref{tbl:variables} gives an overview of the variables in the data that are used in the analysis.
 Notice that variables can vary by monthly observation or by solar power system.
-I will take account of this natural data hierarchy in my modeling.
-There are a total of 171,027 monthly observations used in the analysis.
+I will take account of this natural data hierarchy in my modelling.
+A total of 171,027 monthly observations are used in the analysis.
 These observations come from 3,145 unique solar power systems.
@@
 Production is measured in kilowatt hours (kWh).
-In the analysis, production data is first log transformed and then de-scaled, as shown in equation \ref{eqn:descaled}.
-For production data for each system, $s$, the log production data is demeaned and then divided by 2 times the standard deviation of the production data for that solar power system.
+In the analysis, production data is de-scaled, as shown in equation \ref{eqn:descaled}.
+For production data for each system, $s$, the production data observations for each month, $i$, is demeaned and then divided by the standard deviation of the production data for that solar power system.
 Figure \ref{fig:production} gives a visualization of the transformed production data.
@@
 
-Other continuous variables are transformed in a similar manner, with observation level variables de-scaled at the system level, and system-level variables de-scaled across systems.
+Other continuous variables are transformed in a similar manner, with observation-level variables de-scaled at the system level, and system-level variables de-scaled across systems.
 These transformations avoid confusion that may come when interpreting coefficients on variables with different units and scales.
@@
 For systems with third party owners, the total cost is an estimate reported by the system owner based on the sum of component costs, installation costs and a mark-up.
-I will discuss how this may effect the results.
+I will discuss how this may affect the results.
 
@@
 
-Finally, among the 3000 plus solar power systems, 497 had residential owners, 1759 had commercial owners, 777 had governmental owners and 112 had non-profit owners.
+Finally, among the 3145 plus solar power systems, 497 had residential owners, 1759 had commercial owners, 777 had governmental owners and 112 had non-profit owners.
 A portion of the commercial owners operate as third-party-owners, with hosts in the other sectors.
 Alternatively, I could include host sector in my analysis: 1241 solar panel systems had a commercial host, 501 had a residential host, 1199 had a governmental host and 204 had a non-profit host.
-However, In a study of asymmetric information, it makes most sense to focus on the sector of the owner.
+However, in a study of asymmetric information, it makes most sense to focus on the sector of the owner.
 
@@
 Here $\mathbf{\beta}$ is a p-dimensional coefficient vector (or  ``fixed effects'') of the $n \times p$ matrix $\mathbf{X}$ of (p-1) explanatory variables plus intercept term.
-$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modeled directly with corresponding vector of coefficients, or ``random effects'' $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of ``random'' refers to the fact that these groups are assumed to be specific to the sample.
+$\mathbf{Z}$ is an $n \times q$ model matrix of the variables whose groups are to be modelled directly with corresponding vector of coefficients, or ``random effects'' $\mathbf{B}$ that are fixed at values $\mathbf{b}$.\footnote{The use of ``random'' refers to the fact that these groups are assumed to be specific to the sample.
 Many applications of mixed effects model thus treat the varying group coefficients as group-level error-terms.
-Interpreting coefficients as group varying coefficients or as group error terms is mathematically equivalent.} Here $q$ is a dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$ for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.
+Interpreting coefficients as group varying coefficients or as group error terms is mathematically equivalent.} Here $q$ is dependent on the number of terms, $k$, that are allowed to vary by group and the number of groups, $l$, for each term: $q=\sum_i^k q_i = \sum_i^k l_i p_i$.
 $\mathbf{W}$ is a diagonal matrix of known prior weights.
@@
 I write the restricted comparison model explicitly in equation \ref{eqn:observation_level_rest}.
+Symbols are defined in table \ref{tbl:definitions3p1}.
 
-In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modeled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$.
-Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by the individual system.
-Thus in this model formulation I take into account the groupings by system, explicitly modeling varying relative initial production values (intercepts) and degradation over time (slope).
-Moving up a level, the estimated $a$ and $b$ parameters are in turn modeled as multivariate normal.
+In this model, the transformed production variable, $prod\_scaled_{i}$ for observation $i$ is distributed normally with the mean modelled as a linear function of an intercept term $a_s$, a slope term, $b_s$ on the transformed months of operation variable $months\_operation\_scaled$, and a set of dummy variables representing the month of year, $\mathbf{month}$.
+Notice that the intercept and slope are indexed by $s$, indicating that they are allowed to vary by system.
+Thus, in this model formulation, I take into account the groupings by system, explicitly modelling varying relative initial production values (intercepts) and degradation over time (slope).
+Moving up a level, the estimated $a$ and $b$ parameters are in turn modelled as multivariate normal distributions.
+
+\begin{table}
+  \begin{tabular}{rl}
+  \toprule
+    Symbol &  Definition \\
+    \midrule
+    $prod_i$ & Production from a given system in a given month (observation i), descaled\\
+    $\mathbf{month}$ & Vector of month dummy variables\\
+    $months\_operation_i$ & Number of months of operation of a system at observation i, descaled \\
+    $\mathbf{a_s}$& Intercept parameter for a system, s \\
+    $\mathbf{b_s}$ & Slope parameter for a system, s \\
+    $\sigma^2$ & Variance of the production data \\
+    $\mu^a$ & Mean of the system intercept terms, $a_s$ \\
+    $\mu^b$ & Mean of the system slope terms, $b_s$ \\
+    $\sigma^{2,a}$ & Variance of the system intercept terms, $a_s$ \\
+    $\sigma^{2,b}$ & Variance of the system slope terms, $b_s$ \\
+    $\rho$ & Correlation coefficients between intercept and slope terms \\
+
+  \bottomrule
+  \end{tabular}
+  \caption{Definitions from the restricted model presented in equation \ref{eqn:observation_level_rest}. \label{tbl:definitions3p1}}
+\end{table}
 
@@
 \begin{aligned}
-prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation_{i}, \sigma^2)\\
 \begin{pmatrix}
-  a\\
-  b
+  a_s\\
+  b_s
 \end{pmatrix}
@@
 \begin{pmatrix}
-  \mu_a\\
-  \mu_b
+  \mu^a\\
+  \mu^b
 \end{pmatrix},
 \begin{pmatrix}
-  \sigma_a^2 & \rho \sigma_a \sigma_b \\
-  \rho \sigma_a \sigma_b & \sigma_b^2
+  \sigma_a^2 & \rho \sigma^a \sigma^b \\
+  \rho \sigma^a \sigma^b & \sigma^{2,b}
 \end{pmatrix}\label{eqn:observation_level_rest}
@@
 
-The inclusion of the matrix of month dummies, $\mathbf{\mu^{month}}$, controls for seasonal variation across all systems.
-This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and a more direct modeling of seasonality is not of direct importance, this should suffice.
-As long as un-modeled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.
+The inclusion of the matrix of month dummies, $\mathbf{month}$, controls for seasonal variation across all systems.
+This is a simplified way to model seasonality, but because the data is itself at a monthly frequency, and a more direct modelling of seasonality is not of direct importance, this should suffice.
+As long as un-modelled variation in seasonality is not correlated with the long-run slope of production, the point estimates for the slope should be unbiased.
 
@@
 In this model, the observation-level regression is identical to the restricted comparison model as in equations \ref{eqn:observation_level_rest}.
-The system-level intercepts, $a$, and slopes, $b$, terms are also modeled as multivariate normal.
-But the mean values of the multivariate normal distribution, $\mu_a$ and $\mu_b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.
+New symbols in this model are defined in table \ref{tbl:definitions3p2}.
+The system-level intercepts, $a$, and slopes, $b$, are also modelled as multivariate normal distributions.
+But the mean values of the multivariate normal distribution, $\mu^a$ and $\mu^b$, are now indexed by $m$, as shown in equation \ref{eqn:unrest}, representing that they are allowed to vary by module manufacturer.
+
+\begin{table}
+  \begin{tabular}{rl}
+  \toprule
+    Symbol &  Definition \\
+    \midrule
+    $\mu^a_m$ & Mean of the system intercept terms, $a_s$ by manufacturer, $m$ \\
+    $\mu^b_m$ & Mean of the system slope terms, $b_s$ by manufacturer, $m$ \\
+    $\mu^a_l$ & Mean of the system intercept terms, $a_s$ by ownership type, $l$ (leased/host-owned) \\
+    $\mu^b_l$ & Mean of the system slope terms, $b_s$ by ownership type, $l$ (leased/host-owned) \\
+    $\theta^a$ & Higher level (meta) mean of the distribution of grouped intercept means, $\mu^a_m$ \\
+    $\theta^b$ & Higher level (meta) mean of the distribution of grouped slope means, $\mu^b_m$ \\
+    $\sigma^{2, \mu^a_m}$ & Higher level (meta) variance of the distribution of grouped intercept means, $\mu^a_m$ \\
+    $\sigma^{2, \mu^b_m}$ & Higher level (meta) variance of the distribution of grouped slope means, $\mu^b_m$ \\
+
+  \bottomrule
+  \end{tabular}
+
+  \caption{Definitions from the unrestricted models presented in equation \ref{eqn:unrest}, with means that vary by manufacturer, and in equation \ref{eqn:fullmodel}, with means that vary by owner type (leased or host-owned).
+See table \ref{tbl:definitions3p1} for symbols that are identical in both the restricted and unrestricted models.
+\label{tbl:definitions3p2}}
+\end{table}
 
@@
 \begin{pmatrix}
-  \mu_{a[m]}\\
-  \mu_{b[m]}
+  \mu^a_m\\
+  \mu^b_m
 \end{pmatrix},
 \begin{pmatrix}
-  \sigma_{a[m]}^2 & \rho \sigma_{a} \sigma_{b} \\
-  \rho \sigma_a \sigma_b & \sigma_b^2
+  \sigma^{2,a} & \rho \sigma^{a} \sigma^{b} \\
+  \rho \sigma^a \sigma^b & \sigma^{2,b}
 \end{pmatrix} \Bigg) \\
-\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
-\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:unrest}
+\mu^a_m &\sim N(\theta^a, \sigma^{2, \mu^a_m}) \\
+\mu^b_m & \sim N(\theta^b, \sigma^{2, \mu^b_m}) \label{eqn:unrest}
 \end{aligned}
@@
 The lines represent approximate 95\% confidence intervals around the point estimates.
-One anomaly we can notice is that a handful of the systems have slopes that are estimated to be positive.
-This is clearly problematic if we wish to interpret the slope coefficient as degradation over time.
+One anomaly is that a handful of the systems have slopes that are estimated to be positive.
+This is problematic if we wish to interpret the slope coefficient as degradation over time.
 I address this in the full Bayesian model without resorting to excluding data.
@@
   \caption{Approximate 95\% confidence intervals around maximum likelihood point estimates of system-level slope terms, $b_i$, ordered from lowest to highest.
-Visually, we can see significant differences in the average degradation over time across solar power systems.}
+Visually, we can see significant differences in the average degradation over time across solar power systems.
+\label{sfig:sys_slope_fig}}
   \includegraphics[width=1\linewidth]{figures/sys_slope_fig.png}
-  \label{sfig:sys_slope_fig}
+
 \end{figure}
@@
 Figure \ref{sfig:nested_manuf_fig} gives a summary of the estimated manufacturer-level mean slopes.
-Again the lines represent 95\% confidence intervals around the point estimates.
+Again, the lines represent 95\% confidence intervals around the point estimates.
 Visually, we can see substantial variation in quality between the manufacturer groups.
@@
   \caption{Approximate 95\% confidence intervals around maximum likelihood estimators of manufacturer-level mean slope terms $\mu_b$, ordered from lowest to highest.
-These are the means of the higher level normal distributions that vary by manufacturer that the system-level parameters are assumed to be drawn from.
-The visual impression is of substantial variation between manufacturer groupings.}
+These are the means of the higher-level normal distributions that vary by manufacturer that the system-level parameters are assumed to be drawn from.
+The visual impression is of substantial variation between manufacturer groupings.
+\label{sfig:nested_manuf_fig}}
   \includegraphics[width=1\linewidth]{figures/nested_manuf_fig_final.png}
-  \label{sfig:nested_manuf_fig}
+
 \end{figure}
@@
   \end{tabular}
-  \label{tbl:lm_commodity}
+
   \caption{A likelihood ratio test indicates that the addition of panel manufacturer as a hierarchy significantly improves the predictive power of the model.
-Akaike Information Criterion (AIC) and Bayesian Information Criterion are lower in the model with manufacturer groupings, indicating better out-of-sample predictive fit.}
+Akaike Information Criterion (AIC) and Bayesian Information Criterion are lower in the model with manufacturer groupings, indicating better out-of-sample predictive fit.
+\label{tbl:lm_commodity}}
 \end{table}
 
-The first result, of which the evidence appears to be quite strong, is that solar panel systems do appear to vary substantially in quality between module manufacturers.
+The first result, of which the evidence is strong, is that solar panel systems vary substantially in quality between module manufacturers.
 Given that module quality is a salient differentiating characteristic between panels, it is then difficult to view solar panels as commodities.
@@
 
-From our informal model of asymmetric information where we see host owners of solar systems as being low-information types, while owners of leased panels as being high-information types, we would predict that on average leased panels will display higher quality over time.
+From our informal model of asymmetric information where we see host-owners of solar systems as being low-information types, while owners of leased panels as being high-information types, we would predict that on average leased panels will display higher quality over time.
 To test this directly with the data, I only need to do a slight modification of the model in the previous subsection.
@@
 \begin{aligned}
-log_prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation\_scaled_{i}, \sigma^2)\\
+prod_{i} &\sim N(\mathbf{month} + a_s + b_s months\_operation_{i}, \sigma^2)\\
 \begin{pmatrix}
@@
 \begin{pmatrix}
-  \mu_{a[l]}\\
-  \mu_{b[l]}
+  \mu^a_l\\
+  \mu^b_l
 \end{pmatrix},
 \begin{pmatrix}
-  \sigma_{a}^2 & \rho \sigma_{a} \sigma_{b} \\
-  \rho \sigma_a \sigma_b & \sigma_b^2
+  \sigma^{2,a} & \rho \sigma^{a} \sigma^{b} \\
+  \rho \sigma^a \sigma^b & \sigma^{2,b}
 \end{pmatrix} \Bigg) \\
-\mu_{a} &\sim N(\theta_a, \sigma_{\mu_a}^2) \\
-\mu_{b} & \sim N(\theta_b, \sigma_{\mu_b}^2) \label{eqn:fullmodel}
+\mu^{a} &\sim N(\theta^a, \sigma^{2,\mu_a}) \\
+\mu^{b} & \sim N(\theta^b, \sigma^{2,\mu_b}) \label{eqn:fullmodel}
 \end{aligned}
@@
 
-A visual summary of the results are shown in figure \ref{lease_sys_fig}.
+A visual summary of the results is shown in figure \ref{lease_sys_fig}.
 Again, the vertical lines represent approximate 95\% confidence intervals over the point estimates for the system-level slope terms, $b_s$.
-But here the the slope estimates are grouped into host-owned (left) and leased (right).
+But here the slope estimates are grouped into host-owned (left) and leased (right).
 The estimated group means are shown as horizontal lines.
@@
 The red line represents the zero threshold.
-The horizontal black line on the left represents the mean of the higher level distribution on the slopes for host-owned systems, $\mu_b[l=0]$, while the dotted black line on the right represents the mean of the higher level distribution on the slopes for third-party owned systems, $\mu_b[l=1]$ }
-	\label{lease_sys_fig}
+The horizontal black line on the left represents the mean of the higher level distribution on the slopes for host-owned systems, $\mu_b[l=0]$, while the dotted black line on the right represents the mean of the higher level distribution on the slopes for third-party owned systems, $\mu_b[l=1]$ \label{lease_sys_fig}}
+
 \end{figure}
@@
 \section{Full Bayesian hierarchical model}
-Up until now, I have used relatively parsimonious models with few fixed-effects and only one or two hierarchies ("random effects").
+Up until now, I have used relatively parsimonious models with few fixed effects and only one or two hierarchies (``random effects'').
 Ideally, I would like to estimate a full probability model with all relevant and available covariates and hierarchies.
-First and foremost so that the parameters of interests to the testable implications can be estimated simultaneously from one model.
+First and foremost, so that the parameters of interests to the testable implications can be estimated simultaneously from one model.
 But also, as \citet{barr_random_2013} shows, mixed-effects models that have the maximum random effects structure justified by the model, provide the best out-of-sample prediction.
@@
 I also wish to have a model where both system, module manufacturer, and leasing hierarchies are estimated simultaneously.
-Geographic information on the location of solar panels could also conceivably be correlated with degradation of solar panels, and should be accounted for in the model.
+Geographic information on the location of solar panels could also conceivably be correlated with degradation of solar panels and should be accounted for in the model.
 I am not able to obtain convergence of the maximum likelihood routines under this rich model.
@@
 
-Multilevel models estimated using Bayesian techniques have some nice theoretical properties, as I discussed earlier.
+To estimate the parameters of the model, I use the Stan Bayesian programming language and simulator \citep{stan_development_team_stan_2014}, which uses Hamiltonian MCMC to estimate a joint posterior distribution of the parameters.
+\citet{kruschke_doing_2014} provides an accessible explanation of the Hamiltonian MCMC algorithm, while more detailed technical descriptions are provided by \citet{gelman_bayesian_2013} and \citet{stan_development_team_stan_2014}.
+
+Bayesian methods making use of MCMC simulation are still emerging and this is an active field of research in both the statistical sciences and operations research fields.
+It is worth spending a moment to motivate their use in this case over asymptotic methods that are more common.
+The main reason for using the Bayesian model is the flexible treatment it allows for modeling the inherent hierarchy of the data and allowing for partial pooling in estimation.
+
+In a hierarchical model, lower-level parameters within a certain group are themselves modelled as coming from a distribution characterised by ``meta-parameters''.
+Thus, the estimated group-level parameters are a weighted function of both the observations within the group as well as the full set of observations in the data-set.
+For example, the estimated mean parameter on a group with a few, outlying observations would be pulled towards the full sample mean.
+This partial pooling also serves as a natural form of parameter shrinkage, which mostly eliminates the need for using corrections for multiple comparisons in inference \citep{gelman_data_2006}.
+
+With a Bayesian multilevel model, I also have greater flexibility in specifying the model without having to rely on assumptions like constant group-level variances and a Gaussian noise distribution.
+Because Bayesian simulation techniques result in an estimate of the full joint probability distribution of the model, the inference has the potential to be more informative than the typical point estimates and p-values of the standard hypothesis testing frameworks \citep{kruschke_doing_2014}.
+
+Another advantage is that the posterior distributions of the parameters have a natural and direct interpretation as probabilities, as opposed to standard hypothesis testing where the concepts of p-values and significance are often ill-defined.
+See \citet{kruschke_doing_2014} or \citet{gelman_bayesian_2013} for further discussion.
+
+Hierarchical models estimated using Bayesian techniques have some nice theoretical properties.
 However, the main advantages to using the Bayesian techniques here are practical.
@@
 
-A Bayesian MCMC approach has some additional attractive properties.
-Results are presented as estimates of marginal posterior distributions over coefficients of interest, which can be interpreted directly and intuitively as probabilities.
-Group-level variances are not constrained to be equal and assumptions of normality of the likelihood and posterior distributions can be loosened.
-For further discussion of the Bayesian frameworks and techniques, I refer to discussions in \citet{gelman_bayesian_2013}, \citet{kruschke_doing_2014}, or \citet{mcelreath_statistical_2015}.
-
 The model is described by the equations in \ref{eqn:bayesMod}.
@@
 This response variable is given a normal likelihood with mean, $\hat{y}$ and standard deviation, $\sigma$.
-At the observation level, the fitted values, $\hat{y}$, are modeled as a simple linear regression, with an intercept $a_s$, and a slope $b_s$ which are both allowed to vary by system, $s$.
-In addition a vector of month dummies, $\mathbf{\mu^{month}}$ is included in order to control for seasonality.
+At the observation level, the fitted values, $\hat{y}$, are modelled as a simple linear regression, with an intercept $a_s$, and a slope $b_s$, which are both allowed to vary by system, $s$.
+In addition a vector of month dummies, $\mathbf{month_c}$, is included in order to control for seasonality.
+In contrast to the earlier estimation, I allow the vector of coefficients on the month dummies to vary by county, indexed by $c$.
+California is a large, geographically diverse state, and the seasonal profile of solar generation is bound to vary across counties.
 
-At the system level, the intercept terms, $a_s$ are given a common higher level Cauchy distribution with location parameter $\mu^a$ and scale parameter $\sigma^a$.
-But because the intercept term is not of direct economic importance, it is not modeled further.
+At the system level, the intercept terms, $a_s$ are given a common higher level normal distribution with mean $\mu^a$ and standard deviation $\sigma^a$.
 
-The slope terms, $b_s$ are also modeled as Normal distributions with mean $\mu^{b}_s$, that is allowed to vary by system, and standard deviation $\sigma^{b}$.
-I decompose the variation of $\mu^{b}_s$ into three terms that vary by lease, $\mu^{lease}_l$, host sector $\mu^{sect}_c$, and manufacturer $\mu^m_m$.
-In addition three continuous variables are included in the system-level equation: the year of first production, $first\_prod\_year_s$, capacity, $capacity_s$ and the system cost per kilowatt, $cost_s$ with corresponding coefficients $\beta^{fy}$, $\beta^{size}$, and $\beta^{cost}$.
+The slope terms, $b_s$ are also modelled as normal distributions with mean terms $\mu^{b}_s$ and standard deviations $\sigma^{b}_s$, which are both allowed to vary by system.
+I decompose the variation of $\mu^{b}_s$ by groupings and covariates of interest.
+I account for variation by lease, $\mu^{lease}_l$, host sector $\mu^{sect}_c$, and manufacturer $\mu^{manuf}_m$.
+In addition, three continuous variables are included in the system-level equation: the year of first production, $first\_prod\_year_s$, capacity, $capacity_s$ and the system cost per kilowatt, $cost_s$ with corresponding coefficients $\beta^{fy}$, $\beta^{size}$, and $\beta^{cost}$.
+The coefficients in this equation can be interpreted as interaction effects with the slope terms as in a traditional regression.
+It can also be helpful to think of it as a two-stage regression, where the $b_s$ terms are first estimated and interpreted as a proxy for the quality of a solar power system.
+The estimated $\hat{b_s}$ terms are in turn regressed on the explanatory variables.
+In reality, estimation is done simultaneously and the $b_s$ parameters are appropriately regularized by the higher-level distributions.
+
+\begin{table}
+  \begin{tabular}{rl}
+  \toprule
+    Symbol &  Definition \\
+    \midrule
+    $prod_{i}$ & Production, descaled, from a solar power system at a given \\
+    &month (observation i).\\
+    $months\_operation_{st}$ & Number of months of operation of solar system s at month t \\
+    $\mathbf{month_c}$ & Vector of month dummy variables that can vary by county, c\\
+    $first\_prod\_year_s$ & Year of first production from solar system, s\\
+    $log\_csi\_rating_s$ & Size of solar power system, as measured by log production capacity, \\
+    &descaled, of solar power system, s\\
+    $log\_cost_s$ & Log cost, descaled, of solar power system, s \\
+    $\hat{y_{st}}$ & Mean of the response variable, production for system s at time t\\
+    $\sigma$ & Standard deviation of the response variable, production \\
+    $a_s$& Intercept parameter for a system, s \\
+    $b_s$ & Slope parameter for a system, s \\
+    $\mu^a$ & Mean of the distribution of system intercept terms, $a_s$ \\
+    $\mu^b_s$ & Mean of the distribution of system slope terms, $b_s$.  \\
+    $\sigma^{a}$ & Variance of the distribution of system intercept terms, $a_s$\\
+    $\sigma^{b}_s$ & Variance of the distribution of the system slope terms, $b_s$. \\
+    & Can vary by system, s \\
+    $\mu^{month}$ & Mean of the distribution of the county-varying month indicators\\
+    $\sigma^{month}$ & Variance of the distribution of the county-varying month indicators\\
+    $x_0$ & Mean of the higher level (meta) distribution of the slope \\
+    & standard deviations, $\sigma^b_s$\\
+    $\gamma$ &  Variance of the higher level (meta) distribution of the slope \\
+    & standard deviations, $\sigma^b_s$\\
+    $\mu_l^{lease}$ & Mean parameters for distribution of slopes of \\
+    & leased and host-owned groups, l \\
+    $\mu_e^{sector}$ & Mean parameters for distribution of slopes of host sector groups, e.\\
+    $\mu_m^{manuf}$ & Mean parameters for distribution of slopes of panel manufacturer groups, m. \\
+    $\beta^{fy} $ & Parameter on variable for first production year, $first\_prod\_year$\\
+    $\beta^{size}$ & Parameter on variable for size of solar power system, $log\_csi\_rating$ \\
+    $\beta^{cost}$ & Parameter on variable for size of solar power system, $log\_csi\_rating$ \\
+
+  \bottomrule
+  \end{tabular}
+
+  \caption{Definitions from the full Bayesian model presented in equation \ref{eqn:bayesMod}. \label{tbl:definitions4p1}}
+\end{table}
 
@@
 \begin{aligned}
-prod_{st} &\sim N(\hat{y}_{st}, \sigma)\\ %\label{eqn:likelihood}
-\hat{y}_{st} &= a_s + b_s months\_operation_{st} + \mathbf{\mu^{month}}\\
+prod_{i} &\sim N(\hat{y}_{st}, \sigma)\\ %\label{eqn:likelihood}
+\hat{y}_{st} &= a_s + b_s months\_operation_{st} + \mathbf{month_c}\\
 a_s &\sim N(\mu^a, \sigma^a)  \label{eqn:bayesMod} \\
-b_s &\sim N(\mu_s^b, \sigma^b) \\
-\mu_{s}^{b} & = \mu^{lease}_l + \mu^{sect}_s + \mu^{manuf}_m + \mu^{county}_c  \\
+b_s &\sim N(\mu_s^b, \sigma_s^b) \\
+\mathbf{month_c} &\sim N(\mu^{month}, \sigma^{month})\\
+\sigma_s^{b} &\sim \text{half-Cauchy}(x_0,\gamma) \\
+\mu_{s}^{b} & = \mu^{lease}_l + \mu^{sect}_e + \mu^{manuf}_m  \\
 & +\beta^{fy} first\_prod\_year_s + \beta^{size} log\_csi\_rating_s + \beta^{cost} log\_cost_s
@@
 
-Higher level terms are all given Cauchy prior distributions as summarized in the equations in \ref{eqn:meta}.
+The standard deviation parameters and higher level terms are all given Cauchy prior distributions as summarized in the equations in \ref{eqn:meta}.
 The Cauchy distribution, which is a T-distribution with 1 degree of freedom, has wide tails and thus allows for the occasional outlying coefficient in a lower level regression, while still provides some regularization towards zero \citep{gelman_weakly_2008}.
-
-The $\mu^{lease}_l$, $\mu^{sect}_s$, and $\mu^{manuf}$ terms are all
-modeled as coming from a Cauchy distribution with a location meta-parameter $\mu^{meta}$ and scale parameter $\sigma^{mu}$.
-Similarly, $\beta^{fy}$, $\beta^{size}$ and $\beta^{cost}_l$ are modeled as coming from a $Cauchy(\mu^{\beta}, \sigma^{\beta})$ distribution.
-The location parameters $\mu^{meta}$ and $\mu^{\beta}$ are in turn given a Cauchy(0,5) prior, while the corresponding $\sigma^{meta}$ and $\sigma^{beta}$ terms are given half-Cauchy(0,5) priors--essentially a Cauchy distribution constrained to the positive values.
-
-The $Cauchy(0,5)$ prior is what is referred to as a weakly-informative prior.
-It imposes some weak regularization on the parameters, while also allowing for some outlying parameter estimates.
-At the same time, the prior does not impose any strong prior knowledge that meaningfully effects the posterior distribution, especially in the presence of substantial data.
-The parameters $\beta^{fy}$, $\beta^{size}$, $\beta^{cost}$ and $\mathbf{\mu^{month}}$ are all also given $Cauchy(0,5)$ prior distributions.
+The $\text{Cauchy}(0,1)$ prior for location parameters of de-scaled variables is what is referred to as a weakly-informative prior.
+As opposed to a completely flat prior, it imposes some weak regularization on the parameters in order to avoid over-fitting the data and improve predictive performance.
+At the same time, the prior does not impose any strong prior knowledge.
+The $\text{half-Cauchy}(0,5)$ prior for scale parameters is in practice coded as a $\text{Cauchy}(0,5)$ distribution restricted to a positive support
 
@@
 \begin{aligned}
-\mu^{lease}_l &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\
-\mu^{sect}_s &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \\
-\mu^{manuf}_m &\sim Cauchy(\mu^{meta}, \sigma^{\mu}) \label{eqn:meta} \\
-\mu^{meta} &\sim Cauchy(0, 5) \\
-\mu^{a} &\sim Cauchy(0,5) \\
-\mathbf{\mu^{month}} &\sim Cauchy(0,5) \\
-\beta^{fy} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta}) \\
-\beta^{size} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta})\\
-\beta_l^{cost} &\sim Cauchy(\mu^{\beta}, \sigma^{\beta})\\
-\mu^{\beta} &\sim Cauchy(0, 5) \\
-\sigma &\sim half-Cauchy(0,5) \\
-\sigma^{a} &\sim half-Cauchy(0,5) \\
-\sigma^{b} &\sim half-Cauchy(0,5) \\
-\sigma^{\mu} &\sim half-Cauchy(0,5)\\
-\sigma^{\beta} &\sim half-Cauchy(0,5)\\
+\mu^{lease}_l &\sim \text{Cauchy}(0,1) \\
+\mu^{sect}_s &\sim \text{Cauchy}(0,1) \\
+\mu^{manuf}_m &\sim \text{Cauchy}(0,1) \label{eqn:meta} \\
+\mu^{a} &\sim \text{Cauchy}(0,1) \\
+mon_c &\sim \text{Cauchy}(0,1) \\
+\beta^{fy} &\sim \text{Cauchy}(0,1) \\
+\beta^{size} &\sim \text{Cauchy}(0,1)\\
+\beta_l^{cost} &\sim \text{Cauchy}(0,1)\\
+x_0 &\sim \text{Cauchy}(0,1)\\
+\sigma &\sim \text{half-Cauchy}(0,5) \\
+\sigma^{a} &\sim \text{half-Cauchy}(0,5) \\
+\sigma^{mon} &\sim \text{half-Cauchy}(0,5) \\
+\gamma &\sim \text{half-Cauchy}(0,5)\\
 \end{aligned}
@@
 	\caption{An overview of the structure of the model.
-Starting from the top in panel I, the monthly production data is modeled with a normally distributed likelihood.
-Moving down to II, the fitted values, $\hat{y}_{st}$, are in turn decomposed into system-varying slope and intercept terms, as well as month-dummies that control for seasonality.
-Further down to III, the slope terms, $b_s$ are modeled as normal distributions with varying mean parameters, $\mu_b^s$ that decompose further into terms representing variables of interest and control variables.
-Finally, in IV, all parameters are given higher-level meta-parameters that serve to regularize (or "shrink") the lower-level parameter estimates.}
-	\label{solar_prod_hierarchy}
+Starting from the top in panel I, the monthly production data is modelled with a Normally distributed likelihood.
+Moving down to II, the fitted values, $\hat{y}_{i}$, are in turn decomposed into system-varying slope and intercept terms, as well as month-dummies that control for seasonality.
+Further down to III, the slope terms, $b_s$ are modelled as Normal distributions with varying mean parameters, $\mu_s^b$ that decompose further into terms representing variables of interest and control variables.
+Finally, in IV, all parameters are given higher-level meta-parameters that serve to regularize (or "shrink") the lower-level parameter estimates.
+\label{solar_prod_hierarchy}}
+
 \end{figure}
@@
 In the earlier section, we noticed that some of the system slope parameters were estimated to be positive.
-There are intuitive and plausible reasons for observing higher average production some time after the initial installation.
+There are intuitive and plausible reasons for observing higher average production sometime after the initial installation.
 The placement of solar panels could have been optimized after the fact.
@@
 Several possibilities present themselves.
-I could discard the systems with positive slope parameters from the data used.
-But discarding data and the associated information is generally not a desirable.
+I could discard the systems with positive slope parameters from the data.
+But discarding data is generally not desirable.
 Alternatively, I could keep in all data, allowing for the slope terms to be positive.
-After all, one of the advantages of the hierarchical model with weakly regularizing priors is that it allows for outlying parameter estimates, without unduly effecting the overall results.
-A third solution is to include all the data, but constrain the model by having priors that place zero probability on positive slope parameter values.
-$b_s$ parameters would be given half-Cauchy distributions with support between 0 and negative infinity.
-This has the desired effect of constraining slope parameters to be negative by imposing prior knowledge.
-The downside is that such strong priors can sometimes lead to unintended effects on the posterior distribution and can lead to faulty inference.
+One of the advantages of the hierarchical model with weakly regularizing priors is that it allows for outlying parameter estimates, without unduly affecting the overall results.
+A third solution is to include all the data but constrain the model by specifying priors that place zero probability on positive slope parameter values.
+In particular the $b_s$ parameters would be given Cauchy distributions with support between 0 and negative infinity.
+This has the desired effect of constraining slope parameters to be negative by imposing strong prior knowledge.
+The downside is that such strong priors can sometimes lead to unintended effects on the posterior distribution and faulty inference.
+In the Bayesian literature, such hard constraints are generally not recommended.
 In the proceeding analysis I therefor present results from both constrained and unconstrained models.
@@
 The sampler was run with four chains and 1000 iterations.
-A Gelman-Rubin convergence statistic ($\hat{R}$ of 1 indicated convergence in the model.
+Gelman-Rubin convergence statistics ($\hat{R})$ were used to check convergence in the model.
 2000 samples are taken from the simulated posterior distribution, and model results for the marginal distributions of the parameters of interest are presented in the form of kernel densities of the samples.
 
-Recall that the testable implications from the theory of asymmetric information and quality were that 1.) average quality should be higher in the panels owned by the high-information type (leased panels).
-2.) Price should be more highly correlated with quality in the panels owned by the high-information type.
+Recall that the testable implications from the theory of asymmetric information and quality were that 1.) average quality should be higher in the panels owned by the high-information types (leased panels).
+2.) Price should be more highly correlated with quality in the panels owned by high-information types.
 
-The parameters of interest in our model are then the higher level parameters $\mu_l^{lease}$ and $\beta_l$.
-The first represents the mean value of the system level slope terms, $b_s$ grouped by whether the systems were leased or host-owned.
+The parameters of interest in our model are then the higher-level parameters $\mu_l^{lease}$ and $\beta_l^{cost}$.
+The first represents the mean value of the system-level slope terms, $b_s$, grouped by whether the systems were leased or host-owned.
 The latter is a measure of the correlation between the cost of the system and the system slope terms.
-The $\beta_l$ parameter is also allowed to vary by leased or host-owned.
-It is important to note, that the estimated $\beta_l$ parameters are estimated conditional on the installation date of the solar panels, thus they do not reflect changes in price over time.
+The $\beta_l^{cost}$ parameter is also allowed to vary by leased or host-owned.
+Importantly, the $\beta_l$ parameters are estimated conditional on the installation date of the solar panels; thus they do not reflect changes in price over time.
 This could lead to a heavy bias in the results if, as seems plausible, both prices and quality changed over time.
 
-For the testable implications, the inference we are interested is in the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta_{owned}$.
+For the testable implications, the inference we are interested in is the contrasts, $\mu_{lease, yes} - \mu_{lease, no}$ and $\beta_{leased}- \beta_{owned}$.
 Kernel densities representing the marginal distributions of these contrasts are presented in figure \ref{fig:contrasts}.
 The top panel shows the distribution of the contrast $\mu_{lease, yes} - \mu_{lease, no}$.
-Essentially all of the probability mass is located in the positive range.
+Approximately 99\% of the probability mass is located in the positive range.
 This reinforces the finding from the mixed effects model that leased panels do appear to have substantially higher quality, as the theory of asymmetric information of quality would suggest.
@@
 The evidence for the second testable implication is less conclusive.
-The model estimates that there is approximately a 89\% probability that there is a higher correlation between quality and cost for the high information type owner of leased panels compared to the low information type owner of host-owned panels.
+The model estimates that there is approximately a 85\% probability that there is a higher correlation between quality and cost for the high information type owner of leased panels compared to the low information type owner of host-owned panels.
 
 Figure \ref{fig:contrasts_const} show the corresponding distributions under the constrained model.
-$\mu_{lease, yes} - \mu_{lease, no}$ is pulled somewhat towards zero in magnitude, but nearly all of the probability mass is still in the positive range.
-The distribution of  $\beta_{leased}- \beta_{owned}$ is also pulled towards zero, and now 72\% of the probability mass is in the positive range.
+$\mu_{lease, yes} - \mu_{lease, no}$ is pulled somewhat towards zero in magnitude, but about 95\% of the probability mass is still in the positive range.
+The distribution of  $\beta_{leased}- \beta_{owned}$ is also pulled towards zero, and now is centered close to zero.
 
@@
   \includegraphics[width=1\linewidth]{figures/bayes_hypos.png}
-  \label{fig:contrasts}
  \end{minipage}\qquad
@@
   \includegraphics[width=1\linewidth]{figures/bayes_hypos_const.png}
-  \label{fig:contrasts_const}
+
  \end{minipage}
@@
 \caption{Marginal posterior distributions for the contrasts of interest.
-The top panel shows that the distributions of the slope parameters, $b_s$, are estimated to have means that are significantly different from each other between leased and host-owned systems as defined by the contrast, $\mu_{lease, yes} - \mu_{lease, no}$.
-The lower panel shows the marginal posterior distribution of the contrast between the $\beta$ parameter for leased and host-owned solar systems.
-The results indicate that there is an approximately 89\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned.}
+The top panel shows that the distributions of the slope parameters, $b_s$ between leased and host-owned systems as defined by the contrast, $\mu_{lease, yes} - \mu_{lease, no}$ have approximately 99\% probability of being positively different from each other.
+The lower panel shows the marginal posterior distribution of the contrast between the $\beta_l^{cost}$ parameter for leased and host-owned solar systems.
+The results indicate that there is an approximately 85\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned.}
+ \label{fig:contrasts}
 \end{minipage}\qquad
@@
 \centering
- \caption{Marginal posterior distributions for the contrasts when the $b_s$ paramaters and corresponding higher-level mean parameters are constrained to have a maximum of zero.
-The contrast, $\mu_{lease, yes} - \mu_{lease, no}$ is materially unchanged and continues to have nearly its entire distribution in a positive range, though it is pulled somewhat towards zero.
-The contrast $\beta_{lease, yes} - \beta_{lease, no}$ is also pulled towards zero, and now only approximately 72 \% of the probability is in a positive range.
-The results indicate that there is an approximately 72\% probability that costs are more highly correlated with quality in leased systems compared to those that are host-owned when all higher level mean parameters are constrained to have a maximum of zero.}
+ \caption{Marginal posterior distributions for the contrasts when the $b_s$ parameters and corresponding higher-level mean parameters are constrained to have a maximum of zero.
+The contrast, $\mu_{lease, yes} - \mu_{lease, no}$ is pulled somewhat towards zero, but continues to have approximately 95\% percent of its probability mass in the positive range.
+The contrast $\beta_{lease, yes} - \beta_{lease, no}$ is also pulled towards zero, and the distribution of the contrast is closely centered around zero.}
+ \label{fig:contrasts_const}
 \end{minipage}
@@
 
+Interpretation of the $\beta_l^{cost}$ parameters must be done with some care.
 A direct comparison of costs between leased and host-owned solar power systems is somewhat complicated.
-The prices of host-owned panels are market prices: simply the purchase price that the owner of the solar panel system paid to the contractor.
+The prices of host-owned panels are market prices: Simply the purchase price that the owner of the solar panel system paid to the contractor.
 On the other hand, the price of leased systems is a reported price from the contractors consisting of the component and installation costs plus an estimated mark-up.
@@
 The reported cost should reflect the ability of the high-information owners to judge quality of the panels that they purchase.
-In practice, the reliability of reported cost is a source of added uncertainty, and should be taken into consideration when interpreting the results.
+In practice, the reliability of reported cost is a source of added uncertainty and should be taken into consideration when interpreting the results.
 
@@
 The top panel represents the contrast with commercial sector hosts, the middle with government sector host and the bottom panel with non-profit sector host.
-All three contrasts indicate relatively lower quality for installations for these sectors compared to the residential sector.
+The contrasts for commercial and non-profit hosts relative to residential have distributions centered close to zero.
+There is little evidence that quality varies significantly.
+The distribution for the contrast between government and residential solar panels has about 90 percent of its probability mass in the positive range.
+We can interpret this to say that there is a 90\% probability that solar panels systems with governmental hosts display on average higher quality than similar solar panels systems with residential hosts.
 Results of the constrained model are shown in figure \ref{fig:mu_sectors_const}.
-The results are similar, though the difference in average quality between the government and residential sector is now centered closely around zero.
+The results are similar.
 
-While it is not clear why this should be, an explanation consistent with the theory of asymmetric information of quality is that there exists an agency problem.
-A homeowner will, in some form, have a direct financial incentive to acquire solar panels of good quality.
-On the other hand, A manager or employee of a firm, government agency or non-profit may not have any direct financial incentive in ensuring that a solar panel system is of high quality.
+While it is not clear why solar panel systems with governmental hosts should display higher quality, an explanation consistent with the theory of asymmetric information of quality is that governmental host may be better informed.
+Local or state governments are likely repeat buyers of solar panel systems and have acquired knowledge about quality.
 
@@
   \midrule
-  $\mu^{meta}$     &  0.004' & 0.022' & -0.018' & -0.010' & -0.007' & -0.005' & -0.002' &  0.069' &  0.084' \\
-  $\beta^{meta}$   &  0.003' & 0.838' & -8.120' & -1.707' & -0.074' & -0.007' &  0.075' &  1.738' &  6.787' \\
-  $\mu^{meta,mon}$ &  0.005' & 0.036' & -0.138' & -0.065' & -0.019' &  0.006' &  0.030' &  0.075' &  0.146' \\
-  $\mu^{b_s}$      & -0.000' & 0.015' & -0.043' & -0.028' & -0.012' & -0.002' &  0.011' &  0.031' &  0.044' \\
-  $\mu^{jan}$      & -0.164' & 0.013' & -0.188' & -0.186' & -0.173' & -0.163' & -0.154' & -0.144' & -0.142' \\
-  $\mu^{feb}$      & -0.121' & 0.013' & -0.146' & -0.143' & -0.131' & -0.120' & -0.111' & -0.102' & -0.100' \\
-  $\mu^{mar}$      & -0.054' & 0.013' & -0.078' & -0.076' & -0.064' & -0.053' & -0.044' & -0.035' & -0.033' \\
-  $\mu^{apr}$     &  0.040' & 0.013' &  0.017' &  0.018' &  0.031' &  0.042' &  0.051' &  0.060' &  0.062' \\
-  $\mu^{may}$      &  0.087' & 0.013' &  0.063' &  0.064' &  0.077' &  0.088' &  0.097' &  0.106' &  0.108' \\
-  $\mu^{jun}$      &  0.111' & 0.013' &  0.087' &  0.089' &  0.102' &  0.113' &  0.121' &  0.131' &  0.133' \\
-  $\mu^{jul}$      &  0.108' & 0.013' &  0.084' &  0.086' &  0.099' &  0.110' &  0.118' &  0.128' &  0.130' \\
-  $\mu^{aug}$      &  0.101' & 0.013' &  0.077' &  0.079' &  0.092' &  0.103' &  0.112' &  0.121' &  0.123' \\
-  $\mu^{sep}$     &  0.082' & 0.013' &  0.058' &  0.060' &  0.072' &  0.083' &  0.092' &  0.102' &  0.103' \\
-  $\mu^{oct}$      &  0.026' & 0.013' &  0.002' &  0.004' &  0.016' &  0.027' &  0.036' &  0.045' &  0.047' \\
-  $\mu^{nov}$     & -0.031' & 0.013' & -0.056' & -0.053' & -0.041' & -0.030' & -0.021' & -0.011' & -0.010' \\
-  $\mu^{dec}$     & -0.123' & 0.013' & -0.147' & -0.146' & -0.133' & -0.122' & -0.113' & -0.104' & -0.101' \\
-  $\mu^{lease, no}$   & -0.030' & 0.045' & -0.193' & -0.159' & -0.015' & -0.011' & -0.008' & -0.002' &  0.015' \\
-  $\mu^{lease, yes}$   & -0.022' & 0.045' & -0.186' & -0.151' & -0.007' & -0.003' &  0.000' &  0.006' &  0.023' \\
-  $\mu^{sect, com}$      &  0.002' & 0.022' & -0.024' & -0.015' & -0.010' & -0.007' & -0.004' &  0.066' &  0.087' \\
-  $\mu^{sect, res}$       &  0.012' & 0.022' & -0.014' & -0.005' & -0.000' &  0.003' &  0.006' &  0.075' &  0.096' \\
-  $\mu^{sect, gov}$       &  0.006' & 0.022' & -0.017' & -0.010' & -0.005' & -0.003' &  0.001' &  0.071' &  0.092' \\
-  $\mu^{sect, npr}$      & -0.001' & 0.022' & -0.026' & -0.018' & -0.012' & -0.010' & -0.006' &  0.063' &  0.084' \\
-  $\beta^{cost, l=no}$  & -0.002' & 0.002' & -0.007' & -0.005' & -0.003' & -0.002' & -0.001' &  0.001' &  0.004' \\
-  $\beta^{cost, l=yes}$  &  0.001' & 0.001' & -0.002' & -0.001' &  0.000' &  0.001' &  0.001' &  0.002' &  0.003' \\
-  $\beta^{size}$  &  0.002' & 0.001' & -0.001' &  0.000' &  0.002' &  0.002' &  0.003' &  0.004' &  0.006' \\
-  $\beta^{fy}$    & -0.018' & 0.001' & -0.021' & -0.020' & -0.018' & -0.018' & -0.017' & -0.016' & -0.014' \\
-  $\sigma$      &  0.065' & 0.000' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  0.066' &  0.066' \\
-  $\sigma^{b_s}$    &  0.476' & 0.006' &  0.454' &  0.465' &  0.472' &  0.476' &  0.480' &  0.488' &  0.495' \\
-  $\sigma^{mon}$   &  0.114' & 0.030' &  0.059' &  0.074' &  0.094' &  0.108' &  0.128' &  0.185' &  0.341' \\
-  $\sigma^{\beta}$  &  0.833' & 1.773' &  0.001' &  0.006' &  0.047' &  0.195' &  0.836' &  5.675' & 19.879' \\
-  $\sigma^{mu}$   &  0.006' & 0.001' &  0.003' &  0.004' &  0.005' &  0.006' &  0.006' &  0.008' &  0.011' \\
+  $\mu^a$     &  0.007 & 0.052 & -0.116 & -0.094 & -0.026 &  0.012 &  0.040 &  0.095 &  0.124 \\
+$\sigma^a$   &  0.907 & 0.012 &  0.872 &  0.885 &  0.899 &  0.907 &  0.915 &  0.934 &  0.952 \\
+$\sigma^{cm} $  &  0.299 & 0.009 &  0.268 &  0.280 &  0.293 &  0.298 &  0.305 &  0.315 &  0.331 \\
+$\sigma$     &  0.321 & 0.001 &  0.319 &  0.320 &  0.320 &  0.321 &  0.321 &  0.322 &  0.323 \\
+$\mu^{lease, no} $ & -0.005 & 0.370 & -1.008 & -0.824 & -0.239 & -0.028 &  0.271 &  0.705 &  0.936 \\
+$\mu^{lease, yes}$  &  0.001 & 0.370 & -1.000 & -0.814 & -0.232 & -0.022 &  0.278 &  0.716 &  0.943 \\
+$\mu^{sect, com}$      & -0.048 & 0.348 & -0.995 & -0.752 & -0.300 & -0.026 &  0.163 &  0.707 &  0.887 \\
+$\mu^{sect, res} $     & -0.049 & 0.348 & -0.998 & -0.753 & -0.302 & -0.027 &  0.162 &  0.708 &  0.885 \\
+$\mu^{sect, gov} $    & -0.045 & 0.348 & -0.990 & -0.748 & -0.298 & -0.024 &  0.168 &  0.716 &  0.889 \\
+$\mu^{sect, npr}$      & -0.049 & 0.348 & -0.999 & -0.756 & -0.300 & -0.025 &  0.163 &  0.706 &  0.893 \\
+$\beta^fy$   & -0.006 & 0.002 & -0.013 & -0.010 & -0.007 & -0.006 & -0.005 & -0.003 & -0.001 \\
+$\beta_{own}^{cost}$ & -0.001 & 0.001 & -0.005 & -0.003 & -0.002 & -0.001 &  0.000 &  0.002 &  0.004 \\
+$\beta_{lease}^{cost}$ &  0.001 & 0.002 & -0.005 & -0.002 &  0.000 &  0.001 &  0.002 &  0.004 &  0.006 \\
+$\beta^{size}$ & -0.021 & 0.001 & -0.025 & -0.024 & -0.022 & -0.021 & -0.020 & -0.018 & -0.016 \\
   \bottomrule
   \end{tabular}
-\label{table:higher_level_params}
-\caption{Summary statistics of the estimated posterior distributions of higher level parameters}
+\caption{\label{table:higher_level_params} Summary statistics of the estimated posterior distributions of higher level parameters}
 \end{table}
@@
 \midrule
-beta1\_fy0    & -0.017' & 0.001' & -0.021' & -0.019' & -0.018' & -0.017' & -0.017' & -0.016' & -1.454096e-02 \\
-meta\_mu0     & -0.006' & 0.001' & -0.009' & -0.008' & -0.007' & -0.006' & -0.006' & -0.005' & -3.700574e-03 \\
-meta\_beta0   &  0.012' & 0.798' & -4.910' & -1.877' & -0.083' & -0.009' &  0.080' &  1.989' &  6.342196e+00 \\
-meta\_mu\_mon0 & -0.114' & 0.034' & -0.280' & -0.180' & -0.137' & -0.115' & -0.093' & -0.044' &  3.362246e-02 \\
-sigma\_mu10   &  0.003' & 0.001' &  0.002' &  0.002' &  0.003' &  0.003' &  0.003' &  0.004' &  5.319401e-03 \\
-mu\_b00       &  0.121' & 0.011' &  0.088' &  0.101' &  0.113' &  0.121' &  0.129' &  0.142' &  1.531909e-01 \\
-sigma\_b00    &  0.476' & 0.006' &  0.459' &  0.464' &  0.472' &  0.476' &  0.480' &  0.487' &  4.995173e-01 \\
-sigma\_mon0   &  0.114' & 0.029' &  0.055' &  0.075' &  0.094' &  0.108' &  0.127' &  0.188' &  2.955012e-01 \\
-sigma\_beta0  &  0.800' & 1.495' &  0.006' &  0.010' &  0.051' &  0.201' &  0.814' &  5.056' &  1.976847e+01 \\
-sigma0       &  0.065' & 0.000' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  0.065' &  6.572528e-02 \\
-mu\_mon0      & -0.285' & 0.007' & -0.298' & -0.297' & -0.290' & -0.286' & -0.278' & -0.275' & -2.736025e-01 \\
-mu\_mon1      & -0.243' & 0.007' & -0.256' & -0.254' & -0.248' & -0.243' & -0.235' & -0.233' & -2.310124e-01 \\
-mu\_mon2      & -0.176' & 0.007' & -0.189' & -0.187' & -0.181' & -0.176' & -0.168' & -0.166' & -1.642462e-01 \\
-mu\_mon3      & -0.081' & 0.007' & -0.094' & -0.093' & -0.086' & -0.082' & -0.074' & -0.071' & -6.978557e-02 \\
-mu\_mon4      & -0.035' & 0.007' & -0.048' & -0.047' & -0.040' & -0.035' & -0.028' & -0.025' & -2.350794e-02 \\
-mu\_mon5      & -0.010' & 0.007' & -0.023' & -0.022' & -0.015' & -0.011' & -0.003' & -0.000' & -1.024029e-05 \\
-mu\_mon6      & -0.013' & 0.007' & -0.026' & -0.025' & -0.018' & -0.014' & -0.006' & -0.003' & -1.735978e-03 \\
-mu\_mon7      & -0.020' & 0.007' & -0.034' & -0.031' & -0.025' & -0.021' & -0.013' & -0.010' & -8.326959e-03 \\
-mu\_mon8      & -0.039' & 0.007' & -0.053' & -0.051' & -0.045' & -0.040' & -0.032' & -0.030' & -2.838155e-02 \\
-mu\_mon9      & -0.096' & 0.007' & -0.109' & -0.107' & -0.101' & -0.096' & -0.089' & -0.086' & -8.434625e-02 \\
-mu\_mon10     & -0.152' & 0.007' & -0.166' & -0.164' & -0.158' & -0.153' & -0.145' & -0.143' & -1.412385e-01 \\
-mu\_mon11     & -0.245' & 0.007' & -0.258' & -0.256' & -0.250' & -0.245' & -0.237' & -0.235' & -2.336429e-01 \\
-mu1\_lease0   & -0.005' & 0.001' & -0.008' & -0.007' & -0.006' & -0.005' & -0.005' & -0.004' & -2.602403e-03 \\
-mu1\_lease1   & -0.000' & 0.000' & -0.002' & -0.001' & -0.000' & -0.000' & -0.000' & -0.000' & -5.518242e-08 \\
-mu1\_s0       & -0.002' & 0.001' & -0.005' & -0.004' & -0.003' & -0.002' & -0.002' & -0.001' & -1.790787e-04 \\
-mu1\_s1       & -0.001' & 0.000' & -0.003' & -0.002' & -0.001' & -0.000' & -0.000' & -0.000' & -2.355766e-06 \\
-mu1\_s2       & -0.001' & 0.001' & -0.004' & -0.002' & -0.001' & -0.001' & -0.000' & -0.000' & -1.807113e-06 \\
-mu1\_s3       & -0.007' & 0.002' & -0.014' & -0.010' & -0.008' & -0.007' & -0.006' & -0.004' & -2.390480e-03 \\
-beta1\_cost0  & -0.000' & 0.002' & -0.005' & -0.003' & -0.001' & -0.000' &  0.001' &  0.003' &  4.742583e-03 \\
-beta1\_cost1  &  0.001' & 0.001' & -0.002' & -0.001' &  0.000' &  0.001' &  0.001' &  0.002' &  3.464386e-03 \\
-beta1\_size0  &  0.001' & 0.001' & -0.002' & -0.000' &  0.001' &  0.001' &  0.002' &  0.003' &  3.930082e-03 \\
+$\mu^a$      & -0.037 & 0.035 & -0.129 & -0.104 & -0.065 & -0.031 & -0.010 &  0.021 &  0.045 \\
+$\sigma^a$   &  0.907 & 0.012 &  0.868 &  0.884 &  0.899 &  0.906 &  0.915 &  0.930 &  0.946 \\
+$\sigma^{cm} $    &  0.298 & 0.010 &  0.269 &  0.280 &  0.291 &  0.297 &  0.304 &  0.318 &  0.333 \\
+$\sigma$      &  0.328 & 0.001 &  0.326 &  0.327 &  0.328 &  0.328 &  0.329 &  0.329 &  0.330 \\
+$\mu^{lease, no} $  &  0.163 & 0.221 & -0.589 & -0.404 &  0.060 &  0.181 &  0.291 &  0.578 &  0.695 \\
+$\mu^{lease, yes}$  &  0.165 & 0.221 & -0.585 & -0.399 &  0.060 &  0.183 &  0.294 &  0.579 &  0.700 \\
+$\mu^{sect, com}$      & -0.240 & 0.216 & -0.788 & -0.629 & -0.394 & -0.247 & -0.123 &  0.246 &  0.495 \\
+$\mu^{sect, res} $       & -0.242 & 0.216 & -0.789 & -0.628 & -0.395 & -0.250 & -0.122 &  0.244 &  0.493 \\
+$\mu^{sect, gov} $       & -0.239 & 0.216 & -0.782 & -0.625 & -0.392 & -0.245 & -0.120 &  0.250 &  0.501 \\
+$\mu^{sect, npr}$       & -0.240 & 0.216 & -0.786 & -0.629 & -0.396 & -0.249 & -0.121 &  0.250 &  0.498 \\
+$\beta^fy$   & -0.003 & 0.001 & -0.006 & -0.004 & -0.003 & -0.003 & -0.002 & -0.001 &  0.000 \\
+$\beta_{own}^{cost}$ & -0.000 & 0.001 & -0.002 & -0.002 & -0.001 & -0.000 &  0.000 &  0.001 &  0.002 \\
+$\beta_{lease}^{cost}$ & -0.000 & 0.001 & -0.003 & -0.002 & -0.001 & -0.000 &  0.000 &  0.002 &  0.003 \\
+$\beta^{size}$ & -0.007 & 0.001 & -0.011 & -0.009 & -0.008 & -0.007 & -0.006 & -0.005 & -0.004 \\
 \bottomrule
 \end{tabular}
-\label{table:higher_level_params_const}
-\caption{Summary statistics of the estimated posterior distributions of higher level parameters when higher level mean parameters are constrained to have a maximum of zero.}
+\caption{\label{table:higher_level_params_const}Summary statistics of the estimated posterior distributions of higher level parameters when $b_s$ parameters are constrained to have a maximum of zero.}
 \end{table}
@@
   \includegraphics[width=1\linewidth]{figures/mu_sectors.png}
-  \label{fig:mu_sectors}
+
  \end{minipage}\qquad
@@
   \includegraphics[width=1\linewidth]{figures/mu_sectors_const.png}
-  \label{fig:mu_sectors_const}
+
  \end{minipage}
@@
  \caption{Distribution of the sector contrasts relative to residential sector.
-Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
+Installations for commercial and non-profit hosts do not appear to have on average significant differences in quality compared to residential installations.
+There is estimated approximately 90 percent probability that governmental installations have a higher quality than residential installations}
+   \label{fig:mu_sectors}
 \end{minipage}\qquad
@@
 \centering
- \caption{Distribution of the sector contrasts relative to residential sector when all higher level mean parameters are constrained to have a maximum of zero.
-Installations for commercial, governmental and non-profit hosts are shown to have on average lower quality then installations in residential}
+ \caption{Distribution of the sector contrasts relative to residential sector when all higher level mean parameters are constrained to have a maximum of zero. Results are not materially changed.}
+ \label{fig:mu_sectors_const}
 \end{minipage}
@@
 
-Summary statistics for the parameters on the manufacturer group means as well as the lower-level intercept and slope parameters $a_s$ and $b_s$ are too numerous to present in a table.
+Summary statistics for the parameters on the manufacturer group means as well as the lower-level intercept and slope parameters $a_s$ and $b_s$ are impractical to present in a table.
 Visual summaries are presented in figures \ref{fig:BayManPlot}, \ref{fig:Bay_as} and \ref{fig:Bay_bs}.
-Figure \ref{fig:BayManPlot} shows box plots of the manufacturer distributions, with the box representing 70 percent of probability and the lines representing 95 percent of the probability.
+Figure \ref{fig:BayManPlot} shows line plots of the manufacturer distributions, with the lines representing 95 percent intervals of the probability distribution.
+Figure \ref{fig:BayManPlot_const} shows results from the constrained model.
 Mirroring earlier results, the plot shows most manufacturers with roughly equivalent quality, with a handful with meaningfully lower quality.
@@
 The intercept terms are shown to have considerable variation between systems, as would be expected, but are individually estimated precisely.
+Figures \ref{fig:Bay_as_const} and \ref{fig:Bay_bs_const} show results from the constrained model.
 
-The slope parameters, $b_s$ are modeled explicitly and are also allowed to have varying variance parameters.
+The slope parameters, $b_s$ are modelled explicitly and are also allowed to have varying variance parameters.
 This is evident in the plot.
@@
   \includegraphics[width=1\linewidth]{figures/BayManPlot.png}
-  \label{fig:BayManPlot}
  \end{minipage}\qquad
@@
   \includegraphics[width=1\linewidth]{figures/BayManPlot_const.png}
-  \label{fig:BayManPlot_const}
  \end{minipage}
@@
 \centering
-  \caption{Box plot of the the parameters $\mu^{manuf}$, representing the distributions of the mean parameter for manufacturer groupings.}
+  \caption{Plot of the the parameters $\mu^{manuf}$, representing the distributions of the mean parameter for manufacturer groupings.}
+   \label{fig:BayManPlot}
 \end{minipage}\qquad
@@
 \centering
-  \caption{Box plot of the the parameters $\mu^{manuf}$, when all higher level mean parameters have a maximum of zero.}
-\end{minipage}
-\end{figure}
-
-%new image
-
-\begin{figure}
-\begin{minipage}{.45\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{figures/BayCountyPlot.png}
-	\label{fig:BayCountyPlot}
- \end{minipage}\qquad
-\begin{minipage}{.45\textwidth}
-  \centering
-  \includegraphics[width=1\linewidth]{figures/BayCountyPlot_const.png}
-  \label{fig:BayCountyPlot_const}
- \end{minipage}
-
-\bigskip
-
-\begin{minipage}[t]{.45\textwidth}
-\centering
-  \caption{Box plot of the the parameters $\mu^{county}$, representing the distributions of the mean parameter for county groupings.}
-\end{minipage}\qquad
-\begin{minipage}[t]{.45\textwidth}
-\centering
-  \caption{Box plot of the the parameters $\mu^{county}$, when all higher level mean parameters have a maximum of zero.}
+  \caption{Plot of the the parameters $\mu^{manuf}$, when all higher level mean parameters have a maximum of zero.}
+    \label{fig:BayManPlot_const}
 \end{minipage}
@@
   \includegraphics[width=1\linewidth]{figures/Bay_as.png}
-  \label{sfig:Bay_as}
  \end{minipage}\qquad
@@
   \includegraphics[width=1\linewidth]{figures/Bay_as_const.png}
-  \label{sfig:Bay_as_const}
  \end{minipage}
@@
   \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$.}
+    \label{fig:Bay_as}
 \end{minipage}\qquad
@@
   \caption{95\% interval of the distributions of the system level intercept parameters, $a_s$ when higher level mean parameters are constrained to have a maximum of zero.}
+   \label{fig:Bay_as_const}
 \end{minipage}
@@
   \includegraphics[width=1\linewidth]{figures/Bay_bs.png}
-  \label{sfig:Bay_bs}
  \end{minipage}\qquad
@@
   \includegraphics[width=1\linewidth]{figures/Bay_bs_const.png}
-  \label{sfig:Bay_bs_const}
  \end{minipage}
@@
  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
+  \label{fig:Bay_bs}
 \end{minipage}\qquad
@@
 \centering
-  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$}
+  \caption{95\% interval of the distributions of the system level slope parameter, $b_s$ when distributions are constrained to be at or below 0.}
+   \label{fig:Bay_bs_const}
 \end{minipage}
@@
 
-Historically, the commoditization of goods was a necessary condition for the establishment of sophisticated financial markets for those goods.
-The invention of grain elevators for mixing and thereby standardizing grain was a necessary precursor to the establishment of a futures markets with receipts of delivery traded on the Chicago Board of Trade \citep{cronon_natures_1992}.
-The abstraction between producer and end purchaser of goods was necessary for the functioning of these markets.
-A trader could fulfill a promise to deliver grain of a certain grade by simply buying it on the market or with cash settlement, rather than having produced it oneself.
-
 If solar panels truly have become commodities, it would have meaningful and substantial implications for the solar power industry.
@@
 Yet the findings from this article strongly suggest that solar panels are not commodities.
-In particular, I find that the salient property of quality varies substantially between producers.
-Because solar power is a distributed technology, where assets can be owned by individual homeowners, businesses and other other organizations that do not have the resources to judge quality, the issue of asymmetric information may then enter the market.
+In particular, I find that the salient property of quality can vary substantially between producers.
+Because solar power is a distributed technology, where assets can be owned by individual homeowners, businesses and other organizations that do not have the resources to judge quality, the issue of asymmetric information may be present in the market.
 
@@
 I find that, as the theory would predict, high information owners are able to attain higher-quality panels.
-With about 90 percent probability, I also find that the cost of solar panel systems owned by high-information types is more highly correlated with quality, as theory would also suggest.
+With about 85 percent probability, I also find that the cost of solar panel systems owned by high-information types is more highly correlated with quality, as theory would also suggest.
+It is important to note that while the empirical analysis is consistent with the results predicted by the theory of asymmetric information, asymmetric information is not perfectly identified.
+The results could plausibly be driven by other factors.
+For example, the different ownership models could induce a moral-hazard type behavior when it comes to maintenance of the solar panels, which in turn could effect production over time.
 
@@
 One of the advantages of the linear mixed-effects models and the hierarchical models is that we can also get an idea of the degree of differentiation in quality between different manufacturers.
-It appears that a relatively small tail of producers accounted for the worst quality panels, while the majority of panels had similar quality.
-The market for solar panels is still relatively young, and if low-quality producers either are forced to exit the market or improve their quality to match standards in the industry, then it is possible that solar panels could indeed become commodities in the future, with associated implications for pricing, competition and financing.
+About 20\% of producers accounted for the worst quality panels, while the majority of panels had similar quality.
+If low-quality producers either are forced to exit the market or improve their quality to match standards in the industry, then it is possible that solar panels could indeed become commodities in the future.
 
-\bibliographystyle{chicago}
+However, optimism should be tempered by both theory and experience from other industries.
+Markets that are subject to informational problems have a tendency to experience stasis.
+As \citet{akerlof_market_1970} notes, poor quality producers will tend to push out high quality producers in the presence of asymmetric information.
+In a long-term study of the used car market in Switzerland, \citet{emons_market_2009} find evidence of sustained quality uncertainty among buyers.
+Evidence of long-term quality uncertainty is also found in the market for business aircraft \citep{gilligan_lemons_2004}.
+The market for solar panels may long be characterised by producers differentiated by quality.
+In consequence, the buyers of solar panels will need to take steps to ensure high quality such as paying a premium for panels from a reputed producer or making use of engineering firms specialising in certifying panel quality.
+
+The use of Bayesian MCMC simulation techniques in this article also provides an illustrative example of the use and usefulness of a powerful emerging tool-set.
+The Bayesian frameworks and recently developed software such as Stan provide major practical and theoretical advantages over many traditional statistical methods used in operations research and decision science.
+These methods and tools are also increasingly accessible to practitioners.
+\citet{kruschke_doing_2014}, \citet{mcelreath_statistical_2015}, \citet{gelman_bayesian_2013} and \citet{stan_development_team_stan_2014} provide accessible and thorough introductions and guides.
+
+This article should also be considered as part of a broader literature on the special characteristics of new distributed energy generation technologies.
+The energy investment behavior of home owners, farmers, and small cooperatives are bound to be substantially different than those of large, specialized energy companies that have traditionally done most of the investment in electricity generation.
+Understanding the quickly evolving energy and power industry requires taking into account informational and behavioral factors, and this is a rich field for further research in the field of operations research.
+
+\bibliographystyle{apa}
 \bibliography{solar_prod}

### 2019-07-02 ejor_submission/R&R2/rr2_mauritzenJune19.tex
Final push after acceptance

--- 2019-07-02 ejor_submission/R&R/revision_mauritzenApril19.tex
@@
 \documentclass[a4paper]{article}
+
+%%Version sent in for second revision
 
@@
 
+Solar power has been the fastest growing source of energy generation in recent decades\footnote{\url{https://www.iea.org/topics/renewables/solar/}}, and a large and growing literature within operations research addresses the challenges associated with this technology.
+One branch of the literature considers the operation of electricity markets and transmission planning with large shares of solar and other intermittent energy (see for example \citet{gianfreda_stochastic_2018}, \citet{pritchard_single_settlement_2010}, or \citet{morales_electricity_2014}).
+This article, however, contributes to the branch of the literature that concerns investment decisions, procurement and supply chains for solar power.
+A popular approach has been designing mulitcriteria decision models (MCDM) for solar investment and planning decisions.
+Here, multicriteria takes on  a range of meanings.
+For example it can mean modelling dueling objectives between governments and developers \citep{garcia-bernabeu_photovoltaic_2016}.
+Alternatively, it can refer to the decision of a planner who must balance financial considerations with environmental and social factors \citep{georgopoulou_multicriteria_1997, henao_multicriteria_2012,linares_multiple_2000}, or analysing how multiple sources of risk, such as prices and weather, inform the design decisions of a photovoltaic plant \citep{merzifonluoglu_photovoltaic_2018}.
+Finally, and most closely related to the subject of this article, a multicriteria decision model can refer to the decision of choosing a certain type of solar panel.
+\citet{socorro_garcia-cascales_evaluation_2012}, for example, analyses the decision of choosing panel characteristics such as monocrystalline vs.
+polycrystalline or alternatively, various thin-film type panels.
+The criteria used for choosing a certain type of cell include manufacturing cost, efficiency, and emissions of greenhouse gases during manufacturing.
+Notably, however, an analysis of procuring solar panels of uncertain quality and reliability does not exist in the literature, to the best of my knowledge.
+
 Solar panels have been referred to colloquially as being commodities.\footnote{See for example \url{https://www.greentechmedia.com/articles/read/are-pv-modules-a-commodity\#gs.lx3tMbY}} This is often in reference to the high degree of competition, especially from Chinese manufacturers, that has dramatically driven down prices of solar panels in the last decade.
@@
 Schäffle distinguished between a popular understanding of a commodity as a material input in manufacturing, to the scientific understanding as "exchangeable material goods" \citep{menger_principles_2007}.
-Despite the simplicity of this definition, \citet{gordon_what_1999} show the difficulty in practice of determining whether a product is in fact a commodity.
+In a more recent, but similar definition, \citet{reimann_toward_2010} defines the concept of commoditization as ``occuring when competitors in comparatively stable industries offer increasingly homogenous products.''  Despite the simplicity of these definitions, \citet{gordon_what_1999} show the difficulty in practice of determining whether a product is in fact a commodity.
 The root problem is that a good can be differentiated in a host of ways, some of which are difficult to observe at the time of transaction.
+
+An important differentiating characteristic between solar panels is the quality of these solar panels.
+This can be measured as both the failure rate of these panels as well as the degree of degradation of the panels over time.
+This definition is in line with the general understanding of quality within operations research and economics as "performance quality" \citep{karmarkar_quality_1997}.
+If there were significant differences in quality between manufacturers, then this has implications for the structure and development of the solar panel market.
+Manufacturers with superior quality could differentiate themselves and extract a premium in the market .
+A reputation for quality among established manufacturers could also act as a barrier to entry, thereby reducing long-term pricing pressure \citep{karaer_quality_2015}.
 
@@
 
-An important differentiating characteristic between solar panels is the quality of these solar panels.
-This can be measured as both the failure rate of these panels as well as the degree of degradation of the panels over time.
-If there were significant differences in quality between manufacturers, then this has implications for the structure and development of the solar panel market.
-Manufacturers with superior quality could differentiate themselves and extract a premium in the market.
-A reputation for quality among established manufacturers could also act as a barrier to entry, thereby reducing long-term pricing pressure.
-
-From the perspective of a purchasing manager responsible for procuring solar panels, whether panels can be considered a commodity is of great importance.
+From the perspective of a purchasing or supply chain manager responsible for procuring solar panels, whether panels can be considered a commodity is of great importance.
 If panels are known to be commodities with standardised quality and similar pricing, then the decision for purchasing manager is substantially simplified.
@@
 This eliminates repeat buying as a mechanism for ensuring quality.
-In the literature, warranties are often suggested as a strong signal of quality.
+In the literature, warranties are often suggested as a strong signal of quality \citet{murthy_product_2004, murthy_product_2006}.
 However, warranties may be a relatively weak assurance of quality in the market for solar panels as both contractors and manufacturers are relatively new, tend to be heavily indebted and some have recently shown a tendency to go bankrupt.
@@
 
-Early theoretical work of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller--in other words the seller can choose the level of quality of its product.
+Issues of asymmetric information of quality in supply chains and purchasing manager decisions have become prominent in the operations research literature.
+\citet{quigley_supplier_2018}, noting that validating the quality of suppliers is expensive, develop a model to optimally choose an investment level for ensuring quality of supplies.
+\citet{nagurney_spatial_2014} model a situation with information asymmetry of quality with multiple supply and demand markets, where spatial distance is correlated to information asymmetry.
+This model was motivated by an increasingly globalised supply chain, which of course also describes the sourcing of solar panels.
+
+\citet{bakshi_signaling_2015} develop a game theoretic model of a market where reliability is uncertain, and the seller holds private information.
+The authors show how service contracts can be used as credible signals of product quality.
+\citet{gumus_supply_side_2012} analyses a similar dynamic, where a buyer must choose between a reliable but more expensive supplier and a less reliable but cheaper supplier who also holds private information about quality.
+Here a guarantee is used as a signal of quality.
+Interestingly, in relation to this article, a spot market is allowed to exist for the underlying good, thus implying that the underlying good is a commodity.
+Reliability here is then, presumably, associated not with the good but the supplier.
+Other research in the operations research of procurement with spot- or other financial markets often assume the existence of such markets \citep{zhao_demand_2015, li_buy_2018}.
+But to my knowledge there lacks research analysing the underlying conditions necessary for the formation of such financial markets in the operations research literature.
+In the context of this article, one of those conditions would be that the traded good is a commodity with a certain uniform quality level.
+
+Early theoretical work in economics of particular relevance to this paper includes \citet{chan_prices_1982}, whom model the situation where information on quality is costly rather than directly unavailable and where quality is endogenous to the seller--in other words the seller can choose the level of quality of its product.
 Under these assumptions, the authors predict several equilibrium--essentially markets for both high- and low-quality products depending on the extent to which the buyer is informed.
